---
title: "12.3 coral rarefactions"
author: "TG"
date: "20 12 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(iNEXT)
library(rareNMtests)

```


```{r}
load("knoll_mdata.rdata")
load("coral_cover.rdata")
load("coral_comm.rdata")
```

```{r}
coral_survey_list<-coral_cover %>% group_by(Site,Year) %>% summarise(knoll_number=n_distinct(Knoll))

#write.csv(coral_survey_list,"coral_survey_list.csv")
```


```{r}
stone_coral<- coral_comm %>%
  filter(Category %in% c("Stony Coral","Hydrozoa"))


rare_coral<-stone_coral %>% group_by(site,knoll,year,storm,species) %>% summarise("abundance"=sum(points))

data<-spread(rare_coral,species,abundance,fill = 0)


data <- data %>% filter(year != "2017")

species_data<-data[,5:ncol(data)]
site_data<-data[,1:4]

site_data$Site_Knoll<-paste(site_data$site,site_data$knoll,sep="_")

species_data<-as.data.frame(species_data)
```

vegan - group individual 

```{r}

#create the data

stone_coral<- coral_comm %>%
  filter(Category %in% c("Stony Coral","Hydrozoa"))


rare_coral<-stone_coral %>% group_by(site,knoll,year,storm,species) %>% summarise("abundance"=sum(points))

ibr<-rare_coral %>% group_by(site,species) %>% summarise("abundance"=sum(abundance))

ibr<-spread(ibr,species,abundance,fill = 0)

#playing a bit with the vegan functions

ibr_max<-min(rowSums(ibr[,2:ncol(ibr)]))

rarefy(ibr[,2:ncol(ibr)],ibr_max,se=TRUE)


rare_index_steps<-rarefy(ibr[2:ncol(ibr)],se=T,sample=100)

site_name<-ibr$site
colors<-c("black","blue","red","pink")

rareplot_per_site<-rarecurve(ibr[,2:ncol(ibr)],step=20,sample =ibr_max,se=T,label=F) 
legend("bottomright",legend = site_name, lty=c(1,1))



```


individual rarefaction curve with confidence interval
pooled to the site level


```{r}

steps <- data.frame("ind" =  seq(10, 15875, by=10)) 
rare_list<-list()
loop<-1

for(i in 1:nrow(steps)){
  
  ind <- steps$ind[i]
  
 S<-rarefy(ibr[,2:ncol(ibr)], ind, se = TRUE)
 S<-t(S)
 S<-as.data.frame(S)
 S$Site<-site_name
 S$ind<-rep(ind)

 rare_list[[loop]]  <- S
 
 loop<-loop+1
  
}

rarefection<-bind_rows(rare_list)

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")


ggplot(data = rarefection,aes(x=ind,y=S,color=Site))+geom_line(size=1.5)+
   geom_ribbon(aes(ymin=S-se, ymax=S+se),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Individuals number") + ggtitle("Stony corals")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


#ggsave("hard_coral_ibr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")



```

```{r}
caves<-data %>% filter(site=="Caves")

rarefaction.sample(caves[,5:ncol(caves)], method = "sample-size", q = 0)




```

sample base rarefaction with ci (rarefy version) 


```{r}

srb_list<-list()
loop<-1

for (i in unique(data$site)) {
  
  site<- data %>% filter(site == i)
  
  site_richness<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=0,
                          random=999,
                          mean = F)
  
  site_richness$site<-rep(i)
  site_richness$sample<-1:nrow(site_richness)
  
  
  srb_list[[loop]]<-site_richness
  loop<-loop+1
  
  
}

site_richness<-bind_rows(srb_list)


ggplot(site_richness,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stone coral")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


#ggsave("hard_coral_sbr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")

```

plot for assaf 
add all the sites together to the plot

```{r}

all_site_rare <- Rarefy::rare_alpha(data[5:ncol(data)],
                          method="hill",
                          q=0,
                          random=999,
                          mean = F)

all_site_rare$sample<-1:nrow(all_site_rare)



ggplot(site_richness,
       aes(x= sample,y=Rarefaction,color=site))+
  geom_line(size=1.2)+
  geom_line(data = all_site_rare,
            aes(x=sample,y=Rarefaction),
            color="black",size=1.2)+
  theme_bw()+
  xlab("Samples") + 
  ggtitle("Hard coral")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(name = "Site",
    breaks = c("Caves",
               "Katza",
               "NR",
               "Princess",
               "All sites"),
    values = c("Caves"="#56dde0",
               "Katza"="#33658a",
               "NR"="#f5af2d",
               "Princess"= "#f2641a",
               "All sites"="black"))

ggsave("hard_coral_sbr_with_all_site.png", width = 14, height = 5, dpi = 300)

```


#sbr for stone coral by year


```{r}

l_type<-c("solid","solid","solid","twodash")

srb_list_year<-list()
loop<-1
all_new<-c("#98c1d9","#6b8ead","#003049","#c1121f")


for (i in unique(data$year)) {
  
  year<- data %>% filter(year == i)
  
  year_richness<-Rarefy::rare_alpha(year[5:ncol(year)],
                          method="hill",
                          q=0,
                          random=999,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  srb_list_year[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness<-bind_rows(srb_list_year)

year_richness$Year<-as.factor(year_richness$Year)

ggplot(year_richness,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year, linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Year",values = l_type) +
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stony coral")+
  ylab("Species richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = all_new)

ggsave("hard_coral_sbr_year.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")

```

#shanon diversity by site

```{r}

srb_shn_list<-list()
loop<-1

for (i in unique(data$site)) {
  
  site<- data %>% filter(site == i)
  
  site_shanon<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=1,
                          random=999,
                          mean = F)
  
  site_shanon$site<-rep(i)
  site_shanon$sample<-1:nrow(site_shanon)
  
  
  srb_shn_list[[loop]]<-site_shanon
  loop<-loop+1
  
  
}

site_shanon<-bind_rows(srb_shn_list)


ggplot(site_shanon,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Stony corals")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("hard_coral_shanon_sbr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")



```
# shannon sbr for stone coral by year


```{r}
srb_shanon_list_year<-list()
loop<-1
all_new<-c("#98c1d9","#6b8ead","#003049","#c1121f")


for (i in unique(data$year)) {
  
  year<- data %>% filter(year == i)
  
  year_richness<-Rarefy::rare_alpha(year[5:ncol(year)],
                          method="hill",
                          q=1,
                          random=999,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  srb_shanon_list_year[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness_sh<-bind_rows(srb_shanon_list_year)

year_richness_sh$Year<-as.factor(year_richness_sh$Year)

ggplot(year_richness_sh,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year, linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Year",values = l_type) +
  theme_bw()+xlab("Random sample number") + ggtitle("Stony coral")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = all_new)

ggsave("hard_coral_sbr_year_shanon.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")

```
soft corals

```{r}

soft_coral<- coral_comm %>%
  filter(Category %in% c("Soft Coral","Zoantharia"))


rare_soft<-soft_coral %>% group_by(site,knoll,year,storm,species) %>% summarise("abundance"=sum(points))

data_soft<-spread(rare_soft,species,abundance,fill = 0)


data_soft <- data_soft %>% filter(year != "2017")





```


ind dased

```{r}
ibr_soft<-rare_soft %>% group_by(site,species) %>% summarise("abundance"=sum(abundance))

ibr_soft<-spread(ibr_soft,species,abundance,fill = 0)

#playing a bit with the vegan functions

ibr_soft_max<-min(rowSums(ibr_soft[,2:ncol(ibr_soft)]))


steps_s <- data.frame("ind" =  seq(1, 525, by=5)) 
rare_list_soft<-list()
loop<-1

for(i in 1:nrow(steps_s)){
  
  ind <- steps_s$ind[i]
  
 S<-rarefy(ibr_soft[,2:ncol(ibr_soft)], ind, se = TRUE)
 S<-t(S)
 S<-as.data.frame(S)
 S$Site<-site_name
 S$ind<-rep(ind)

 rare_list_soft[[loop]]  <- S
 
 loop<-loop+1
  
}

rarefection_soft<-bind_rows(rare_list_soft)

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")


ggplot(data = rarefection_soft,aes(x=ind,y=S,color=Site))+geom_line(size=1.5)+
   geom_ribbon(aes(ymin=S-se, ymax=S+se),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Individuals number") + ggtitle("Soft corals")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
      plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("soft_coral_ibr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")


```
# Shannon


```{r}
srb_shn_list_soft<-list()
loop<-1

for (i in unique(data_soft$site)) {
  
  site<- data_soft %>% filter(site == i)
  
  site_shanon<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=1,
                          random=999,
                          mean = F)
  
  site_shanon$site<-rep(i)
  site_shanon$sample<-1:nrow(site_shanon)
  
  
  srb_shn_list_soft[[loop]]<-site_shanon
  loop<-loop+1
  
  
}

site_shanon_soft<-bind_rows(srb_shn_list_soft)


ggplot(site_shanon_soft,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Soft corals")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("soft_coral_shanon_sbr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")

```


sample based 

```{r}
srb_soft_list<-list()
loop<-1

for (i in unique(data_soft$site)) {
  
  site<- data_soft %>% filter(site == i)
  
  site_richness<-Rarefy::rare_alpha(site[5:ncol(site)],
                          method="hill",
                          q=0,
                          random=999,
                          mean = F)
  
  site_richness$site<-rep(i)
  site_richness$sample<-1:nrow(site_richness)
  
  
  srb_soft_list[[loop]]<-site_richness
  loop<-loop+1
  
  
}

site_soft_richness<-bind_rows(srb_soft_list)


ggplot(site_soft_richness,aes(x= sample,y=Rarefaction,color=site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Soft coral")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("soft_coral_sbr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```

# Hill numbers


## Stony corals

```{r}
renyi_profile_site<-renyi(ibr[,2:ncol(ibr)],scales = c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf), hill = T)

renyi_data<-data.frame(t(renyi_profile_site))

rownames(renyi_data)=NULL

renyi_data$q<-c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf)

colnames(renyi_data)<-c("Caves","Katza","NR","Princess","Q")

renyi_data<-gather(renyi_data,"Site","richness",1:4)

renyi_data$Q<-ordered(renyi_data$Q,levels = c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf))

ggplot(renyi_data,aes(x=Q,y=richness,color = Site))+
  geom_point(size=2)+
  scale_color_manual(values = color)+
  xlab("Diversity number")+
  ylab("Diversity")+theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  ggtitle("Stony coral")
  

ggsave("hill_profile_stonecoral.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")


```


