---
title: "14.1 knoll based analyses"
author: "TG"
date: "17 12 2021"
output: html_document
---

```{r}
library(tidyverse)
library(ggplot2)
library(wesanderson)
library(RColorBrewer)
library(plotrix)
```

```{r}
load("knoll_mdata.rdata")
load("coral_cover.rdata")
load("coral_comm.rdata")
load("fish_full_data.rdata")

```

# Corals


```{r}
yearly_cover<- coral_cover %>% select(Site,Year,storm,Surface_Area,Knoll,Stony.Coral,Hydrozoa,Soft.Coral,Algae,Zoantharia) %>% mutate(all_stone = Stony.Coral+Hydrozoa) %>% mutate(all_soft = Soft.Coral+Zoantharia)


k_new<-c("#669bbc","#336683","#1a4b66","#003049","#c1121f")
all_new<-c("#669bbc","#336683","#003049","#c1121f")
```


## coral cover %

stony coral and hydrozoa

```{r}

yearly_cover_no_k<-yearly_cover %>% filter(Site != "Katza")

for (i in unique(yearly_cover_no_k$Site)) {
 
test<-yearly_cover_no_k %>% filter(Site == i) %>% drop_na(Stony.Coral)

gp1 <- wes_palettes$GrandBudapest2 
gp2 <- wes_palettes$GrandBudapest1

plot_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\cover per knoll\\stone"

comb_bud<-c(gp1[c(2,3,4)],gp2[4])


p2<-test %>%
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,all_stone, .fun='max' )) %>%
  ggplot( aes(x=Knoll, y=all_stone, color = as.factor(Year))) +   geom_point(size=2) +
  ggtitle(unique(test$Site))+
  xlab("Knoll id") +
  ylab("Stony coral cover (%)")+
  scale_y_continuous(limits = c(0,85),breaks=seq(0,85,by=5))+
  theme_bw()+
  theme(title = element_text(size = 20),
    axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))   +scale_color_manual(values=all_new)+
   labs(color = "Year")

p2_name<-paste("stoney_cover_v2_",unique(test$Site),".png")

ggsave(p2_name, width = 14, height = 5, dpi = 300,path = plot_path)
  

p3<-test %>%
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,Soft.Coral, .fun='max' )) %>%
  ggplot( aes(x=Knoll, y=Soft.Coral, color = as.factor(Year))) +   geom_point(size=2) +
  ggtitle(unique(test$Site))+
  xlab("Knoll id") +
  ylab("Soft coral cover (%)")+
  scale_y_continuous(breaks=seq(0,50,by=1))+
  theme_bw()+
  theme(title = element_text(size = 20),
    axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+          labs(color = "Year")+
  scale_color_manual(values=comb_bud)
 

p3_name<-paste("soft_cover_",unique(test$Site),".png")

p3_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\cover per knoll\\soft"

ggsave(p3_name, width = 14, height = 4, dpi = 300,path = p3_path)

}


```

bar plot of soft and stone corals


```{r}



cover_bar<-yearly_cover %>% group_by(Site,Year) %>% summarise("mean_stone"=mean(all_stone,na.rm =T),
                                                   "mean_soft"=mean(all_soft,na.rm =T),
                                                   "n"=n(),
                                                   "SE_stone"=std.error(all_stone),
                                                   "SE_soft"=std.error(all_soft))


cover_bar<-gather(cover_bar,"coral_type","cover",3:4)


cover_bar$Year<-as.factor(cover_bar$Year)

cover_bar<-cover_bar %>% group_by(Site,Year) %>% 
  mutate("se_positon"= case_when(coral_type =="mean_stone" ~ cover + SE_stone,
                                 coral_type=="mean_soft" ~ cumsum(cover)+SE_soft))

cover_bar<-cover_bar %>% group_by(Site,Year) %>% 
  mutate("se_positon_min"= case_when(coral_type =="mean_stone" ~ cover ,
                                 coral_type=="mean_soft" ~ cumsum(cover)))


ggplot(data = cover_bar,aes(x=Year,y = cover, fill = coral_type))+
    geom_bar(position="stack", stat="identity",color ="black")+
  geom_errorbar(aes(ymax = se_positon,ymin = se_positon_min),width=.1)+
    facet_wrap(Site~.)+
    theme_bw()+ylab("Mean cover(%)")+ 
  geom_text(data = cover_bar,
        aes(label = n, y = 5),size=6)+
  scale_fill_manual(values = c("#fed0bb","#d77a7d"),name = "Coral type",labels= c("Soft coral","Stony coral"))+
    theme(title = element_text(size = 20),
    axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
    strip.text = element_text(size=22),
    strip.background =element_rect(fill="#faf5ef"))
  
#ggsave("coral_cover_stacked_bar.png" , width = 10, height = 8, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021") 
 
```




only kazta (separate for the different color pallet)

```{r}

only_kazta<-yearly_cover %>% filter(Site == "Katza")
test<-only_kazta %>%  drop_na(Stony.Coral)

gp1 <- wes_palettes$GrandBudapest1 
gp2 <- wes_palettes$GrandBudapest2
k_pallert<-c(gp1[c(1,2)],gp2[4],gp1[c(4,3)])


plot_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\cover per knoll\\stone"


p2<-test %>%
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,all_stone, .fun='max' )) %>%
  ggplot( aes(x=Knoll, y=all_stone, color = as.factor(Year))) +   geom_point(size=2) +
  ggtitle(unique(test$Site))+
  xlab("Knoll id") +
  ylab("Stony coral cover (%)")+
  scale_y_continuous(limits = c(0,85),breaks=seq(0,85,by=5))+
  theme_bw()+
  theme(title = element_text(size = 20),
        axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))   +scale_color_manual(values=k_new)+
   labs(color = "Year")

p2_name<-paste("stoney_cover_v2_",unique(test$Site),".png")

ggsave(p2_name, width = 14, height =5, dpi = 300,path = plot_path)
  

#p3<-test %>%
#  mutate(Knoll = as.factor(Knoll)) %>% 
 # mutate(Knoll = fct_reorder(Knoll,Soft.Coral, .fun='max' )) %>%
  #ggplot( aes(x=Knoll, y=Soft.Coral, color = as.factor(Year))) +   geom_point(size=2) +
  #ggtitle(unique(test$Site))+
  #xlab("Knoll id") +
  #ylab("Soft coral cover (%)")+
  #scale_y_continuous(breaks=seq(0,50,by=1))+
  #theme_bw()+
  #theme(axis.text.x = element_text(size=20),
  #      axis.text.y = element_text(size =16),
  #      axis.title.x = element_text(size=22),
  #      axis.title.y = element_text(size=22),
  #      legend.title = element_text(size=22),
  #      legend.text = element_text(size=20))+
  #labs(color = "Year")+
  #scale_color_manual(values=comb_bud)
 

#p3_name<-paste("soft_cover_",unique(test$Site),".png")

#p3_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\cover per knoll\\soft"

#ggsave(p3_name, width = 14, height =5, dpi = 300,path = p3_path)


```

# Coral species richness

only stoney corals and hydorozoa

1. only katza

#669bbc 2015
#336683 2016
#1a4b66 2017
#003049 2018

red

#c1121f 2020

```{r}
stony_richness<-coral_comm %>%
filter(Category =="Stony Coral"|Category =="Hydrozoa") %>% 
  group_by(site,knoll,year,storm) %>% 
  summarise("species" = n_distinct(species))

gp1 <- wes_palettes$GrandBudapest1 
gp2 <- wes_palettes$GrandBudapest2
k_pallert<-c(gp1[c(1,2)],gp2[4],gp1[c(4,3)])
k_new<-c("#669bbc","#336683","#1a4b66","#003049","#c1121f")

r1<-stony_richness %>% filter(site =="Katza") %>% 
  mutate(knoll = as.factor(knoll)) %>% 
  mutate(knoll = fct_reorder(knoll,species,.fun='max')) %>%
  ggplot( aes(x=fct_reorder(knoll,species,.fun='max'), y=species, color = as.factor(year))) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  scale_y_continuous(limits = c(0,30),breaks=seq(0,30,by=5))+
  ggtitle("Katza")+
  xlab("Knoll id") +
  ylab("Stony coral species richness")+
 theme_bw()+
  theme(title = element_text(size = 20),
        axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
   scale_color_manual(values=k_new)+
   labs(color = "Year")
r1  

r1_name<-"stoney_richness_v1_Katza.png"
r1_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\coral richness per knoll"
ggsave(r1_name, width = 14, height =5, dpi = 300,path = r1_path)


```


2. loop the other sites

```{r}

loop_data<-stony_richness %>% filter(site !="Katza")

for (i in unique(loop_data$site)) {
  
gp1 <- wes_palettes$GrandBudapest1 


data<-loop_data %>% filter(site == i) 

r1<-data %>%
  mutate(knoll = as.factor(knoll)) %>% 
  ggplot( aes(x=fct_reorder(knoll,species,.fun='max'), y=species, color = as.factor(year))) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  ggtitle(unique(data$site))+
  scale_y_continuous(limits = c(0,30),breaks=seq(0,30,by=5))+
  xlab("Knoll id") +
  ylab("Stony coral species richness")+
 theme_bw()+
  theme(title = element_text(size = 20),
        axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=all_new)+
   labs(color = "Year")
  
r1_name<-paste("stoney_richness_v1_",unique(data$site),".png")
r1_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\coral richness per knoll"
ggsave(r1_name, width = 14, height =5, dpi = 300,path = r1_path)

  
}


```



## fish richness per knoll

### general steps:

1) filter confidence = 1

for the years 2015-2017 where there is no confidance value i set the value 0

2) remove irregular sites

3) keep one observer

4) keep only residence fish

5) count the species number

6) add sample season


```{r}
#1

sp_area<-fish_full_data
sp_area$Confidence[is.na(sp_area$Confidence)]<-(0)
sp_area<- sp_area %>% filter(Confidence < 1)

#2

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

sp_area<-sp_area %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")

#3

one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(sp_area$survey_id)) {
  
  data <- sp_area %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
# 4

only_cryptic <- one_observer %>% filter(Status != "T")

# 5

fish_richness<- only_cryptic %>% group_by(Site,Knoll,Date,year_month,storm) %>% summarise("species"= n_distinct(Species_2015))

fish_richness$sample_name<-ifelse(lubridate::month(fish_richness$Date)==6,paste("Spring",lubridate::year(fish_richness$Date),sep=" "),paste("Fall",lubridate::year(fish_richness$Date),sep=" "))

fish_richness$sample_name<-as.factor(fish_richness$sample_name)

fish_richness$sample_name<-ordered(fish_richness$sample_name, levels = c("Fall 2015", "Fall 2016", "Fall 2017", "Fall 2018","Spring 2020", "Fall 2020" ,"Spring 2021"))






```


plot Katza

```{r}
fish_katza<-fish_richness %>% filter(Site=="Katza")

gp1 <- wes_palettes$GrandBudapest1 
gp2 <- wes_palettes$GrandBudapest2

reg_pallet<-c("")
k_pallert<-c(gp1[c(1,2)],gp2[4],gp1[c(4,3)])

fish_k<-c("#669bbc","#336683","#1a4b66","#003049","#ed3140","#c1121f","#710912")
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")


f1<-fish_katza %>% 
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,species,.fun='max')) %>%
  ggplot( aes(x=fct_reorder(Knoll,species,.fun='max'), y=species, color = sample_name)) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  ggtitle("Katza")+
  scale_y_continuous(limits = c(0, 25),breaks = seq(0, 25, 5))+
  xlab("Knoll id") +
  ylab("Resident fish richness")+
 theme_bw()+
  theme(title = element_text(size = 20),
    axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=fish_k)+
   labs(color = "Year")
f1


f1_name<-"fish_richness_Katza.png"
f1_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\fish richness per knoll"
ggsave(f1_name, width = 14, height =5, dpi = 300,path = f1_path)





```



for the rest of the sites

```{r}

fish_no_k<-fish_richness %>% filter(Site != "Katza")

for (i in unique(fish_no_k$Site)) {
  
  data<-fish_no_k %>% filter(Site == i)

  f1<-data %>% 
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,species,.fun='max')) %>%
  ggplot( aes(x=fct_reorder(Knoll,species,.fun='max'), y=species, color = sample_name)) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  ggtitle(unique(data$Site))+
  scale_y_continuous(limits = c(0, 25),breaks = seq(0, 25, 5))+
  xlab("Knoll id") +
  ylab("Resident fish richness")+
 theme_bw()+
  theme(title = element_text(size = 20),
       axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=fish_all)+
   labs(color = "Year")
f1


f1_name<-paste("fish_richness_",unique(data$Site),".png")
f1_path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021\\fish richness per knoll"
ggsave(f1_name, width = 14, height =5, dpi = 300,path = f1_path)
    
}



```


# stone coral cover by year


```{r}

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")


mean_cover<-yearly_cover %>% group_by(Site,Year) %>% summarise("mean"=mean(all_stone,na.rm=T),"se" = std.error(all_stone),n=n(),                                                               "ci_upper"=mean+((1.96*se)/sqrt(n)),
"ci_lower"=mean-((1.96*se)/sqrt(n)),
"ci" = (1.96*se)/sqrt(n))


ggplot(data = mean_cover,aes(x=Year,y=mean,color = Site))+geom_point(size=3,position = position_dodge(0.3))+
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), width=.1,position = position_dodge(0.3))+
  scale_color_manual(values = color)+theme_bw()+geom_line(size=1,position = position_dodge(0.3))+
  xlab("Year") +
  ylab("Mean stony coral cover (%)")+
 theme_bw()+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))

ggsave("stone_mean_cover_w_ci.png", width = 14, height =5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")


```


```{r}

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")




mean_soft_cover<-yearly_cover %>% group_by(Site,Year) %>% summarise("mean"=mean(Soft.Coral,na.rm=T),"se" = std.error(Soft.Coral),n=n(),                                                               "ci_upper"=mean+((1.96*se)/sqrt(n)),
"ci_lower"=mean-((1.96*se)/sqrt(n)),
"ci" = (1.96*se)/sqrt(n))


ggplot(data = mean_soft_cover,aes(x=Year,y=mean,color = Site))+geom_point(size=3,position = position_dodge(0.3))+
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper,color = Site), width=.1,position = position_dodge(0.3))+
 theme_bw()+geom_line(size=1,position = position_dodge(0.3))+
  xlab("Year") +
  ylab("Mean soft coral cover (%)")+
 theme_bw()+scale_y_continuous(limits = c(0,14),breaks = seq(0,14,by=2))+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("soft_mean_cover_w_ci.png", width = 14, height =5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```