---
title: "16.1 zeta diversity"
author: "TG"
date: "10 1 2022"
output: html_document
---


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(Rarefy)
library(rareNMtests)
library(zetadiv)
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
```

```{r}
load("fish_full_data.rdata")
```

1) set one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 1

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <1)

```

3) filter the north beach
```{r}
one_observer<-one_observer %>%
  filter(Site !="North beach",
         ! Knoll %in% c("Tamar","South bridge reserve","south bridge reserve",
                        "North bridge reserve","north bridge reserve"))
                                      
```



4) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```

# lets create the data for one site at one year

NR 2018, convert to wide foramt and change to presence absence


```{r}

nr_2018<-one_observer %>%
  filter(Site =="NR",Year =="2018") %>% 
  select(Knoll,DD_X,DD_Y,Species,Abundance) %>%
  mutate(Knoll = as.numeric(Knoll)) %>% 
  group_by(Knoll,DD_X,DD_Y,Species) %>% summarise(Abundance=sum(Abundance)) %>% arrange(Knoll)

nr_2018<-spread(nr_2018,Species,Abundance,fill = 0)
nr_2018<-as.data.frame(nr_2018)

nr_2018_species<-nr_2018[,4:ncol(nr_2018)]
nr_2018_species[nr_2018_species>0]<-1



nr_2018_coord<-nr_2018[,2:3]


nr_2018_coord<-nr_2018_coord %>% rename(x=DD_X,y=DD_Y)


nr_2018_species<-as.data.frame(nr_2018_species)
nr_2018_coord<-as.data.frame(nr_2018_coord)
```



lets try the function

```{r}
zeta_nr_18<-Zeta.decline.mc(nr_2018_species,
                nr_2018_coord,
                orders = 1:17,
                sam = 100,
                NON = T)
```

```{r}

nr_2021<-one_observer %>%
  filter(Site =="NR",Year =="2021") %>% 
  select(Knoll,DD_X,DD_Y,Species,Abundance) %>%
  mutate(Knoll = as.numeric(Knoll)) %>% 
  group_by(Knoll,DD_X,DD_Y,Species) %>% summarise(Abundance=sum(Abundance)) %>% arrange(Knoll)

nr_2021<-spread(nr_2021,Species,Abundance,fill = 0)


nr_2021_species<-nr_2021[,4:ncol(nr_2021)]
nr_2021_species[nr_2021_species>0]<-1



nr_2021_coord<-nr_2021[,2:3]


nr_2021_coord<-nr_2021_coord %>% rename(x=DD_X,y=DD_Y)


nr_2021_species<-as.data.frame(nr_2021_species)
nr_2021_coord<-as.data.frame(nr_2021_coord)
```



lets try the function

```{r}
zeta_nr_21<-Zeta.decline.mc(nr_2021_species,
                nr_2021_coord,
                orders = 1:17,
                sam = 100,
                NON = T)
```

try to plot toghter

```{r}
nr_zeta<-data.frame("z_order"=zeta_nr_18[["zeta.order"]],
                       "zeta_value_2018"=zeta_nr_18[["zeta.val"]],
                       "sd_18"=zeta_nr_18[["zeta.val.sd"]],
                       "zeta_value_2021"=zeta_nr_21[["zeta.val"]],
                       "sd_21"=zeta_nr_21[["zeta.val.sd"]])

nr_zeta<-gather(nr_zeta,"Year","Value",2:3)


ggplot(data =nr_zeta,aes(x=z_order,y=Value,color =Year) )+geom_point()+geom_line()
```


write a loop that do this for each site and each year



```{r}



one_observer$site_sample<-paste(one_observer$Site,
                                one_observer$Year,                          month(one_observer$Date),
                                sep=" ")

one_observer_spatial<-one_observer %>% drop_na(DD_X)

one_observer_spatial<-one_observer_spatial %>% filter(Year >2017)

zeta_data_list<-list()
zeta_list<-list()
aic_list<-list()
loop<-1

for (i in unique(one_observer_spatial$site_sample)) {
  
  one_data <- one_observer_spatial %>% filter(site_sample==i)
  
  t_Year<-unique(one_data$Year)
  t_year_month<-unique(one_data$year_month)
  t_month<-unique(month(one_data$Date))
  t_Site<-unique(one_data$Site)
  
  
  one_zeta_data <- one_data %>%
  select(Knoll,DD_X,DD_Y,Species,Abundance) %>%
  mutate(Knoll = as.numeric(Knoll)) %>% 
  group_by(Knoll,DD_X,DD_Y,Species) %>% 
    summarise(Abundance=sum(Abundance)) %>%    arrange(Knoll)

one_zeta_data<-spread(one_zeta_data,
                      Species,
                      Abundance,
                      fill = 0)

one_species<-one_zeta_data[,4:ncol(one_zeta_data)]
one_species[one_species>0]<-1
one_species<-as.data.frame(one_species)


one_coord<-one_zeta_data[,2:3]
one_coord<-one_coord %>% rename(x=DD_X,y=DD_Y)
one_coord<-as.data.frame(one_coord)

zeta <-Zeta.decline.mc(one_species,
                one_coord,
                orders = 1:nrow(one_species),
                sam = 100,
                NON = T)

data<-data.frame("z_order"=zeta[["zeta.order"]],
                       "zeta_value"=zeta[["zeta.val"]],
                       "zeta_sd"=zeta[["zeta.val.sd"]],
                 "zeta_ratio"=c(NA,zeta[["ratio"]]),
                 "Site"=rep(t_Site),
                 "Year"= rep(t_Year),
                 "year_month"=rep(t_year_month),
                 "month"=rep(t_month),
                 "sample_id"=rep(i))


 aic_table<-as.data.frame(zeta[["aic"]])
 aic_table$Site<-rep(t_Site)  
 aic_table$year_month<- rep(t_year_month)
 aic_table$method<-rownames(aic_table)
 
 aic_table$method[aic_table$method=="zeta$zeta.exp"]<-"exp"

 aic_table$method[aic_table$method=="zeta$zeta.pl"]<-"plog" 
 
# calculte the delta and keep only the best model
 
 aic_table <- aic_table %>% mutate("delta_aic"= min(AIC)-max(AIC)) %>% filter(AIC == min(AIC))
   
 aic_list[[loop]]<-aic_table
 zeta_list[[i]]<-zeta
 zeta_data_list[[loop]]<-data


loop<-loop+1

  
  
}
  
  
zeta_divesity<-bind_rows(zeta_data_list)

aic_table<-bind_rows(aic_list)

ggplot(data =zeta_divesity,
       aes(x=z_order,y=zeta_value,color =year_month))+
  geom_line()+
  geom_point()+
  facet_wrap(Site ~ .,nrow = 2,ncol=2)+
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, by = 1))+
  theme_bw()+xlab("Zeta order")+ylab("Zeta diversity")


ggplot(data =zeta_divesity,
       aes(x=z_order,y=zeta_ratio,color =year_month))+
  geom_line()+
  geom_point()+
  facet_wrap(Site ~ .,nrow = 2,ncol=2)+
  theme_bw()+xlab("Zeta order")+ylab("Zeta ratio")



for (i in unique(zeta_divesity$Site)) {
  
  zeta_plot<-zeta_divesity %>% filter(Site==i)
  
  plot<-ggplot(data =zeta_plot,aes(x=z_order,y=zeta_value,color =year_month) )+geom_line()+geom_point(size=2,position = position_dodge(0.3))+ theme_classic()+xlab("Zeta order")+ylab("Zeta diversity")+geom_errorbar(aes(ymin = zeta_value-zeta_sd,ymax = zeta_value+zeta_sd),width = 0.4,alpha = 0.5,position = position_dodge(0.3))+ggtitle(i)+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        plot.title = element_text(size=24))
  
 print(plot)
  
}


```

# all sites toghther diffrent years 

```{r}

all_site <- one_observer_spatial %>% 
  select(site_sample,Site,year_month,DD_X,DD_Y,Species,Abundance) %>%
  group_by(site_sample,Site,year_month,DD_X,DD_Y,Species) %>% 
    summarise(Abundance=sum(Abundance))

all_site_wide <- spread(all_site,
                      Species,
                      Abundance,
                      fill = 0)



zeta_data_all_list<-list()
zeta_all_list<-list()
aic_list_all<-list()
loop_all<-1

for (i in unique(all_site_wide$year_month)) {
  
  one_data <- all_site_wide %>% filter(year_month == i )
  
  t_year_month<-unique(one_data$year_month)
  t_Site<-unique(one_data$Site)
  
  
one_species<-one_data[,6:ncol(one_data)]
one_species[one_species>0]<-1
one_species<-as.data.frame(one_species)


one_coord <- one_data[,4:5]
one_coord <- one_coord %>% rename(x=DD_X,y=DD_Y)
one_coord <- as.data.frame(one_coord)

zeta <- Zeta.decline.mc(one_species,
                one_coord,
                orders = 1:nrow(one_species),
                sam = 100,
                NON = T)


data <- data.frame("z_order"=zeta[["zeta.order"]],
                       "zeta_value"=zeta[["zeta.val"]],
                       "zeta_sd"=zeta[["zeta.val.sd"]],
                 "zeta_ratio"=c(NA,zeta[["ratio"]]),
                 "year_month"=rep(t_year_month))


 aic_table<-as.data.frame(zeta[["aic"]])
 aic_table$year_month<- rep(t_year_month)
 aic_table$method<-rownames(aic_table)
 
 aic_table$method[aic_table$method=="zeta$zeta.exp"]<-"exp"

 aic_table$method[aic_table$method=="zeta$zeta.pl"]<-"plog" 
 
# calculte the delta and keep only the best model
 
 aic_table <- aic_table %>% mutate("delta_aic"= min(AIC)-max(AIC)) %>% filter(AIC == min(AIC))
   
 aic_list_all[[loop_all]]<-aic_table
 zeta_all_list[[i]]<-zeta
 zeta_data_all_list[[loop_all]]<-data


loop_all<-loop_all+1

  
  
}
  
zeta_divesity_all<-bind_rows(zeta_data_all_list)

aic_table_all<-bind_rows(aic_list_all)

ggplot(data =zeta_divesity_all,
       aes(x=z_order,y=zeta_value,color =year_month))+
  geom_line()+
  geom_point()+
  theme_bw()+xlab("Zeta order")+ylab("Zeta diversity")+
  geom_errorbar(aes(ymin = zeta_value-zeta_sd,ymax = zeta_value+zeta_sd),width = 0.2,alpha = 0.6)+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        plot.title = element_text(size=24))+
  scale_x_continuous(limits = c(0, 15), breaks = seq(0,15, by = 5))+
  scale_color_manual(values = fish_all[3:6],name = "Year",
                           labels = c(                                  "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))

#ggsave("zeta_2018.pdf" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")

#ggsave("zeta_2018.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")
  
ggplot(data =zeta_divesity_all,
       aes(x=z_order,y=zeta_ratio,color =year_month))+
  geom_line()+
  geom_point()+
  theme_bw()+xlab("Zeta order")+ylab("Zeta ratio")+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        plot.title = element_text(size=24))+
    scale_color_manual(values = fish_all[3:6],name = "Year",
                           labels = c(                                  "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))
#ggsave("zeta_ratio_2018.pdf" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")

#ggsave("zeta_ratio_2018.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")


```


# all sites toghther diffrent years - but with 2015 data

```{r}

all_site_2015 <- one_observer %>% 
  select(site_sample,Site,year_month,DD_X,DD_Y,Species_2015,Abundance) %>%
  group_by(site_sample,Site,year_month,DD_X,DD_Y,Species_2015) %>% 
    summarise(Abundance=sum(Abundance))

all_site_2015_wide <- spread(all_site_2015,
                      Species_2015,
                      Abundance,
                      fill = 0)


all_site_2015_wide<-all_site_2015_wide %>% filter(year_month!="2017 b")


zeta_data_all_list_15<-list()
zeta_all_list_15<-list()
aic_list_all_15<-list()
loop_all_15<-1

for (i in unique(all_site_2015_wide$year_month)) {
  
  one_data <- all_site_2015_wide %>% filter(year_month == i )
  
  t_year_month<-unique(one_data$year_month)
  t_Site<-unique(one_data$Site)
  
  
one_species<-one_data[,6:ncol(one_data)]
one_species[one_species>0]<-1
one_species<-as.data.frame(one_species)


one_coord <- one_data[,4:5]
one_coord <- one_coord %>% rename(x=DD_X,y=DD_Y)
one_coord <- as.data.frame(one_coord)

zeta <- Zeta.decline.mc(one_species,
                one_coord,
                orders = 1:nrow(one_species),
                sam = 100,
                NON = T)


data <- data.frame("z_order"=zeta[["zeta.order"]],
                       "zeta_value"=zeta[["zeta.val"]],
                       "zeta_sd"=zeta[["zeta.val.sd"]],
                 "zeta_ratio"=c(NA,zeta[["ratio"]]),
                 "year_month"=rep(t_year_month))


 aic_table<-as.data.frame(zeta[["aic"]])
 aic_table$year_month<- rep(t_year_month)
 aic_table$method<-rownames(aic_table)
 
 aic_table$method[aic_table$method=="zeta$zeta.exp"]<-"exp"

 aic_table$method[aic_table$method=="zeta$zeta.pl"]<-"plog" 
 
# calculte the delta and keep only the best model
 
 aic_table <- aic_table %>% mutate("delta_aic"= min(AIC)-max(AIC)) %>% filter(AIC == min(AIC))
   
 aic_list_all_15[[loop_all_15]]<-aic_table
 zeta_all_list_15[[i]]<-zeta
 zeta_data_all_list_15[[loop_all_15]]<-data


loop_all_15<-loop_all_15+1

  
  
}
  
zeta_divesity_all_15<-bind_rows(zeta_data_all_list_15)

aic_table_all_15<-bind_rows(aic_list_all_15)

ggplot(data =zeta_divesity_all_15,
       aes(x=z_order,y=zeta_value,color =year_month))+
  geom_line()+
  geom_point()+
  theme_bw()+xlab("Zeta order")+ylab("Zeta diversity")+
  geom_errorbar(aes(ymin = zeta_value-zeta_sd,ymax = zeta_value+zeta_sd),width = 0.2,alpha = 0.6)+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        plot.title = element_text(size=24))+
  scale_x_continuous(limits = c(0, 20), breaks = seq(0,20, by = 5))+
  scale_color_manual(values = fish_all,name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))

#ggsave("zeta_2015.pdf" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")

#ggsave("zeta_2015.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")

  
ggplot(data =zeta_divesity_all_15,
       aes(x=z_order,y=zeta_ratio,color =year_month))+
  geom_line()+
  geom_point()+
  theme_bw()+xlab("Zeta order")+ylab("Zeta ratio")+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        plot.title = element_text(size=24))+
     scale_color_manual(values = fish_all,name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("zeta_ratio_2015.pdf" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")

#ggsave("zeta_ratio_2015.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\zeta")
```

