---
title: "18.2 dissimilarity model"
author: "TG"
date: "5 4 2022"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(rareNMtests)
library(zetadiv)
library(car)

```

color scales definitions 

```{r}
color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")
```


# Data prepartion

```{r}
load("fish_full_data.rdata")
load("coral_comm.rdata")
select <- dplyr::select
```

## general steps:

```{r}
breakage <- read.csv("data\\breakage 2020 - 2018.csv")

fish_prep<-fish_full_data

```



1. final cleaning stages

(to be removed to the data prep script asap)

change mistakes in the confidence level of survey "Princess_6_2020-10-13" 

```{r}
fish_prep$Confidence<-ifelse(fish_prep$survey_id == "Princess_6_2020-10-13" & fish_prep$Species != "Gobiodon reticulatus",0,fish_prep$Confidence)
```

change the status of Pseudanthias squamipinnis to C here but i need to do it in the data prep script

```{r}

fish_prep$Status<-ifelse(fish_prep$Species %in% c("Pseudanthias squamipinnis",
                                                  "Neopomacentrus miryae",
                                                  "Chromis pembae",
                                                  "Chromis weberi"),"C",fish_prep$Status)

```

2. keep only 2018 and 2020 spring (storm)

```{r}

fish_prep<- fish_prep %>% filter(year_month %in% c("2018 b","2020 a"))

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```


7. keep only the knolls that were surveyd in both years

```{r}

knolls_2018<-fish_prep %>% filter(Year %in% c("2018")) %>% distinct(knoll_id)

knolls_2020<-fish_prep %>% filter(Year %in% c("2020")) %>% distinct(knoll_id)

united_knolls<-intersect(knolls_2018,knolls_2020)

fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

8. separate to all fish and resident fish only


```{r}

all_fish <- fish_prep 
resident_fish <- fish_prep %>% filter(Status != "T")

```


10. convert to wide format


all fish

```{r}

all_fish <- all_fish %>% 
  select(Site,year_month,Knoll,knoll_id,Surface_Area,coral_cover,Max_depth,Mean_Complexity,Hight,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,knoll_id,Surface_Area,coral_cover,Mean_Complexity,Max_depth,Hight,Species) %>% 
    summarise(Abundance=sum(Abundance))

all_fish <- spread(all_fish,
                      Species,
                      Abundance,
                      fill = 0)



```

resident fish

```{r}
resident_fish <- resident_fish %>% 
  select(Site,year_month,Knoll,knoll_id,Surface_Area,coral_cover,Mean_Complexity,Max_depth,Hight,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,knoll_id,Surface_Area,coral_cover,Mean_Complexity,Max_depth,Hight,Species) %>% 
    summarise(Abundance=sum(Abundance))

resident_fish <- spread(resident_fish,
                      Species,
                      Abundance,
                      fill = 0)



```


Im removing knoll_id 'Caves_5' for now because there is no species in 2020

```{r}
resident_fish<-resident_fish %>% filter(knoll_id != "Caves_5")
```



# calculate the disimilarity by Bray curtis index


Add the breakage data


use only the 2020 breakage data as variable but check the correlation between 2018 and 2020
```{r}


breakage<-breakage %>% select(1:5) %>% mutate("precentage"=Broken*100/Total)

breakage_wide <- breakage %>%
  mutate(knoll_id = paste(Site,Knoll,sep="_")) %>%
  select(knoll_id,Year,Site,precentage) %>% 
  filter(knoll_id != "NR_NA")

breakage_wide <- spread(breakage_wide,Year,precentage)
breakage_wide<- na.omit(breakage_wide)


ggplot(data = breakage_wide,aes(x=breakage_wide$`2018`,y=breakage_wide$`2020`,color = Site))+ geom_point()+geom_smooth(method = lm)


breakage_2020<-breakage %>% filter(Year =="2020") %>% na.omit() %>% mutate(knoll_id = paste(Site,Knoll,sep="_")) %>% select(knoll_id,precentage)
```

no correlation between 2018 and 2020 in broken coloneys so i can use only 2020 as my variable


## add coral richness data

```{r}
stone <- coral_comm %>% 
  filter(Category %in% c("Hydrozoa","Stony Coral"),
         year %in% c("2018","2020")) %>% 
  select(1:7) %>% 
  mutate(knoll_id = paste(site,knoll,sep="_")) %>% 
  filter(knoll_id %in% united_knolls$knoll_id) %>% 
  group_by(year,knoll_id) %>% summarise(c_richness =n_distinct(species)) %>% group_by(knoll_id) %>% 
  summarise(mean_c_rich = round(mean(c_richness),0))
```



#  only resident fish species



```{r}

bray_list_resid<-list()


for (i in unique(resident_fish$knoll_id)) {
  
  one_knoll <- resident_fish %>% filter(knoll_id == i) # compare one knoll
  
  species <-one_knoll[,10:ncol(one_knoll)] # species matrix
  
  dissimilarity <- vegdist(species,method = "bray") # calc dissimilarity
  
  # save to data frame
  
  bray <- data.frame("knoll_id" = i,
                     "dissimilarity" = as.numeric(dissimilarity)) 
  
  # save dataframe to list
  
  bray_list_resid[[i]] <- bray
  
}

bray_diss_resid<-bind_rows(bray_list_resid)

```

connect the dissimilarity to the data

```{r}
resident_fish <- resident_fish %>% 
  select(Site,year_month,Knoll,knoll_id,Surface_Area,coral_cover,Max_depth,Hight)

variables_resi<- resident_fish %>% 
  group_by(Site,knoll_id) %>%
  summarise("mean_Surface_Area"=mean(Surface_Area),
            "mean_coral_cover"=mean(coral_cover),
            "mean_Complexity"=mean(Mean_Complexity),
            "mean_Max_depth"=mean(Max_depth),
            "mean_Hight"=mean(Hight))

variables_resi$category_depth<-ifelse(variables_resi$mean_Max_depth>6,"deep","shallow")


bray_diss_resid<-left_join(variables_resi,bray_diss_resid)


```

# Add the breakage data



```{r}
bray_diss_resid<-left_join(bray_diss_resid,breakage_2020)
```


filter knolls without breakage data (or any other na)


```{r}

bray_diss_resid_small <- bray_diss_resid %>% na.omit()
  
```


some plots of the variables

```{r}
ggplot(data = bray_diss_resid,aes(x=log(mean_Surface_Area),y=dissimilarity))+geom_point()
ggplot(data = bray_diss_resid,aes(x=mean_Max_depth,y=dissimilarity))+geom_point()
ggplot(data = bray_diss_resid, aes(x=log(mean_Surface_Area),y=mean_Complexity))+geom_point()
ggplot(data = bray_diss_resid,aes(x=Site,y=dissimilarity,color=Site,label=knoll_id))+geom_point()+geom_text()

ggplot(data = bray_diss_resid,aes(x=Site,y=precentage,color = dissimilarity))+geom_point()

ggplot(data = bray_diss_resid,aes(precentage,dissimilarity,color = Site))+geom_point()

```


some plots of the variables without the NA rows

```{r}
ggplot(data = bray_diss_resid_small,aes(x=log(mean_Surface_Area),y=dissimilarity))+geom_point()
ggplot(data = bray_diss_resid_small,aes(x=mean_Max_depth,y=dissimilarity))+geom_point()
ggplot(data = bray_diss_resid_small, aes(x=log(mean_Surface_Area),y=mean_Complexity))+geom_point()
ggplot(data = bray_diss_resid_small,aes(x=Site,y=dissimilarity,color=Site,label=knoll_id))+geom_point()+geom_text()

ggplot(data = bray_diss_resid_small,aes(x=Site,y=precentage,color = dissimilarity))+geom_point()

ggplot(data = bray_diss_resid_small,aes(precentage,dissimilarity,color = Site))+geom_point()


ggplot(data = bray_diss_resid_small,aes(log(mean_Max_depth),dissimilarity,color = Site))+geom_point()
```

## models

###  global models

1. without interaction
2. with interaction

im using only the shallow knolls (<6m)
```{r}
bray_diss_resid_small$Site <-as.factor(bray_diss_resid_small$Site)
bray_diss_resid_small$category_depth <-as.factor(bray_diss_resid_small$category_depth)

#1

resident_diss<-lm(formula =  dissimilarity ~ Site + mean_Surface_Area + mean_coral_cover + mean_Complexity + precentage + log(mean_Max_depth) ,data = bray_diss_resid_small, na.action = "na.fail")


car::vif(resident_diss)

summary(resident_diss)
plot(resident_diss)



#2

resident_diss_interaction<-lm(formula =  dissimilarity ~ (Site + mean_Surface_Area +mean_coral_cover + mean_Complexity + precentage + log(mean_Max_depth))^2 ,data = bray_diss_resid_small, na.action = "na.fail")

summary(resident_diss_interaction)
plot(resident_diss_interaction)

car::vif(resident_diss_interaction)

```

### dreage  no interaction


```{r}

dredge_table_resident<-dredge(global.model = resident_diss)

#write.csv(dredge_table_resident,"dredge_table_resident_new.csv")


# first model

model_1<-lm(formula =  dissimilarity ~ Site + precentage,data = bray_diss_resid_small, na.action = "na.fail")

summary(model_1)
plot(model_1)

tukey_m1<-multcomp::glht(model_1,linfct = multcomp::mcp(Site="Tukey"))
summary(tukey_m1)

# second best model

model_2<-lm(formula =  dissimilarity ~ Site  +mean_coral_cover  + precentage ,data = bray_diss_resid_small, na.action = "na.fail")

summary(model_2)

# third best model

model_3<-lm(formula =  dissimilarity ~ Site  + precentage + category_depth ,data = bray_diss_resid_small, na.action = "na.fail")


summary(model_3)



```

### dredghe with interactions

```{r}
dredge_interaction_table_resident<-dredge(global.model = resident_diss_interaction)

int_aic<-dredge_interaction_table_resident %>% filter(delta<2.4508215
) 

int_aic <- int_aic[ , colSums(is.na(int_aic)) < nrow(int_aic)]  

#write.csv(int_aic,"dredge_intraction_table_resident_new.csv")

model_1_int<-lm(formula =  dissimilarity ~ Site  + log(mean_Max_depth),data = bray_diss_resid_small, na.action = "na.fail")

summary(model_1_int)
plot(model_1_int)

vif(model_1_int)

#tab_model(model_1_int,file = "diss_models//model_1_int.doc")

model_2_int<-lm(formula =  dissimilarity ~ Site  +log(mean_Max_depth) +precentage,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_2_int)

plot(model_2_int)

vif(model_2_int)

#tab_model(model_2_int,file = "diss_models//model_2_int.doc")


model_3_int<-lm(formula =  dissimilarity ~ Site  + log(mean_Max_depth) +mean_coral_cover,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_3_int)

plot(model_3_int)

vif(model_3_int)


#tab_model(model_3_int,file = "diss_models//model_3_int.doc")

model_4_int<-lm(formula =  dissimilarity ~ Site  +log(mean_Max_depth) +mean_coral_cover+precentage,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_4_int)

plot(model_4_int)

vif(model_4_int)

#tab_model(model_4_int,file = "diss_models//model_4_int.doc")


model_5_int<-lm(formula =  dissimilarity ~ Site  + mean_coral_cover*precentage,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_5_int)

plot(model_5_int)
vif(model_5_int)


#tab_model(model_5_int,file = "diss_models//model_5_int.doc")

model_6_int<-lm(formula =  dissimilarity ~ Site +log
                (mean_Max_depth) + mean_coral_cover*precentage,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_6_int)

vif(model_6_int)

plot(model_6_int)

#tab_model(model_6_int,file = "diss_models//model_5_int.doc")

model_7_int<-lm(formula =  dissimilarity ~ Site + precentage,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_7_int)

vif(model_7_int)

model_8_int<-lm(formula =  dissimilarity ~ Site + precentage + mean_coral_cover,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_8_int)

vif(model_8_int)


model_9_int<-lm(formula =  dissimilarity ~ Site + log(mean_Max_depth) * mean_Surface_Area,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_9_int)

vif(model_9_int)

model_10_int<-lm(formula =  dissimilarity ~ Site + log(mean_Max_depth) + mean_Surface_Area,data = bray_diss_resid_small, na.action = "na.fail")
summary(model_10_int)

vif(model_10_int)


```





# all fish species

- no data standardization



```{r}

bray_list<-list()


for (i in unique(all_fish$knoll_id)) {
  
  one_knoll <- all_fish %>% filter(knoll_id == i) # compare one knoll
  
  species <-one_knoll[,10:ncol(one_knoll)] # species matrix
  
  dissimilarity <- vegdist(species,method = "bray") # calc dissimilarity
  
  # save to data frame
  
  bray <- data.frame("knoll_id" = i,
                     "dissimilarity" = as.numeric(dissimilarity)) 
  
  # save dataframe to list
  
  bray_list[[i]] <- bray
  
}

bray_diss<-bind_rows(bray_list)

```

connect the dissimilarity to the data

```{r}
all_fish <- all_fish %>% 
  select(Site,year_month,Knoll,knoll_id,Surface_Area,coral_cover,Max_depth,Hight)

variables<- all_fish %>% 
  group_by(Site,knoll_id) %>%
  summarise("mean_Surface_Area"=mean(Surface_Area),
            "mean_coral_cover"=mean(coral_cover),
            "mean_Complexity"=mean(Mean_Complexity),
            "mean_Max_depth"=mean(Max_depth),
            "mean_Hight"=mean(Hight))

variables$category_depth<-ifelse(variables$mean_Max_depth>6,"deep","shallow")


bray_diss<-left_join(variables,bray_diss)


```



```{r}
bray_diss<-left_join(bray_diss,breakage_2020)
```




some plots of the variables

```{r}
ggplot(data = bray_diss,aes(x=log(mean_Surface_Area),y=dissimilarity,color = Site))+geom_point()
ggplot(data = bray_diss,aes(mean_Hight,dissimilarity,color = Site))+geom_point()
ggplot(data = bray_diss,aes(mean_Hight,log(mean_Surface_Area),color = Site))+geom_point()


ggplot(data = bray_diss,aes(x=mean_Max_depth,y=dissimilarity))+geom_point()
ggplot(data = bray_diss, aes(x=log(mean_Surface_Area),y=mean_Complexity))+geom_point()
ggplot(data = bray_diss,aes(x=Site,y=dissimilarity,color=Site,label=knoll_id))+geom_point()+geom_text()

ggplot(data = bray_diss,aes(x=Site,y=dissimilarity,color=precentage))+geom_point()

ggplot(data = bray_diss,aes(precentage,dissimilarity,color = Site))+geom_point()


```


```{r}

bray_diss_small <- bray_diss %>% filter(category_depth=="shallow") %>% na.omit()
  
```


```{r}
diss_interaction<-lm(formula =  dissimilarity ~ (Site + mean_Surface_Area +mean_coral_cover + mean_Complexity + precentage+ mean_Max_depth)^2 ,data = bray_diss_small, na.action = "na.fail")

diss<-lm(formula =  dissimilarity ~ Site + mean_Surface_Area +mean_coral_cover + mean_Complexity + precentage+mean_Max_depth ,data = bray_diss_small, na.action = "na.fail")

diss_no_site<-lm(formula =  dissimilarity ~  mean_Surface_Area +mean_coral_cover + mean_Complexity + precentage+mean_Max_depth ,data = bray_diss_small, na.action = "na.fail")


dredge_table<-dredge(global.model = diss)
dredge_table_no_site<-dredge(global.model = diss_no_site)
dredge_interaction_table<-dredge(global.model = diss_interaction)





#summary(diss)
#plot(diss)

#diss_dredge <- lm(formula =  dissimilarity ~ Site + mean_Surface_Area ,data = bray_diss_small, na.action = "na.fail")


#summary(diss_dredge)



```
