---
title: "19.1 commoness index"
author: "TG"
date: "17 4 2022"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(Rarefy)
library(rareNMtests)
library(zetadiv)
library(FuzzyQ)
library(SingleCaseES)

```

test fo single case ES

```{r}
data(Schmidt2007)
batch_calc_ES(dat = Schmidt2007,
              grouping = c(Behavior_type, Case_pseudonym, Phase_num),
              condition = Condition,
              outcome = Outcome,
              ES = c("LRRi","LRRd"),
              improvement = direction,
              scale = Metric,
              bias_correct = TRUE,
              confidence = NULL,
              format = "wide")


data(McKissick)
batch_calc_ES(McKissick,
              grouping = Case_pseudonym,
              condition = Condition,
              outcome = Outcome,
              ES = c("LRRd","LRRi"),
              improvement = "decrease",
              scale = "count",
              observation_length = 20,
              format = "long")


batch_calc_ES(dat = Schmidt2007,
              grouping = c(Behavior_type, Case_pseudonym),
              aggregate = Phase_num,
              weighting = "1/V",
              condition = Condition,
              outcome = Outcome,
              ES = c("LRRi", "LRRd", "SMD", "Tau"),
              improvement = direction,
              scale = "count",
              bias_correct = TRUE,
              confidence = NULL,
              format = "long")
```




color scales definitions 

```{r}
color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")
```


# Data prepartion

```{r}
load("fish_full_data.rdata")
select <- dplyr::select



```

change the values of: 
Tylosurus choram (from 100 ind * 2cm to 2ind*100cm )
Pomacentrus trichrourus (50 to 5 cm)


```{r}

fish_full_data$Length<-ifelse(fish_full_data$Species=="Tylosurus choram" &
fish_full_data$Abundance == 100 &
fish_full_data$survey_id =="Caves_14_2020-06-14",
100,
fish_full_data$Length)

fish_full_data$Abundance<-ifelse(fish_full_data$Species=="Tylosurus choram" &
fish_full_data$Abundance == 100 &
fish_full_data$survey_id =="Caves_14_2020-06-14",
2,
fish_full_data$Abundance)


fish_full_data$Length<-ifelse(fish_full_data$Species=="Pomacentrus trichrourus" &
fish_full_data$Length == 50 &
fish_full_data$survey_id =="Princess_5_2020-06-14",
5,
fish_full_data$Length)

fish_prep<-fish_full_data
```



1. final cleaning stages

(to be removed to the data prep script asap)

change mistakes in the confidence level of survey "Princess_6_2020-10-13" 

```{r}
fish_prep$Confidence<-ifelse(fish_prep$survey_id == "Princess_6_2020-10-13" & fish_prep$Species != "Gobiodon reticulatus",0,fish_prep$Confidence)
```

change the status of Pseudanthias squamipinnis to C here but i need to do it in the data prep script

```{r}

fish_prep$Status<-ifelse(fish_prep$Species %in% c("Pseudanthias squamipinnis",
                                                  "Neopomacentrus miryae",
                                                  "Chromis pembae",
                                                  "Chromis weberi"),"C",fish_prep$Status)

```




2. keep only data from 2018 
```{r}

fish_prep<- fish_prep %>% filter(Year > 2018)

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```



7. keep only the knolls that were surveyed in both years

(otherwise of coarse 2018 will have higher abundances...more knolls)

```{r}

knolls_2018<-fish_prep %>% filter(Year %in% c("2018")) %>% distinct(knoll_id)

knolls_2020<-fish_prep %>% filter(year_month %in% c("2020 a")) %>% distinct(knoll_id)

knolls_2020_b<-fish_prep %>% filter(year_month %in% c("2020 b")) %>% distinct(knoll_id)


united_knolls<-intersect(knolls_2018,knolls_2020,knolls_2020_b)

fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

7. separate to all fish and resident fish only


```{r}

all_fish <- fish_prep 
resident_fish <- fish_prep %>% filter(Status != "T")

```

8. prepre rank data

based on fuzzyQ of 2018 (all fish)


```{r}

fuzzy_matrix <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
    summarise(Abundance=sum(Abundance))

fuzzy_matrix <- spread(fuzzy_matrix,
                      Species,
                      Abundance,
                      fill = 0)

fuzzy_matrix<-fuzzy_matrix[,4:ncol(fuzzy_matrix)]

fuzzy <- fuzzyq(fuzzy_matrix)

rank_data<-fuzzy[["spp"]]

rank_data$Species<-rownames(rank_data)

rownames(rank_data)<-NULL

rank_data<-rank_data %>%
  arrange(desc(Common.I)) %>%
  mutate("rank"=1:nrow(rank_data)) %>% 
  select(Species,rank,Common.I)


```
9.

```{r}

```


# log ratios 


9. calculate log ratio for 2018 spring 2020


```{r}

log_data <- all_fish %>%
  filter(year_month %in% c("2018 b","2020 a")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data <- spread(log_data,year_month,Abundance)

log_data <- left_join(rank_data,log_data)

log_data <- log_data %>%
  rename("before" = `2018 b`,"after"= `2020 a`) %>%
  mutate(ratio = after/before, log_ratio = log(ratio))

log_data<-na.omit(log_data)

```


plot

```{r}
log_data$Species <- reorder(log_data$Species, -log_data$rank)

log_data

ggplot(data = log_data,aes(x=log_ratio,y=Species,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("spring 2020/ Fall 2018")

#ggsave("log_ratio_storm_2018.pdf" , width = 15, height = 18, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

```

some conclusions:


- most species show decline in abundance (80 out of 120)

- most rare species has 0 abundance in one of the samples and thus were omitted

there might be an seasonal effect...

- Coris aygula (rank 35) - seasonal effect mostly, most ind in 2020 are juvenile

- Coris caudimaculate (rank 73) - seasonal effect mostly, many ind in 2020 are juvenile

- this is true also the other way around - most ind of  Chlorurus sordidus at 2018 are juveniles



# make the same comperison but for fall 2018 and fall 2020

```{r}
rm(list=setdiff(ls(), c("united_knolls", "fish_full_data")))
fish_prep<-fish_full_data
```



1. final cleaning stages

(to be removed to the data prep script asap)

change mistakes in the confidence level of survey "Princess_6_2020-10-13" 

```{r}
fish_prep$Confidence<-ifelse(fish_prep$survey_id == "Princess_6_2020-10-13" & fish_prep$Species != "Gobiodon reticulatus",0,fish_prep$Confidence)
```

change the status of Pseudanthias squamipinnis to C here but i need to do it in the data prep script


2. keep only 2018 and 2020 fall 

```{r}

fish_prep<- fish_prep %>% filter(year_month %in% c("2018 b","2020 b"))

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number




```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```




7. keep only the knolls that were surveyed in all years



```{r}

fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

7. separate to all fish and resident fish only


```{r}

all_fish <- fish_prep 
resident_fish <- fish_prep %>% filter(Status != "T")

```

8. prepre rank data

based on fuzzyQ of 2018 (all fish)


```{r}

fuzzy_matrix <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
    summarise(Abundance=sum(Abundance))

fuzzy_matrix <- spread(fuzzy_matrix,
                      Species,
                      Abundance,
                      fill = 0)

fuzzy_matrix<-fuzzy_matrix[,4:ncol(fuzzy_matrix)]

fuzzy <- fuzzyq(fuzzy_matrix)

rank_data<-fuzzy[["spp"]]

rank_data$Species<-rownames(rank_data)

rownames(rank_data)<-NULL

rank_data<-rank_data %>%
  arrange(desc(Common.I)) %>%
  mutate("rank"=1:nrow(rank_data)) %>% 
  select(Species,rank,Common.I)


```

9. calculate log ratio


```{r}

log_data <- all_fish %>% group_by(year_month,Species) %>% summarise(Abundance = sum(Abundance))

log_data <- spread(log_data,year_month,Abundance)

log_data <- left_join(rank_data,log_data)

log_data <- log_data %>%
  rename("before" = `2018 b`,"after"= `2020 b`) %>%
  mutate(ratio = after/before, log_ratio = log(ratio))

log_data<-na.omit(log_data)

```


plot

```{r}
log_data$Species <- reorder(log_data$Species, -log_data$rank)

ggplot(data = log_data,aes(x=log_ratio,y=Species))+
  geom_bar(stat="identity",color = "black",fill = "#eddb6c")+
  theme_classic()+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Fall 2020/ Fall 2018")

#ggsave("log_ratio_F18_F20.pdf" , width = 15, height = 18, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

```


# make the same comperison but for spring 2020 and spring 2021 

```{r}
rm(list=setdiff(ls(), c("united_knolls", "fish_full_data")))
fish_prep<-fish_full_data
```



1. final cleaning stages

(to be removed to the data prep script asap)

change mistakes in the confidence level of survey "Princess_6_2020-10-13" 

```{r}
fish_prep$Confidence<-ifelse(fish_prep$survey_id == "Princess_6_2020-10-13" & fish_prep$Species != "Gobiodon reticulatus",0,fish_prep$Confidence)
```

change the status of Pseudanthias squamipinnis to C here but i need to do it in the data prep script


2. keep only 2020 spring (storm) and 2021 spring 

```{r}

fish_prep<- fish_prep %>% filter(year_month %in% c("2020 a","2021 a"))

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number




```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```




7. keep only the knolls that were surveyed in all years



```{r}

fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

7. separate to all fish and resident fish only


```{r}

all_fish <- fish_prep 
resident_fish <- fish_prep %>% filter(Status != "T")

```

8. prepre rank data

based on fuzzyQ of 2018 (all fish)


```{r}

fuzzy_matrix <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
    summarise(Abundance=sum(Abundance))

fuzzy_matrix <- spread(fuzzy_matrix,
                      Species,
                      Abundance,
                      fill = 0)

fuzzy_matrix<-fuzzy_matrix[,4:ncol(fuzzy_matrix)]

fuzzy <- fuzzyq(fuzzy_matrix)

rank_data<-fuzzy[["spp"]]

rank_data$Species<-rownames(rank_data)

rownames(rank_data)<-NULL

rank_data<-rank_data %>%
  arrange(desc(Common.I)) %>%
  mutate("rank"=1:nrow(rank_data)) %>% 
  select(Species,rank,Common.I)


```

9. calculate log ratio


```{r}

log_data <- all_fish %>% group_by(year_month,Species) %>% summarise(Abundance = sum(Abundance))

log_data <- spread(log_data,year_month,Abundance)

log_data <- left_join(rank_data,log_data)

log_data <- log_data %>%
  rename("storm" = `2020 a`,"after"= `2021 a`) %>%
  mutate(ratio = after/storm, log_ratio = log(ratio))

log_data<-na.omit(log_data)

```


plot

```{r}
log_data$Species <- reorder(log_data$Species, -log_data$rank)

ggplot(data = log_data,aes(x=log_ratio,y=Species))+
  geom_bar(stat="identity",color = "black",fill = "#de1e2b")+
  theme_classic()+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("spring 2021/ spring 2020")

#ggsave("log_ratio_S20_S21.pdf" , width = 15, height = 18, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

```