---
title: "19.2 commoness index"
author: "TG"
date: "2 5 2022"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(Rarefy)
library(rareNMtests)
library(zetadiv)
library(FuzzyQ)
library(SingleCaseES)
library(extrafont)

#font_import()

```

test fo single case ES

```{r eval=FALSE, warning=TRUE, include=FALSE}
data(Schmidt2007)
batch_calc_ES(dat = Schmidt2007,
              grouping = c(Behavior_type, Case_pseudonym, Phase_num),
              condition = Condition,
              outcome = Outcome,
              ES = c("LRRi","LRRd"),
              improvement = direction,
              scale = Metric,
              bias_correct = TRUE,
              confidence = NULL,
              format = "wide")


data(McKissick)
batch_calc_ES(McKissick,
              grouping = Case_pseudonym,
              condition = Condition,
              outcome = Outcome,
              ES = c("LRRd","LRRi"),
              improvement = "decrease",
              scale = "count",
              observation_length = 20,
              format = "long")


batch_calc_ES(dat = Schmidt2007,
              grouping = c(Behavior_type, Case_pseudonym),
              aggregate = Phase_num,
              weighting = "1/V",
              condition = Condition,
              outcome = Outcome,
              ES = c("LRRi", "LRRd", "SMD", "Tau"),
              improvement = direction,
              scale = "count",
              bias_correct = TRUE,
              confidence = NULL,
              format = "long")

rm(Schmidt2007)
rm(McKissick)

```




color scales definitions 

```{r}
color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")
rank_filter <-51
```


# Data prepartion

```{r}
load("fish_full_data.rdata")
select <- dplyr::select

fish_prep<-fish_full_data
```


2. keep only data from 2018 
```{r}

fish_prep<- fish_prep %>% filter(Year > 2017)

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
  filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
  
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```



7. keep only the knolls that were surveyed all  years

(otherwise of coarse 2018 will have higher abundances...more knolls)

```{r}

knolls_2018<-fish_prep %>% filter(Year %in% c("2018")) %>% distinct(knoll_id)

knolls_2020_a<-fish_prep %>% filter(year_month %in% c("2020 a")) %>% distinct(knoll_id)

knolls_2020_b<-fish_prep %>% filter(year_month %in% c("2020 b")) %>% distinct(knoll_id)

knolls_2021_a<-fish_prep %>% filter(year_month %in% c("2021 a")) %>% distinct(knoll_id)

united_knolls<-intersect(knolls_2018,knolls_2020_a,knolls_2020_b,knolls_2021_a)

fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

7. separate to all fish and resident fish only


```{r}

all_fish <- fish_prep 
resident_fish <- fish_prep %>% filter(Status != "T")

```

## prepre rank data

based on fuzzyQ of all fish from 2018 and on


```{r}

fuzzy_matrix <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
  summarise(Abundance=sum(Abundance))

fuzzy_matrix <- spread(fuzzy_matrix,
                       Species,
                       Abundance,
                       fill = 0)

fuzzy_matrix_species<-fuzzy_matrix[,4:ncol(fuzzy_matrix)]

fuzzy <- fuzzyq(fuzzy_matrix_species)

rank_data<-fuzzy[["spp"]]

rank_data$Species<-rownames(rank_data)

rownames(rank_data)<-NULL

rank_data<-rank_data %>%
  arrange(desc(Common.I)) %>%
  mutate("rank"=1:nrow(rank_data)) %>% 
  select(Species,rank,Common.I)


family_data<-all_fish %>% select(Family,Species) %>% distinct(Species,.keep_all = T)

rank_data<-left_join(rank_data,family_data)
```

## prepare log ratios

recreate long data frame with all the species in all the knolls including 0 for species we didnt see

```{r}

ci_data<-fuzzy_matrix

ci_data_all<-gather(ci_data,key = "Species",value = "Abundance",4:ncol(ci_data))




```


# log ratios for 2018 spring 2020

#### prepare the the ci

- for the function to work I need to filter only the data from the relevant years for the log ratio.

- the function also calculate the log ratio but i also keeping my previous log ratio calculation for control


```{r}
ci_data<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2018 b","2020 a"))

calc_species_se = batch_calc_ES(dat = ci_data, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")
```


```{r}

log_data <- all_fish %>%
  filter(year_month %in% c("2018 b","2020 a")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data <- spread(log_data,year_month,Abundance)

log_data <- left_join(rank_data,log_data)

log_data <- left_join(log_data,calc_species_se)


log_data <- log_data %>%
  rename("before" = `2018 b`,"after"= `2020 a`) %>%
  mutate(ratio = after/before, log_ratio = log(ratio))

#log_data<-na.omit(log_data)


log_data<-log_data %>% mutate(test = lag(rank)-rank)




```
# plots

```{r}
log_data$Species <- reorder(log_data$Species, -log_data$rank)

log_data$Species_rank<-paste0(log_data$Species," (",log_data$rank,")")
log_data$Species_rank <- reorder(log_data$Species_rank, -log_data$rank)


log_data_short <- log_data %>% filter(rank < rank_filter)


ggplot(data = log_data,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  geom_errorbar(aes( xmin= CI_lower, xmax= CI_upper),width = 0.5)+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.text = element_text(size =20,face = "italic",family = "mono"))+
  ggtitle("spring 2020/ Fall 2018")

ggsave("log_ratio_storm_2018_v1.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

ggplot(data = log_data_short,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.text = element_text(size =20,face = "italic",family = "mono"))+
  ggtitle("spring 2020/ Fall 2018")

ggsave("log_ratio_storm_2018_sh1.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

log_data$Species <- reorder(log_data$Species, -log_data$Est)
log_data$Species_rank <- reorder(log_data$Species_rank, -log_data$Est)

log_data_short <- log_data %>% filter(rank < rank_filter)


ggplot(data = log_data,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme(text = element_text(family = "serif"),
    axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.text = element_text(size =20,face = "italic"))+
  ggtitle("spring 2020/ Fall 2018")

ggsave("log_ratio_storm_2018_v2.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

ggplot(data = log_data_short,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+xlim(-5,5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.text = element_text(size =22,face = "italic"))+
  ggtitle("spring 2020/ Fall 2018")

ggsave("log_ratio_storm_2018_sh2.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")
                
cor.test(log_data$Common.I,log_data$Est)

```
                
                
                
                
### log ratios + ci for 2018 fall to 2020 fall
                
                
#### prepare the the ci
                
- for the function to work I need to filter only the data from the relevant years for the log ratio.
                
- the function also calculate the log ratio but i also keeping my previous log ratio calculation for control
                
                
```{r}
ci_data_2<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2018 b","2020 b"))

calc_species_se_2 = batch_calc_ES(dat = ci_data_2, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")

```
                
                
    
    
                
```{r}

log_data_2 <- all_fish %>%
  filter(year_month %in% c("2018 b","2020 b")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data_2 <- spread(log_data_2,year_month,Abundance)

log_data_2 <- left_join(rank_data,log_data_2)
log_data_2 <- left_join(log_data_2,calc_species_se_2)


log_data_2 <- log_data_2 %>%
  rename("before" = `2018 b`,"after"= `2020 b`) %>%
  mutate(ratio = after/before, log_ratio = log(ratio))

#log_data_2<-na.omit(log_data_2)


log_data_2<-log_data_2 %>% mutate(test = lag(rank)-rank)
```

# plot

```{r}
log_data_2$Species <- reorder(log_data_2$Species, -log_data_2$rank)


log_data_2$Species_rank<-paste0(log_data_2$Species," (",log_data_2$rank,")")
log_data_2$Species_rank <- reorder(log_data_2$Species_rank, -log_data_2$rank)


log_data_short_2 <- log_data_2 %>% filter(rank < rank_filter)

ggplot(data = log_data_2,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  scale_y_discrete(position = "right")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.text = element_text(size =20,face = "italic",family = "mono"))+
  ggtitle("Fall 2020/ Fall 2018")

ggsave("log_ratio_F18_F20_v1.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

ggplot(data = log_data_short_2,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower, xmax= CI_upper),width = 0.5)+
  theme_classic()+
  scale_y_discrete(position = "right")+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Fall 2020/ Fall 2018")

ggsave("log_ratio_F18_F20_sh1.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

log_data_2$Species <- reorder(log_data_2$Species, -log_data_2$Est)
log_data_2$Species_rank <- reorder(log_data_2$Species_rank, -log_data_2$Est)

log_data_short_2 <- log_data_2 %>% filter(rank < rank_filter)


ggplot(data = log_data_2,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  scale_y_discrete(position = "right")+xlim(-5,5)+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Fall 2020/ Fall 2018")

ggsave("log_ratio_F18_F20_v2.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")


ggplot(data = log_data_short_2,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+
  scale_y_discrete(position = "right")+xlim(-5,5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Fall 2020/ Fall 2018")


ggsave("log_ratio_F18_F20_sh2.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

```
                                                
                                                
                                                
### log ratios for 2020 spring to 2021 spring
                                                
- for the function to work I need to filter only the data from the relevant years for the log ratio.
                                                
- the function also calculate the log ratio but i also keeping my previous log ratio calculation for control
                                                
                                                
```{r}
ci_data_3<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2020 a","2021 a"))

calc_species_se_3 = batch_calc_ES(dat = ci_data_3, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")
```
                                                
                                                
                                                
```{r}
log_data_3 <- all_fish %>%
  filter(year_month %in% c("2020 a","2021 a")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data_3 <- spread(log_data_3,year_month,Abundance)

log_data_3 <- left_join(rank_data,log_data_3)
log_data_3 <- left_join(log_data_3,calc_species_se_3)
log_data_3 <- log_data_3 %>%
  rename("storm" = `2020 a`,"after"= `2021 a`) %>%
  mutate(ratio = after/storm, log_ratio = log(ratio))


log_data_3<-log_data_3 %>% mutate(test = lag(rank)-rank)

```
                                                
                                                
# plot
                                                
```{r}
log_data_3$Species <- reorder(log_data_3$Species, -log_data_3$rank)

log_data_3$Species_rank<-paste0(log_data_3$Species," (",log_data_3$rank,")")
log_data_3$Species_rank <- reorder(log_data_3$Species_rank, -log_data_3$rank)

log_data_short_3 <- log_data_3 %>% filter(rank < rank_filter)

ggplot(data = log_data_3,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Spring 2020/ Spring 2021")

ggsave("log_ratio_S20_S21_v1.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

ggplot(data = log_data_short_3,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Spring 2020/ Spring 2021")


ggsave("log_ratio_S20_S21_sh1.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

log_data_3$Species <- reorder(log_data_3$Species, -log_data_3$Est)
log_data_3$Species_rank <- reorder(log_data_3$Species_rank, -log_data_3$Est)

log_data_short_3 <- log_data_3 %>% filter(rank < rank_filter)


ggplot(data = log_data_3,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  theme_classic()+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme(axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =20,face = "italic",family = "mono"),
                                   axis.title.x = element_text(size=22),
                                   axis.title.y = element_text(size=22))+
          ggtitle("Spring 2020/ Spring 2021")

ggsave("log_ratio_S20_S21_v2.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")

ggplot(data = log_data_short_3,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+xlim(-5,5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Spring 2020/ Spring 2021")


ggsave("log_ratio_S20_S21_sh2.pdf" , width = 15, height = 22, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\log_ratio")


```
                                                                                
                                                                                