---
title: "2.1 species area relationship"
author: "TG"
date: "22/11/2020"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(sars)
```

# species - area relationship

1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0
```{r}

sp_area<-one_observer

sp_area$Confidence[is.na(sp_area$Confidence)]<-(0)

sp_area<-sp_area %>% filter(Confidence <2)
```




the mean knoll size in each site

```{r}
knolls_size<-sp_area %>% distinct(knoll_id,.keep_all = T) %>% select(knoll_id,Site,Knoll,projected_area_ellipse,Surface.Area)


porj_area_violin<-ggplot(knolls_size, aes(x=Site,y=projected_area_ellipse ,fill=Site)) +
  geom_violin()+ 
  stat_summary(fun=mean, geom="point")+ 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1,shape=16)+
  geom_hline(yintercept=mean(knolls_size$projected_area_ellipse,na.rm = T),linetype="dashed",color="red")+
  theme_bw()

porj_area_violin

+geom_boxplot(width=0.1) geom_jitter(shape=16, position=position_jitter(0.2))+
```


## by site:

```{r}

sp_area<-sp_area %>% filter(Status =="C") %>% filter(Site != "North beach")

sp_for_survey<-sp_area %>% group_by(survey_id,Site,Year,Knoll,Surface.Area,projected_area_ellipse,Species) %>% summarise(species_number = sum(Abundance))


sp_for_survey_sum<-sp_for_survey %>% group_by(survey_id,Site,Year,Knoll,Surface.Area,projected_area_ellipse) %>% summarise(species_number = n())

sp_for_survey_sum<-na.omit(sp_for_survey_sum)
```

1. ill keep all the years as repetitions...

*this is for projected area

```{r}
fit_list<-list()
loop<-1

for( i in unique(sp_for_survey_sum$Site)){
 
  print(i)
  
  site_data <- sp_for_survey_sum %>% 
    filter(Site == i) 

  fit_list[[loop]]<- sar_loga(site_data[,6:7])
  
  loop<-loop+1
}

plot(fit_list[[1]],ModTitle  = "Caves - log model")
summary(fit_list[[1]])

plot(fit_list[[2]],ModTitle  = "Katza - log model")
summary(fit_list[[2]])

plot(fit_list[[3]],ModTitle  = "NR - log model")
summary(fit_list[[3]])

plot(fit_list[[4]],ModTitle  = "Princess - log model")
summary(fit_list[[4]])

```


2. ill use the mean number of species for each knoll from all the years


```{r}

sp_for_survey_mean<-sp_for_survey_sum %>% 
  group_by(Site,Knoll,Surface.Area,projected_area_ellipse) %>%
  summarise(species_number = mean(species_number))

fit_list_mean<-list()
loop<-1

for( i in unique(sp_for_survey_mean$Site)){
 
  print(i)
  
  site_data_mean <- sp_for_survey_mean %>% 
    filter(Site == i) 

  fit_list_mean[[loop]]<- sar_loga(site_data_mean[,4:5])
  
  loop<-loop+1
}

plot(fit_list_mean[[1]],ModTitle  = "Caves - log model")
summary(fit_list_mean[[1]])

plot(fit_list_mean[[2]],ModTitle  = "Katza - log model")
summary(fit_list_mean[[2]])

plot(fit_list_mean[[3]],ModTitle  = "NR - log model")
summary(fit_list_mean[[3]])

plot(fit_list_mean[[4]],ModTitle  = "Princess - log model")
summary(fit_list_mean[[4]])

```

















### with 'sp.' species & transient species

```{r}

by_site <- sp_area %>% group_by(survey_id,Site,Knoll,projected_area_ellipse,Surface.Area) %>% distinct(Species,.keep_all = T)
  
  
by_site<-by_site %>% group_by(Site,Knoll,projected_area_ellipse,Surface.Area) %>%
   summarise(species_number = n()) %>% arrange(Surface.Area)
  
by_site<-na.omit(by_site)

by_site$Knoll<-as.numeric(by_site$Knoll)

ggplot(by_site,aes(log(Surface.Area),log(species_number),color=Site))+geom_smooth(method = lm,se=F)+geom_point()

ggplot(by_site,aes(log(projected_area_ellipse),log(species_number),color=Site))+geom_smooth(method = lm,se=F)+geom_point()


```
 
test for sars
```{r}
for_sars<- sp_area %>%filter(Site=="Princess") %>%  group_by(Site,Knoll,Surface.Area) %>%
   summarise(species_number = n()) %>% arrange(Surface.Area) 

fit_loga <- sar_loga(data = for_sars[,3:4])
fit_helg<-sar_heleg(data = for_sars[,3:4])	
plot(fit_loga)
summary(fit_loga)
plot(fit_loga_1)
```

plot SAR for each site 

* not sure if i need to use mean number of species per knoll (ffrom all the years) or keep them as repeatitions 

```{r}

```


### with 'sp.' species & NO transient species





### without 'sp.' species

```{r}

```


## for each site by year



```{r}

```

