---
title: "2.2 species area relationship"
author: "TG"
date: "26 1 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# species - area relationship

```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(sars)
library(lme4)
library(lmerTest)
library(MuMIn)
```

## Data preperation

1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

sp_area<-one_observer
sp_area$Confidence[is.na(sp_area$Confidence)]<-(0)
sp_area<-sp_area %>% filter(Confidence <2)

```

3) filter Transient species

```{r}
sp_area<-sp_area %>% filter(Status == "C")

```


4) remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

sp_area<-sp_area %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")

```


5) plot the knolls size in each site

```{r}
knolls_size<-sp_area %>% distinct(knoll_id,.keep_all = T) %>% select(knoll_id,Site,Knoll,projected_area_ellipse,Surface.Area)


porj_area_violin<-ggplot(knolls_size, aes(x=Site,y=projected_area_ellipse ,fill=Site)) +
  geom_violin()+ 
  stat_summary(fun=mean, geom="point")+ 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1,shape=16)+
  geom_hline(yintercept=mean(knolls_size$projected_area_ellipse,na.rm = T),linetype="dashed",color="red")+
  theme_bw()

porj_area_violin


surf_area_violin<-ggplot(knolls_size, aes(x=Site,y= Surface.Area ,fill=Site)) +
  geom_violin()+ 
  stat_summary(fun=mean, geom="point")+ 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=1,shape=16)+
  geom_hline(yintercept=mean(knolls_size$Surface.Area,na.rm = T),linetype="dashed",color="red")+
  theme_bw()

surf_area_violin

```


6) set the number of species per knoll

* note - Im using here the "down-graded" data - fish are grouped according to 2016 sp list


```{r}

species_knoll<-sp_area %>% group_by(survey_id) %>% mutate(Species_number = n_distinct(Species_grouped)) %>% select(-Species,-Species_grouped,-Family,-Abundance,-Length,-Confidence,-Status) %>%  ungroup() %>% distinct(.keep_all = T)

```



## MODELS

### with lme4 package

1) Use the Surface Area parameter

* year - random effect
* Max depth - random effect
* log(surface area) in interaction with site

```{r}

sp_surface<-lmer(formula = log(Species_number) ~ log(Surface.Area):Site + Max.depth +(1|Year),data = species_knoll)


summary(sp_surface)
r.squaredGLMM(sp_surface)

plot(sp_surface)

qqnorm(resid(sp_surface)) 
qqline(resid(sp_surface), col = "red")

plot(fitted(sp_surface), resid(sp_surface, type = "pearson"))# this will create the plot
abline(0,0, col="red")

ggplot(data = species_knoll, 
       aes(x   = log(Surface.Area),
           y   = log(Species_number), 
           col = as.factor(Site)))+
  geom_point(size     = 1, 
             alpha    = .7, 
             position = "jitter")+
  geom_smooth(method   = lm,
              se       = T, 
              size     = 1.5, 
              linetype = 1, 
              alpha    = .2)+
  theme_minimal()
```


2) Use the projected area parameter

* year - random effect
* Max depth - random effect
* log(surface area) in interaction with site

```{r}

sp_projected<-lmer(formula = log(Species_number) ~ log(projected_area_ellipse):Site + Max.depth +(1|Year),data = species_knoll)


summary(sp_projected)
r.squaredGLMM(sp_projected)

plot(sp_projected)

qqnorm(resid(sp_projected)) 
qqline(resid(sp_projected), col = "red")

plot(fitted(sp_projected), resid(sp_projected, type = "pearson"))# this will create the plot
abline(0,0, col="red")

ggplot(data = species_knoll, 
       aes(x   = log(projected_area_ellipse),
           y   = log(Species_number), 
           col = as.factor(Site)))+
  geom_point(size     = 1, 
             alpha    = .7, 
             position = "jitter")+
  geom_smooth(method   = lm,
              se       = T, 
              size     = 1.5, 
              linetype = 1, 
              alpha    = .2)+
  theme_minimal()
```

### Sars Models

#### Log models

c+z×log(A)

```{r}
fit_list<-list()
loop<-1

for( i in unique(species_knoll$Site)){
 
  print(i)
  
  site_data <- species_knoll %>% 
    filter(Site == i) %>% select(Surface.Area,Species_number) 
  
  site_data<-na.omit(site_data)

  fit_list[[i]]<- sar_loga(site_data)
  
  
}

plot(fit_list[[1]],ModTitle  = "NR - log model")
summary(fit_list[[1]])

plot(fit_list[[2]],ModTitle  = "Caves - log model")
summary(fit_list[[2]])

plot(fit_list[[3]],ModTitle  = "Princess - log model")
summary(fit_list[[3]])

plot(fit_list[[4]],ModTitle  = "Katza - log model")
summary(fit_list[[4]])
```


```{r}

sar_log_sum<-data.frame("Site"= c("NR","Caves","Princess","Katza"))

sar_log_sum$z<-c(fit_list[[1]][["par"]][["z"]],
                 fit_list[[2]][["par"]][["z"]],
                 fit_list[[3]][["par"]][["z"]],
                 fit_list[[4]][["par"]][["z"]])

sar_log_sum$c<-c(fit_list[[1]][["par"]][["c"]],
                 fit_list[[2]][["par"]][["c"]],
                 fit_list[[3]][["par"]][["c"]],
                 fit_list[[4]][["par"]][["c"]])

sar_log_sum$r<-c(fit_list[[1]][["R2"]],
                 fit_list[[2]][["R2"]],
                 fit_list[[3]][["R2"]],
                 fit_list[[4]][["R2"]])

sar_log_sum
```

#### Power model

c×A^z

```{r}
power_fit_list<-list()


for( i in unique(species_knoll$Site)){
 
  print(i)
  
  site_data <- species_knoll %>% 
    filter(Site == i) %>% select(Surface.Area,Species_number) 
  
  site_data<-na.omit(site_data)

  power_fit_list[[i]]<- sar_power(site_data)
  
  
}

plot(power_fit_list[[1]],ModTitle  = "NR - power model")
summary(power_fit_list[[1]])

plot(power_fit_list[[2]],ModTitle  = "Caves - power model")
summary(power_fit_list[[2]])

plot(power_fit_list[[3]],ModTitle  = "Princess - power model")
summary(power_fit_list[[3]])

plot(power_fit_list[[4]],ModTitle  = "Katza - power model")
summary(power_fit_list[[4]])
```



```{r}

sar_power_sum<-data.frame("Site"= c("NR","Caves","Princess","Katza"))

sar_power_sum$z<-c(power_fit_list[[1]][["par"]][["z"]],
                 power_fit_list[[2]][["par"]][["z"]],
                 power_fit_list[[3]][["par"]][["z"]],
                 power_fit_list[[4]][["par"]][["z"]])

sar_power_sum$c<-c(power_fit_list[[1]][["par"]][["c"]],
                 power_fit_list[[2]][["par"]][["c"]],
                 power_fit_list[[3]][["par"]][["c"]],
                 power_fit_list[[4]][["par"]][["c"]])

sar_power_sum$r<-c(power_fit_list[[1]][["R2"]],
                 power_fit_list[[2]][["R2"]],
                 power_fit_list[[3]][["R2"]],
                 power_fit_list[[4]][["R2"]])

sar_power_sum

```

### Sar models with mean values per knoll:

```{r}

mean_species_knoll<-species_knoll %>% group_by(knoll_id) %>% mutate(Mean_species_number = mean(Species_number)) %>% ungroup() %>% distinct(knoll_id,.keep_all = T)

```


log model


```{r}
fit_list_mean<-list()


for( i in unique(mean_species_knoll$Site)){
 
  print(i)
  
  site_data <- mean_species_knoll %>% 
    filter(Site == i) %>% select(Surface.Area,Species_number) 
  
  site_data<-na.omit(site_data)

  fit_list_mean[[i]]<- sar_loga(site_data)
  
  
}

plot(fit_list_mean[[1]],ModTitle  = "NR - mean log model")
plot(fit_list_mean[[2]],ModTitle  = "Caves - meanlog model")
plot(fit_list_mean[[3]],ModTitle  = "Princess - mean log model")
plot(fit_list_mean[[4]],ModTitle  = "Katza - mean log model")




sar_mean_log_sum<-data.frame("Site"= c("NR","Caves","Princess","Katza"))

sar_mean_log_sum$z<-c(fit_list_mean[[1]][["par"]][["z"]],
                 fit_list_mean[[2]][["par"]][["z"]],
                 fit_list_mean[[3]][["par"]][["z"]],
                 fit_list_mean[[4]][["par"]][["z"]])

sar_mean_log_sum$c<-c(fit_list_mean[[1]][["par"]][["c"]],
                 fit_list_mean[[2]][["par"]][["c"]],
                 fit_list_mean[[3]][["par"]][["c"]],
                 fit_list_mean[[4]][["par"]][["c"]])

sar_mean_log_sum$r<-c(fit_list_mean[[1]][["R2"]],
                 fit_list_mean[[2]][["R2"]],
                 fit_list_mean[[3]][["R2"]],
                 fit_list_mean[[4]][["R2"]])

sar_mean_log_sum
```

