---
title: "2.3 species area relationship"
author: "TG"
date: "31 1 2021"
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)


```


### general steps:

1) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

sp_area<-fish_full_data
sp_area$Confidence[is.na(sp_area$Confidence)]<-(0)
sp_area<-sp_area %>% filter(Confidence <2)

```


2) remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

sp_area<-sp_area %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")

```

3) keep one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(sp_area$survey_id)) {
  
  data <- sp_area %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```

# first combination

* all fish(transient + cryptic)
* only one observer
* y = surface area



4) set the number of species per knoll

* note - Im using here the "down-graded" data - fish are grouped according to 2016 sp list


```{r}

species_knoll<- one_observer %>% group_by(survey_id) %>% mutate(Species_number = n_distinct(Species_2015)) %>% select(-Species,-Species_grouped,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%  ungroup() %>% distinct(.keep_all = T)

```

5) ill leave 2017 where only kazza was sampled out

```{r}
species_knoll<-species_knoll %>% filter(Year != 2017)


```

6) split 2020 to two surveys

```{r}
species_knoll <- species_knoll %>% 
                 mutate(year_month = ymd(paste(Year,month(Date),"1")))
```



7) model

```{r}

sp_surface_area<-list()

sp_surface_area[[1]]<-lmer(formula = log(Species_number) ~ log(Surface.Area)*Site + Max.depth +(1|year_month),data = species_knoll)

sp_surface_area[[2]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site + Max.depth + year_month,data = species_knoll)

sp_surface_area[[3]]<-lmer(formula = log(Species_number) ~ log(Surface.Area)*Site + (1|year_month),data = species_knoll)

sp_surface_area[[4]]<-lmer(formula = log(Species_number) ~ log(Surface.Area):Site + Max.depth +(1|year_month),data = species_knoll)

sp_surface_area[[5]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site*year_month + Max.depth ,data = species_knoll)


```

```{r message=FALSE, warning=FALSE, include=FALSE}
summary(sp_surface_area[[1]])
r.squaredGLMM(sp_surface_area[[1]])

summary(sp_surface_area[[2]])
r.squaredGLMM(sp_surface_area[[2]])

summary(sp_surface_area[[3]])
r.squaredGLMM(sp_surface_area[[3]])

summary(sp_surface_area[[4]])
r.squaredGLMM(sp_surface_area[[4]])

summary(sp_surface_area[[5]])
r.squaredGLMM(sp_surface_area[[5]])

aic<-model.sel(sp_surface_area[[1]],sp_surface_area[[2]],sp_surface_area[[3]],sp_surface_area[[4]],sp_surface_area[[5]])
```

```{r}
aic
```


```{r}
hist(sp_area$Max.depth)

mean(sp_area$Max.depth,na.rm=T)
median(sp_area$Max.depth,na.rm=T)

hist(sp_area$Surface.Area)



```

###the chosen model:

log(Species_number) ~ log(Surface.Area)*Site + (1|year_month)

create predict table:

```{r}

surface_area = rep(seq(from = 5,to = 40,by = 5),24) # Length = 8
site = rep(unique(species_knoll$Site),each = 48) # Length = 4
year = rep(rep(unique(species_knoll$year_month),each = 8),4) # Length = 6

  
predict_sa<-data.frame(Site = site,
                       year_month = year,
                       Surface.Area= surface_area,
                       Max.depth = rep(4))

predict_sa$Surface.Area_log<-log(predict_sa$Surface.Area)

predict_sp<-predict(sp_surface_area[[3]],predict_sa,type="response",se.fit = T)

predicted_surface<-cbind(predict_sa,predict_sp)
predicted_surface$species_number<-exp(predicted_surface$predict_sp)

predicted_surface$site_year<-paste(predicted_surface$Site,predicted_surface$year_month,sep="_")

predicted_surface$year_month<- as.factor(predicted_surface$year_month)

```


```{r echo=FALSE}
ggplot(data = predicted_surface,aes(x= Surface.Area_log,y = predict_sp,color= site_year))+
  geom_line(size=1)+theme_classic()


ggplot(data = predicted_surface,aes(x= Surface.Area,y = species_number,color= site_year))+
  geom_line(size=1)+theme_classic()

```

for each site


```{r echo=FALSE, message=FALSE, warning=FALSE}


for (i in unique(predicted_surface$Site)) {
  data<-predicted_surface %>% filter(Site == i)
  
  print(ggplot(data = data,aes(x= Surface.Area,y = species_number,color= year_month))+
  geom_line(size=1)+
    ylim(0,60)+
    theme_classic()+
    ggtitle(i)+
    scale_colour_manual(values=c("#56dde0", "#33658a", "#f5af2d","#f2641a","#5da446","#c9e32a")))
  
}



  
  
```



for each year

```{r}


for (i in unique(predicted_surface$year_month)) {
  data<-predicted_surface %>% filter(year_month == i)
  print(ggplot(data = data,aes(x= Surface.Area,y = species_number,color= Site))+
  geom_line(size=1)+ylim(0,50)+theme_classic()+ggtitle(i)+
    scale_colour_manual(values=c("#ed6028", "#159ed2", "#16a647","#c40032","#262d35"))
    )
  
}


```


```{r}
rm(list=setdiff(ls(), c("one_observer","species_knoll","sp_area")))

```

# Second combination

* without Transient species
* only one observer
* y = surface area


```{r}
only_cryptic <- one_observer %>% filter(Status != "T")

```


3) set the number of species per knoll


```{r}

species_knoll_cry<- only_cryptic %>% group_by(survey_id) %>% mutate(Species_number = n_distinct(Species_2015)) %>% select(-Species,-Species_grouped,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%  ungroup() %>% distinct(.keep_all = T)

```

* ill leave 2007 where only kazza was sampled out

```{r}
species_knoll_cry<-species_knoll_cry %>% filter(Year != 2017)


```

4) split 2020 to two surveys

```{r}
species_knoll_cry <- species_knoll_cry %>% 
                 mutate(year_month = ymd(paste(Year,month(Date),"1")))
```


 

4) model

```{r}

cryptic_list<-list()

cryptic_list[[1]]<-lmer(formula = log(Species_number) ~ log(Surface.Area)*Site + Max.depth +(1|year_month),data = species_knoll_cry)

cryptic_list[[2]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site + Max.depth + year_month,data = species_knoll_cry)

cryptic_list[[3]]<-lmer(formula = log(Species_number) ~ log(Surface.Area)*Site + (1|year_month),data = species_knoll_cry)

cryptic_list[[4]]<-lmer(formula = log(Species_number) ~ log(Surface.Area):Site + Max.depth +(1|year_month),data = species_knoll_cry)

cryptic_list[[5]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site*year_month + Max.depth ,data = species_knoll_cry)


```

```{r message=FALSE, warning=FALSE, include=FALSE}
summary(cryptic_list[[1]])
r.squaredGLMM(cryptic_list[[1]])

summary(cryptic_list[[2]])
r.squaredGLMM(cryptic_list[[2]])

summary(cryptic_list[[3]])
r.squaredGLMM(cryptic_list[[3]])

summary(cryptic_list[[4]])
r.squaredGLMM(cryptic_list[[4]])

summary(cryptic_list[[5]])
r.squaredGLMM(cryptic_list[[5]])

aic_cry<-model.sel(cryptic_list[[1]],cryptic_list[[2]],cryptic_list[[3]],cryptic_list[[4]],cryptic_list[[5]])

```

### the chosen model:

log(Species_number) ~ log(Surface.Area)*Site + (1|year_month),data = species_knoll_cry

create predict table

```{r}

surface_area = rep(seq(from = 5,to = 40,by = 5),20) # Length = 8
site = rep(unique(species_knoll$Site),each = 40) # Length = 4
year = rep(rep(unique(species_knoll$year_month),each = 8),4) # Length = 5

  
predict_cryptic<-data.frame(Site = site,
                       year_month = year,
                       Surface.Area= surface_area)




predict_cryptic$Surface.Area_log<-log(predict_cryptic$Surface.Area)


predict_species_cryptic<-predict(cryptic_list[[3]],predict_cryptic,type="response")

cyptic_predicted<-cbind(predict_cryptic,predict_species_cryptic)

cyptic_predicted$site_year<-paste(cyptic_predicted$Site,cyptic_predicted$year_month,sep="_")

cyptic_predicted$species_number<-exp(cyptic_predicted$predict_species_cryptic)


cyptic_predicted$year_month<- as.factor(cyptic_predicted$year_month)



```



```{r}
ggplot(data = cyptic_predicted,aes(x= Surface.Area_log,y = predict_species_cryptic,color= site_year))+
  geom_line(size=1)+theme_classic()

ggplot(data = cyptic_predicted,aes(x= Surface.Area,y = species_number,color= site_year))+
  geom_line(size=1)+theme_classic()


  
```

for each site


```{r}


for (i in unique(cyptic_predicted$Site)) {
  data<-cyptic_predicted %>% filter(Site == i)
  
  print(
    ggplot(data = data,aes(x= Surface.Area,y = species_number,color= year_month))+
    geom_line(size=1)+ylim(1,30)+
    theme_classic()+
    ggtitle(i)+
    scale_colour_manual(values=c("#56dde0", "#33658a", "#f5af2d","#f2641a","#5da446")))
  
  
  ggsave(filename =paste(i,".png"))
  
}



  
```



for each year

```{r}


for (i in unique(cyptic_predicted$year_month)) {
  data<-cyptic_predicted %>% filter(year_month == i)
  print(ggplot(data = data,aes(x= Surface.Area,y = species_number,color= Site))+
  geom_line(size=1)+ylim(1,40)+theme_classic()+ggtitle(i)+
    scale_colour_manual(values=c("#ed6028", "#159ed2", "#16a647","#c40032")))
  
}


```




```{r}
rm(list=setdiff(ls(), c("one_observer","species_knoll","sp_area","species_knoll_cry")))

```


# third comperison - storm

before and after the storm...2018 & 2020 
only cryptic species
but ill use the full species name and not the group ones

```{r}
storm_knolls<- one_observer %>% filter(Status != "T") %>%  group_by(survey_id) %>% mutate(Species_number = n_distinct(Species)) %>% select(-Species,-Species_grouped,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%  ungroup() %>% distinct(.keep_all = T)
```

```{r}
storm_knolls<- storm_knolls %>% filter(Year == 2020|Year ==2018) %>% mutate(month = month(Date))

storm_knolls <- storm_knolls %>% mutate(year_month = ymd(paste(Year,month,"1")))
storm_knolls$year_month<-as.factor(storm_knolls$year_month)

storm_knolls$season<-ifelse(storm_knolls$month == 6 ,"summer","fall")
storm_knolls$storm <- ifelse(storm_knolls$month == 9 ,"before","after")
```

```{r}

storm_list<-list()

storm_list[[1]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site + Max.depth + year_month,data = storm_knolls)

storm_list[[2]]<-lmer(formula = log(Species_number) ~ log(Surface.Area)*Site + Max.depth + (1|year_month) ,data = storm_knolls)

storm_list[[3]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site*year_month,data = storm_knolls)

storm_list[[4]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*Site*year_month +Max.depth ,data = storm_knolls)


```

```{r message=FALSE, warning=FALSE, include=FALSE}

aic_storm<-model.sel(storm_list[[1]],storm_list[[2]],storm_list[[3]],storm_list[[4]])
```

```{r}
aic_storm
```

### chosen model

log(Species_number) ~ log(Surface.Area)*Site + year_month + Max.depth

create predict table

```{r}
surface_area = rep(seq(from = 5,to = 40,by = 5),12) # Length = 8
site = rep(unique(storm_knolls$Site),each = 24) # Length = 4
year_month = rep(rep(unique(storm_knolls$year_month),each = 8),4) # Length = 3

  
predict_storm<-data.frame(Site = site,
                       year_month = year_month,
                       Surface.Area=surface_area,
                       Max.depth = rep(4))

predict_storm$Surface.Area_log<-log(predict_storm$Surface.Area)

predict_storm_sp<-predict(storm_list[[1]],predict_storm,type="response")

predict_storm_sp<-cbind(predict_storm,predict_storm_sp)

predict_storm_sp$year_month<-as.factor(predict_storm_sp$year_month)

predict_storm_sp$date_site<-paste(predict_storm_sp$Site,predict_storm_sp$year_month,sep=" ")


predict_storm_sp$species_number<-exp(predict_storm_sp$predict_storm_sp)

ggplot(data = predict_storm_sp,aes(x= Surface.Area_log,y = predict_storm_sp,color= date_site))+
  geom_line(size=1)+theme_classic()
```

```{r}
ggplot(data = predict_storm_sp,aes(x= Surface.Area_log,y = predict_storm_sp,color= date_site))+
  geom_line(size=1)+theme_classic()

ggplot(data = predict_storm_sp,aes(x= Surface.Area,y = species_number,color= date_site))+
  geom_line(size=1)+theme_classic()
```

for each site


```{r echo=FALSE, message=FALSE, warning=FALSE}


for (i in unique(predict_storm_sp$Site)) {
  data<-predict_storm_sp %>% filter(Site == i)
  
  print(ggplot(data = data,aes(x= Surface.Area,y = species_number,color= year_month))+
  geom_line(size=1)+
    ylim(0,60)+
    theme_classic()+
    ggtitle(i)+
    scale_colour_manual(values=c("#56dde0", "#33658a", "#f5af2d","#f2641a","#5da446")))
  
}



  
  
```

for each year

```{r}


for (i in unique(predict_storm_sp$year_month)) {
  data<-predict_storm_sp %>% filter(year_month == i)
  print(ggplot(data = data,aes(x= Surface.Area,y = species_number,color= Site))+
  geom_line(size=1)+ylim(0,40)+theme_classic()+ggtitle(i)+
    scale_colour_manual(values=c("#ed6028", "#159ed2", "#16a647","#c40032"))
    )
  
}


```

 
 
```{r}
rm(list=setdiff(ls(), c("one_observer","species_knoll","sp_area","species_knoll_cry")))

```


 
# forth comperison - Katza


```{r}
species_kazza_knoll<- one_observer %>%filter(Status!= "T") %>% group_by(survey_id) %>% mutate(Species_number = n_distinct(Species_2015)) %>% select(-Species,-Species_grouped,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%  ungroup() %>% distinct(.keep_all = T)
```


```{r}
species_kazza_knoll<- species_kazza_knoll %>% filter(Site == "Katza") %>% filter()

species_kazza_knoll <- species_kazza_knoll %>% 
                 mutate(year_month = ymd(paste(Year,month(Date),"1")))

species_kazza_knoll$opening <- ifelse(year(species_kazza_knoll$year_month) < 2017 ,"before","after")


```



```{r}
katza_list<-list()

katza_list[[1]]<-lmer(formula = log(Species_number) ~ log(Surface.Area) + Max.depth  +(1|year_month),data = species_kazza_knoll)

katza_list[[2]]<-lm(formula = log(Species_number) ~ log(Surface.Area) + Max.depth + year_month + opening,data = species_kazza_knoll)

katza_list[[3]]<-lmer(formula = log(Species_number) ~ log(Surface.Area) + (1|year_month),data = species_kazza_knoll)

katza_list[[4]]<-lm(formula = log(Species_number) ~ log(Surface.Area)*year_month + Max.depth ,data = species_kazza_knoll)
```


```{r message=FALSE, warning=FALSE, include=FALSE}
summary(katza_list[[1]])
r.squaredGLMM(katza_list[[1]])

summary(katza_list[[2]])
r.squaredGLMM(katza_list[[2]])

summary(katza_list[[3]])
r.squaredGLMM(katza_list[[3]])

aic_katza<-model.sel(katza_list[[1]],katza_list[[2]],katza_list[[3]],katza_list[[4]])
```

```{r}
aic_katza
```

```{r}
surface_area = rep(seq(from = 5,to = 40,by = 5),6) # Length = 8

year_month = rep(unique(species_kazza_knoll$year_month),each = 8) # Length = 6

  
predict_katza<-data.frame(year_month = year_month,
                         Surface.Area = surface_area,
                         Max.depth = rep(4))

predict_katza$Surface.Area_log<-log(predict_katza$Surface.Area)

predict_katza_sp<-predict(katza_list[[3]],predict_katza,type="response")

predict_katza_sp<-cbind(predict_katza,predict_katza_sp)

predict_katza_sp$year_month<-as.factor(predict_katza_sp$year_month)

predict_katza_sp$species_number<-exp(predict_katza_sp$predict_katza_sp)

ggplot(data = predict_katza_sp,aes(x= Surface.Area,y = species_number,color= year_month))+
  geom_line(size=1)+theme_classic()
```

