---
title: "2.5 fish species area"
author: "TG"
date: "23 12 2021"
output: html_document
---

 

```{r}
load("fish_full_data.rdata")
select <- dplyr::select
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)

fish_k<-c("#669bbc","#336683","#1a4b66","#003049","#ed3140","#c1121f","#710912")
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")


```


### general steps:

1) filter confidence = 1

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

sp_area<-fish_full_data
sp_area$Confidence[is.na(sp_area$Confidence)]<-(0)
sp_area<- sp_area %>% filter(Confidence < 1)

```


2) remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

sp_area<-sp_area %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")

```

3) keep one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(sp_area$survey_id)) {
  
  data <- sp_area %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```



```{r}
sp_list<-fish_full_data %>% distinct(Species,.keep_all = T) %>% select(Species,Species_2015,Family,Status) %>% arrange(Family,Species)

```


the connection between resident fish and surface area


```{r}
only_cryptic <- one_observer %>% filter(Status != "T")

```


3) set the number of species per knoll


```{r}

species_knoll<- only_cryptic %>% group_by(survey_id) %>% mutate(Species_number = n_distinct(Species_2015)) %>% select(-Species,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%  ungroup() %>% distinct(.keep_all = T)

```

* ill leave 2007 where only kazza was sampled out

```{r}
species_knoll<-species_knoll %>% filter(Year != 2017)


```

add month 

```{r}
species_knoll$month<-month(species_knoll$Date)
species_knoll$month<-as.factor(species_knoll$month)

species_knoll$storm<-as.factor(species_knoll$storm)
species_knoll$Season <- ifelse(species_knoll$month == 6,"Spring","Fall")

species_knoll$Year<-as.factor(species_knoll$Year)
species_knoll$Site<-as.factor(species_knoll$Site)
```



remove outlier knolls

```{r}

no_nr_outliar<-species_knoll %>% filter(knoll_id != "Katza_19",knoll_id != "NR_17")

```

best model from previos scrips (2.4 species area)


```{r}

model<-lmer(formula = log(Species_number) ~ log(Surface_Area)*Site +(1|year_month)+Max_depth ,data = no_nr_outliar)
summary(model)
r.squaredGLMM(model)

model_2<-lm(formula = log(Species_number) ~ log(Surface_Area)*Site +year_month + Max_depth ,data = no_nr_outliar)
summary(model_2)


model_3<-lm(formula = log(Species_number) ~ log(Surface_Area)*year_month + Site + Max_depth ,data = no_nr_outliar)
summary(model_3)
r.squaredGLMM(model_3)

model_5<-lm(formula = log(Species_number) ~ log(Surface_Area)*year_month ,data = no_nr_outliar)
summary(model_5)
r.squaredGLMM(model_5)

model_5_gg=ggeffects::ggpredict(model_5,terms = c('Surface_Area[1:66 by=1]','year_month'))

ggplot()+
  theme_classic()+
  geom_line(data = model_5_gg,aes(x= log(x),y = log(predicted),color= group),size=1.5)+ 
  geom_point(data = species_knoll,aes(log(Surface_Area),log(Species_number),color = year_month),alpha = 0.5)+geom_ribbon(data = model_5_gg,aes(x=log(x),ymin=log(conf.low),ymax=log(conf.high),fill=group),alpha=0.1)+
  ylab("log(richness)")+xlim(1,4)+ylim(-1,4)+
  xlab(bquote('log(Knoll surface area '~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))+
        scale_fill_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))

#ggsave("residence_fish_area.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts")


ggplot()+
  theme_classic()+
  geom_line(data = model_5_gg,aes(x= x,y = predicted,color= group),size=1.5)+ 
  geom_point(data = species_knoll,aes(Surface_Area,Species_number,color = year_month),alpha = 0.5)+geom_ribbon(data = model_5_gg,aes(x=x,ymin=conf.low,ymax=conf.high,fill=group),alpha=0.1)+
  ylab("Richness")+xlim(1,40)+
  xlab(bquote('Knoll surface area '~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))+
        scale_fill_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))

sjPlot::plot_model(model_5,type="eff",terms = c('Surface_Area','year_month'),show.data = T)+theme_classic()


model_4<-lm(formula = log(Species_number) ~ log(Surface_Area)+
              year_month + Site + Max_depth ,data = no_nr_outliar)
summary(model_4)
r.squaredGLMM(model_4)

model.sel(model_2,model_3,model_4)



```

```{r}

surface_area = rep(seq(from = 1,to = 60,by = 1),24) # Length = 60
site = rep(unique(species_knoll$Site),each = 360) # Length = 4
year = rep(rep(unique(species_knoll$year_month),each = 60),4) # Length = 6

predict_no_out<-data.frame(Site = site,
                       year_month = year,
                       Surface_Area= surface_area,
                       Max_depth = rep(4.6))


predict_no_out$Surface_Area_log<-log(predict_no_out$Surface_Area)


predict_species_no_out<-predict(model_3,predict_no_out,type="response")

predict_no_out<-cbind(predict_no_out,predict_species_no_out)

predict_no_out$site_year<-paste(predict_no_out$Site,predict_no_out$year_month,sep="_")

predict_no_out$species_number<-exp(predict_no_out$predict_species_no_out)


predict_no_out$year_month<- as.factor(predict_no_out$year_month)

l_type<-c("solid","solid","solid","twodash","solid","solid")

ggplot(data = predict_no_out,aes(x= Surface_Area,y = species_number,color= site))+
  theme_classic()+geom_smooth(alpha = 0.2)+
  geom_point(data = no_nr_outliar,aes(Surface_Area,Species_number,color = Site),alpha = 0.3)+
  ylab("Richness")+
  xlab(bquote('Knoll surface area '~(m^2)))+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 02))+
  scale_color_manual(name = "Site",values = color)

#ggsave("residence_fish_sp_area.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")


ggplot(data = predict_no_out,aes(x= Surface_Area,y = species_number,color= year_month))+
  theme_classic()+geom_smooth(alpha = 0.2)+   geom_point(data = species_knoll,aes(Surface_Area,Species_number,color = year_month),alpha = 0.3)+
  ylab("Richness")+
  xlab(bquote('Knoll surface area '~(m^2)))+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))


ggplot(data = predict_no_out,aes(x= Surface_Area_log,y = predict_species_no_out,color= year_month))+
  theme_classic()+geom_smooth(alpha = 0.2)+ 
  geom_point(data = species_knoll,aes(log(Surface_Area),log(Species_number),color = year_month),alpha = 0.5)+
  ylab("log(richness)")+
  xlab(bquote('log(Knoll surface area '~(m^2)))+
  xlim(1,4)+
  theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))

#ggsave("residence_fish_sp_area_year_interaction.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts")
```

now do one for the diffreny years

```{r}
surface_area = rep(seq(from = 1,to = 60,by = 1),24) # Length = 60
site = rep(unique(species_knoll$Site),each = 360) # Length = 4
year = rep(rep(unique(no_nr_outliar$year_month),each = 60),4) # Length = 6




mean(species_knoll$Max_depth)
median(species_knoll$Max_depth)
  
predict_cryptic<-data.frame(Site = site,
                       year_month = year,
                       Surface_Area= surface_area,
                       Max_depth = rep(4.6))




predict_cryptic$Surface_Area_log<-log(predict_cryptic$Surface_Area)

predict_species_cryptic<-predict(sp_ar_model,predict_cryptic,type="response")

cyptic_predicted<-cbind(predict_cryptic,predict_species_cryptic)

cyptic_predicted$site_year<-paste(cyptic_predicted$Site,cyptic_predicted$year_month,sep="_")

cyptic_predicted$species_number<-exp(cyptic_predicted$predict_species_cryptic)


cyptic_predicted$year_month<- as.factor(cyptic_predicted$year_month)


ggplot(data = cyptic_predicted,aes(x= Surface_Area,y = species_number,color= site))+
  theme_classic()+geom_smooth()+geom_point()




ggplot(data = cyptic_predicted,aes(x= Surface_Area,y = species_number,color= year_month))+
  theme_classic()+geom_smooth()+geom_point(data = species_knoll,aes(Surface_Area,Species_number,color = year_month),alpha = 0.3)+
  ylab("Species number")+
  xlab(bquote('Knoll surface area '~(m^2)))+
  theme(axis.text.x = element_text(size=16),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title = element_text(size=18),
        legend.text = element_text(size=16),
        strip.text.x = element_text(size = 16))+
  scale_color_manual(values=fish_all)
```



