---
title: "20.1 nMDS for ms"
author: "TG"
date: "17 4 2022"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
#library(Rarefy)
library(rareNMtests)
library(zetadiv)
library(FuzzyQ)

```



```{r}
pairwise.adonis <- function(x,factors, sim.function = 'vegdist', sim.method = 'bray', p.adjust.m ='bonferroni',reduce=NULL,perm=999)
{
 
  co <- combn(unique(as.character(factors)),2)
  pairs <- c()
  Df <- c()
  SumsOfSqs <- c()
  F.Model <- c()
  R2 <- c()
  p.value <- c()
 
 
  for(elem in 1:ncol(co)){
    if(inherits(x, 'dist')){
      x1=as.matrix(x)[factors %in% c(as.character(co[1,elem]),as.character(co[2,elem])),
                      factors %in% c(as.character(co[1,elem]),as.character(co[2,elem]))]
      }
 
    else  (
      if (sim.function == 'daisy'){
            x1 = daisy(x[factors %in% c(co[1,elem],co[2,elem]),],metric=sim.method)
        }
      else{x1 = vegdist(x[factors %in% c(co[1,elem],co[2,elem]),],method=sim.method)}
    )
   
    x2 = data.frame(Fac = factors[factors %in% c(co[1,elem],co[2,elem])])
   
    ad <- adonis2(x1 ~ Fac, data = x2,
                 permutations = perm);
    pairs <- c(pairs,paste(co[1,elem],'vs',co[2,elem]));
    Df <- c(Df,ad$Df[1])
SumsOfSqs <- c(SumsOfSqs,ad$SumOfSqs[1])
F.Model <- c(F.Model,ad$F[1]);
    R2 <- c(R2,ad$R2[1]);
    p.value <- c(p.value,ad$`Pr(>F)`[1])
  }
  p.adjusted <- p.adjust(p.value,method=p.adjust.m)
 
  sig = c(rep('',length(p.adjusted)))
  sig[p.adjusted <= 0.05] <-'.'
  sig[p.adjusted <= 0.01] <-'*'
  sig[p.adjusted <= 0.001] <-'**'
  sig[p.adjusted <= 0.0001] <-'***'
  pairw.res <- data.frame(pairs,Df,SumsOfSqs,F.Model,R2,p.value,p.adjusted,sig)
 
  if(!is.null(reduce)){
    pairw.res <- subset (pairw.res, grepl(reduce,pairs))
    pairw.res$p.adjusted <- p.adjust(pairw.res$p.value,method=p.adjust.m)
   
    sig = c(rep('',length(pairw.res$p.adjusted)))
  sig[pairw.res$p.adjusted <= 0.1] <-'.'
sig[pairw.res$p.adjusted <= 0.05] <-'*'
sig[pairw.res$p.adjusted <= 0.01] <-'**'
sig[pairw.res$p.adjusted <= 0.001] <-'***'
    pairw.res <- data.frame(pairw.res[,1:7],sig)
  }
  class(pairw.res) <- c("pwadonis", "data.frame")
  return(pairw.res)
}

#exp_ado_genus = adonis(exp_diss_mat~genus, data = exp_mds_df,                       permutations = 1000); exp_ado_genus$aov.tab  #this is basically a #PERMANOVA analysis

#pairwise_exp_genus = pairwise.adonis(x=exp_diss_mat,factors=exp_mds_df$genus,                                     p.adjust.m='bonferroni')


```

color scales definitions 

```{r}
color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")
  
save_path<-"C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\nmds"  

```


# Data prepartion

```{r}
load("fish_full_data.rdata")
select <- dplyr::select



```



```{r}

fish_prep<-fish_full_data
```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
filter(Site !="North beach")

```

4. keep data from 2018 and on

```{r}
fish_prep<-fish_prep %>% filter(Year >2017)
```


4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```


7. keep only the knolls that were surveyed in 2020 spring (storm)



```{r}



knolls_2020<-fish_prep %>% filter(Year %in% c("2020")) %>% distinct(knoll_id)



fish_prep<-fish_prep %>% filter(knoll_id %in% knolls_2020$knoll_id)

```

7. separate to all fish and resident fish only


```{r}

all_fish <- fish_prep 
resident_fish <- fish_prep %>% filter(Status != "T")

```

8. wide data


```{r}

wide <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
    summarise(Abundance=sum(Abundance))

wide <- spread(wide,
               Species,
               Abundance,
               fill = 0)




```

## nmds no transformation:

```{r}
 
  
  nmds <- wide[,4:ncol(wide)] #only species

  ord<- metaMDS(nmds,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
  print(ord$stress)
  Year <- factor(wide$year_month)
  stressplot(ord)

pdf(file = "C:\\Users\\TalG\\My Drive\\PHD\\chapter 1 strom effect\\R Scripts\\plots for ms\\nmds\\allsitenmds.pdf",   # The directory you want to save the file in
    width = 5, # The width of the plot in inches
    height = 5)

ordiplot(ord, type = "n")
orditorp(ord, label = T, display = "sites", col = color_op6[Year],pch = 16)
ordiellipse(ord, groups = wide$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_op6)
legend("bottomright", legend = levels(Year), bty = "n", col = color_op6, pch = 15,cex=1.5)

dev.off()



dis <- vegdist(nmds,method="bray")

mod <- betadisper(dis,wide$year_month)

anova(mod)


plot(TukeyHSD(mod))


plot(mod)

adon.results<-adonis(dis ~  year_month,data = wide, method="bray",perm=999)
adon.results

pairwise_asonid = pairwise.adonis(x=dis,factors=wide$year_month,
                                     p.adjust.m='bonferroni')
str(nmds)
#ano<-anosim(x = nmds_data, grouping = wide$year_month, permutations = 9999, distance = "bray") 
#ano

```
### split by site

```{r}


tucky_res<-list()


for (i in unique(wide$Site)) {
  
  data <- wide  %>% filter(Site == i)
  
  nmds_data <- data[,4:ncol(data)] #only species

  ord<- metaMDS(nmds_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
 
  print(i)
  print(ord$stress)
  Year <- factor(data$year_month)
  stressplot(ord)
  
plot_name<-paste0(save_path,"\\",i,"_no_trans_all_fish",".pdf")

#pdf(file = plot_name,   # The directory you want to save the file in
#    width = 5,          # The width of the plot in inches
#    height = 5) 

ordiplot(ord, type = "n",main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_op6[Year],pch = 16)
ordiellipse(ord, groups = data$year_month,kind = "sd", draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_op6)
legend("bottomright", legend = levels(Year), bty = "n", col = color_op6, pch = 15,cex=1.5)

#dev.off()

print(i)

dis <- vegdist(nmds_data,method="bray")
adon.results<-adonis(dis ~  year_month,data = data, method="bray")

pairwise_asonid = pairwise.adonis(x=nmds_data,factors=data$year_month,
                                     p.adjust.m='bonferroni')

print(pairwise_asonid)
print(adon.results) 

print("adonis end")




mod <- betadisper(dis,data$year_month)

print(anova(mod))

res<-TukeyHSD(mod)
res<- as.data.frame(res[["group"]])
res$sample <- rownames(res)



tucky_res[[i]]<-res

print(TukeyHSD(mod))

plot(TukeyHSD(mod))

plot(mod)
  
  
}

#openxlsx::write.xlsx(tucky_res, file = 'tucky_res.xlsx')
```

* adonis - test differences in centroids and data dispersion

https://chrischizinski.github.io/rstats/adonis/


### split by site for resident fish only


```{r}
wide_res <- resident_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
    summarise(Abundance=sum(Abundance))

wide_res <- spread(wide_res,
               Species,
               Abundance,
               fill = 0)
```


```{r}

for (i in unique(wide_res$Site)) {
  
  data <- wide_res  %>% filter(Site == i)
  
  nmds_data <- data[,4:ncol(data)] #only species

  ord<- metaMDS(nmds_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
 
   print(i)
  print(ord$stress)
  Year <- factor(data$year_month)
  stressplot(ord)
  
plot_name<-paste0(save_path,"\\",i,"_no_trans_resident_fish",".pdf")

#pdf(file = plot_name,   # The directory you want to save the file in
#    width = 5,          # The width of the plot in inches
#    height = 5) 

ordiplot(ord, type = "n",main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_op6[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_op6)
legend("bottomright", legend = levels(Year), bty = "n", col = color_op6, pch = 15,cex=1.5)

#dev.off()

print(i)
adon.results<-adonis(nmds_data ~  year_month,data = data, method="bray",perm=999)
print(adon.results) 
  
  
}


```


## nmds log transformation:

less weight for abundaces


```{r}


log_nmds <- decostand(nmds,"log", logbase = 10)

log_ord<- metaMDS(log_nmds,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
log_ord$stress
stressplot(log_ord)

  
ordiplot(log_ord, type = "n")
orditorp(log_ord, label = T, display = "sites", col = color_op6[Year],pch = 16)
ordiellipse(log_ord,groups = wide$year_month, draw = "polygon",alpha = 0.35,label=F,lty = 1, col = color_op6)
legend("bottomright", legend = levels(Year), bty = "n", col = color_op6, pch = 15,cex=1.5)


  
ordiplot(log_ord, type = "n")
orditorp(log_ord, label = T, display = "sites", col = color_op6[Year],pch = 16)
ordihull(log_ord, groups = wide$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_op6)
legend("bottomright", legend = levels(Year), bty = "n", col = color_op6, pch = 15,cex=1.5)

adon.log.results<-adonis(log_nmds ~  year_month,data = wide, method="bray",perm=999)
print(i)
adon.log.results


```
#for each site

```{r}

for (i in unique(wide$Site)) {
  
  data <- wide  %>% filter(Site == i)
  

  nmds_data <- data[,4:ncol(data)] #only species
  
  nmds_log_tran<-decostand(nmds_data,"log", logbase = 10)

  ord<- metaMDS(nmds_log_tran,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
  print(ord$stress)
  Year <- factor(data$year_month)
  stressplot(ord)

  
ordiplot(ord, type = "n",main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_op6[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_op6)
legend("bottomright", legend = levels(Year), bty = "n", col = color_op6, pch = 15,cex=1.5)

print(i)
adon.results<-adonis(nmds_log_tran ~  year_month,data = data, method="bray",perm=999)
print(adon.results) 
  
  
}
```

