---
title: "3.1 Size spectra"
author: "TG"
date: "26 1 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(sars)
library(lme4)
library(lmerTest)
library(MuMIn)
```


1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```


3) separate all abundance >1 to do biomass per fish estimation

```{r}
to_sep<-one_observer %>% filter(Abundance>1)
to_sep$Abundance<-as.integer(to_sep$Abundance)
sum(to_sep$Abundance)

to_sep <- to_sep %>% uncount(Abundance)
to_sep<-to_sep %>% tibble::add_column(Abundance=rep(1),.after = 9)
```


```{r}
individual_data<-one_observer %>% filter(Abundance==1)
individual_data<-bind_rows(individual_data,to_sep)
individual_data<-individual_data %>% arrange(Site,Date,Knoll)
rownames(individual_data)<-1:nrow(individual_data)
```

4) calculate the weight (g) per individual

```{r}
individual_data<- individual_data %>% mutate("weight" =  a*(Length^b),.after = 11)
```


5) plot size distrabiotions


```{r}
qplot(individual_data$weight,
      geom="histogram",
      main = "Fish weights distribution", 
      xlab = "Fish weight (g)",
      ylab = "Abundance",
      fill=I("blue"), 
      col=I("blue"), 
      alpha=I(.5))+theme_classic()


weight_violin<-ggplot(individual_data, aes(x=Site,y=log(weight) ,fill=Site)) +
  geom_violin()+ 
  stat_summary(fun=mean, geom="point")

weight_violin

Length_violin<-ggplot(individual_data, aes(x=Site,y=log(Length) ,fill=Site)) +
  geom_violin()+ 
  stat_summary(fun=mean, geom="point")

Length_violin

qplot(log(individual_data$weight),
      geom="histogram",
      main = "Fish weights distribution", 
      xlab = "log Fish weight (g)",
      ylab = "Abundance",
      fill=I("blue"), 
      col=I("blue"), 
      alpha=I(.5))+theme_classic()
```



# Size spectra by site & year

```{r}

size_sp_data<-individual_data

size_sp_data$class <- with(size_sp_data,(ceiling(log(weight,2)))) 


size_sp_group <- size_sp_data%>% 
  group_by(class,Site,Year) %>% 
  dplyr::summarize(sum_biomass = sum(weight),individuals_in_class= n())

size_sp_group<-na.omit(size_sp_group)

size_sp_group<-size_sp_group %>% unite("site_year",Site,Year,sep = "_")

size_sp_group$midpoint <- ((2^size_sp_group$class)+(2^(size_sp_group$class-1)))/2
size_sp_group$Width <- (2^((size_sp_group$class)))-((2^(size_sp_group$class-1)))
size_sp_group$corrected_biomass <- with(data = size_sp_group,expr = sum_biomass/Width)
size_sp_group$midpoint_log<- log10(size_sp_group$midpoint)



ggplot(size_sp_group,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(size_sp_group,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_line()+geom_point()


filter_size_sp_group<-size_sp_group %>% filter(class>(-2))%>% filter(class<(12))

ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_line()+geom_point()

```



# Size spectra by site 

```{r}
size_sp_data<-individual_data

size_sp_data$class <- with(size_sp_data,(ceiling(log(weight,2)))) 


size_sp_group <- size_sp_data%>% 
  group_by(class,Site) %>% 
  dplyr::summarize(sum_biomass = sum(weight),individuals_in_class= n())

size_sp_group<-na.omit(size_sp_group)



size_sp_group$midpoint <- (2^size_sp_group$class+2^(size_sp_group$class-1))/2
size_sp_group$Width <- (2^((size_sp_group$class)))-((2^(size_sp_group$class-1)))
size_sp_group$corrected_biomass <- with(data = size_sp_group,expr = sum_biomass/Width)
size_sp_group$midpoint_log<- log10(size_sp_group$midpoint)



ggplot(size_sp_group,aes(midpoint,log10(corrected_biomass),color = Site))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(size_sp_group,aes(midpoint,log10(corrected_biomass),color = Site))+geom_line()+geom_point()

filter_size_sp_group<-size_sp_group %>% filter(class>(0))%>% filter(class<(11))

ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = Site))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = Site))+geom_line()+geom_point()
```





# Size spectra by YEAR 

```{r}
size_sp_data<-individual_data

size_sp_data$class <- with(size_sp_data,(ceiling(log(weight,2)))) 


size_sp_group <- size_sp_data%>% 
  group_by(class,Year) %>% 
  dplyr::summarize(sum_biomass = sum(weight),individuals_in_class= n())

size_sp_group<-na.omit(size_sp_group)



size_sp_group$midpoint <- (2^size_sp_group$class+2^(size_sp_group$class-1))/2
size_sp_group$Width <- (2^((size_sp_group$class)))-((2^(size_sp_group$class-1)))
size_sp_group$corrected_biomass <- with(data = size_sp_group,expr = sum_biomass/Width)
size_sp_group$midpoint_log<- log10(size_sp_group$midpoint)



ggplot(size_sp_group,aes(midpoint,log10(corrected_biomass),color = Year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(size_sp_group,aes(midpoint,log10(corrected_biomass),color = Year))+geom_line()+geom_point()

filter_size_sp_group<-size_sp_group %>% filter(class>(0))%>% filter(class<(11))

ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = Year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = Year))+geom_line()+geom_point()


```


* probably there is an artifact since in 2016-2015 the length were log2...

maybe i should group all lenghts to log 2 and then try it again

add lenght class


```{r}


individual_data$lenght_class <- with(individual_data,(ceiling(log(Length,2))))

individual_data<-individual_data %>%
  mutate(lenght_class =(2^lenght_class))


individual_data<-individual_data %>%
  mutate(weight_by_class = a*(lenght_class^b))

```






```{r}

size_sp_data_gg<-individual_data

size_sp_data_gg$class <- with(size_sp_data_gg,(ceiling(log(weight_by_class,2)))) 


size_sp_data_gg <- size_sp_data_gg%>% 
  group_by(class,Site,Year) %>% 
  dplyr::summarize(sum_biomass = sum(weight_by_class),individuals_in_class= n())

size_sp_data_gg<-na.omit(size_sp_data_gg)

size_sp_data_gg<-size_sp_data_gg %>% unite("site_year",Site,Year,sep = "_")

size_sp_data_gg$midpoint <- (2^size_sp_data_gg$class+2^(size_sp_data_gg$class-1))/2

size_sp_data_gg$Width <- (2^((size_sp_data_gg$class)))-((2^(size_sp_data_gg$class-1)))
size_sp_data_gg$corrected_biomass <- with(data = size_sp_data_gg,expr = sum_biomass/Width)
size_sp_data_gg$midpoint_log<- log10(size_sp_data_gg$midpoint)



ggplot(size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_line()+geom_point()


filter_size_sp_group<-size_sp_data_gg %>% filter(class>(-2))%>% filter(class<(12))

ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(filter_size_sp_group,aes(midpoint,log10(corrected_biomass),color = site_year))+geom_line()+geom_point()

```



# Size spectra by site 

```{r}
size_sp_data_gg<-individual_data

size_sp_data_gg$class <- with(size_sp_data_gg,(ceiling(log(weight_by_class,2)))) 


size_sp_data_gg <- size_sp_data_gg%>% 
  group_by(class,Site) %>% 
  dplyr::summarize(sum_biomass = sum(weight_by_class),individuals_in_class= n())

size_sp_data_gg<-na.omit(size_sp_data_gg)



size_sp_data_gg$midpoint <- (2^size_sp_data_gg$class+2^(size_sp_data_gg$class-1))/2
size_sp_data_gg$Width <- (2^((size_sp_data_gg$class)))-((2^(size_sp_data_gg$class-1)))
size_sp_data_gg$corrected_biomass <- with(data = size_sp_data_gg,expr = sum_biomass/Width)
size_sp_data_gg$midpoint_log<- log10(size_sp_data_gg$midpoint)



ggplot(size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Site))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Site))+geom_line()+geom_point()

filter_size_sp_data_gg<-size_sp_data_gg %>% filter(class>(-9))%>% filter(class<(13))

ggplot(filter_size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Site))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(filter_size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Site))+geom_line()+geom_point()
```





# Size spectra by YEAR 

```{r}
size_sp_data_gg<-individual_data

size_sp_data_gg$class <- with(size_sp_data_gg,(ceiling(log(weight_by_class,2)))) 


size_sp_data_gg <- size_sp_data_gg%>% 
  group_by(class,Year) %>% 
  dplyr::summarize(sum_biomass = sum(weight_by_class),individuals_in_class= n())

size_sp_data_gg<-na.omit(size_sp_data_gg)



size_sp_data_gg$midpoint <- (2^size_sp_data_gg$class+2^(size_sp_data_gg$class-1))/2
size_sp_data_gg$Width <- (2^((size_sp_data_gg$class)))-((2^(size_sp_data_gg$class-1)))
size_sp_data_gg$corrected_biomass <- with(data = size_sp_data_gg,expr = sum_biomass/Width)
size_sp_data_gg$midpoint_log<- log10(size_sp_data_gg$midpoint)



ggplot(size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Year))+geom_line()+geom_point()

filter_size_sp_data_gg<-size_sp_data_gg %>% filter(class>(0))%>% filter(class<(11))

ggplot(filter_size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Year))+geom_smooth(method = lm,se=F)+geom_point()


ggplot(filter_size_sp_data_gg,aes(midpoint,log10(corrected_biomass),color = Year))+geom_line()+geom_point()


```

















