---
title: "3.2 Size spectra"
author: "TG"
date: "7 2 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
```



```{r}
#install.packages("devtools")
#devtools::install_github("andrew-edwards/sizeSpectra")
library(sizeSpectra)

```

1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```


3) separate all abundance >1 to do biomass per fish estimation

```{r}
to_sep<-one_observer %>% filter(Abundance>1)
to_sep$Abundance<-as.integer(to_sep$Abundance)
sum(to_sep$Abundance)

to_sep <- to_sep %>% uncount(Abundance)
to_sep<-to_sep %>% tibble::add_column(Abundance=rep(1),.after = 10)
```


```{r}
individual_data<-one_observer %>% filter(Abundance==1)
individual_data<-bind_rows(individual_data,to_sep)
individual_data<-individual_data %>% arrange(Site,Date,Knoll)
rownames(individual_data)<-1:nrow(individual_data)
```

4) unify the fish lenghts (to be log2 scale)

```{r}
individual_data$class <- with(individual_data,(ceiling(log(Length,2))))




individual_data<-individual_data %>%
  mutate(lenght_class =(2^class)) %>% relocate(lenght_class,.after = 12)

individual_data$midpoint <- ((2^individual_data$class) + (2^(individual_data$class-1)))/2

individual_data<-individual_data %>% relocate(midpoint,.after = 13) %>% relocate(class,.after = 14)


```


5) calculate the weight (g) per individual

```{r}
individual_data<- individual_data %>% mutate("true_weight" =  a*Length^b,.after = 14) # to use only when compering years
individual_data<- individual_data %>% mutate("class_weight" =  a*midpoint^b,.after = 15)# to use only when compering sites


```




plot biomass per knoll

```{r}
biomass_knoll<-individual_data %>% group_by(Site,Knoll,year_month) %>% summarise("total_biomass_class"=sum(class_weight),"total_biomss_true" = sum(true_weight)) %>% filter(Site != "North beach") %>% filter(year_month !="2017 b")

biomass_knoll$Knoll<-as.numeric(biomass_knoll$Knoll)

biomass_knoll<-na.omit(biomass_knoll)


ggplot(biomass_knoll, aes(x=Site,y= total_biomss_true/1000 ,fill=Site)) +
  geom_boxplot()+theme_bw()+ylab("Biomass (kg)")

ggplot(biomass_knoll, aes(x=Site,y= total_biomass_class/1000 ,fill=Site)) +
  geom_boxplot()+theme_bw()+ylab("Biomass (kg)")+ scale_y_continuous(breaks=seq(0,55,10),limits = c(0,55))


ggplot(biomass_knoll, aes(x=Site,y= sum(total_biomss_true)/1000 ,fill=Site)) +
  geom_boxplot()+theme_bw()+ylab("Biomass (kg)")

```


no schools

Pseudanthias squamipinnis (Pseudanthias spp.)
Neopomacentrus miryae
Chromis viridis
Atherinidae
```{r}

biomass_knoll_noschool<-individual_data %>% 
  filter(Site != "North beach")%>%
  filter(year_month!="2017 b") %>% 
  filter(!Species_2015 %in% c("Pseudanthias spp.",
                              "Neopomacentrus miryae",
                              "Chromis viridis",
                              "Atherinidae")) %>%
  group_by(Site,Knoll,year_month) %>%
  summarise("total_biomass_class"=sum(class_weight),
            "total_biomss_true" = sum(true_weight)) 

biomass_knoll_noschool$Knoll<-as.numeric(biomass_knoll_noschool$Knoll)

biomass_knoll_noschool<-na.omit(biomass_knoll_noschool)


ggplot(biomass_knoll_noschool, aes(x=Site,y= total_biomss_true/1000 ,fill=Site)) +
  geom_boxplot()+theme_bw()+ylab("Biomass (kg)")

ggplot(biomass_knoll_noschool, aes(x=Site,y= total_biomass_class/1000 ,fill=Site)) +
  geom_boxplot()+theme_bw()+ylab("Biomass (kg)")+ 
  scale_y_continuous(breaks=seq(0,55,10),limits = c(0,55))

```

Quick check of knolls total area before and after the storm

```{r}
load("knoll_mdata.rdata")

g_knoll<-knoll_mdata %>% group_by(storm,Site) %>% summarise("total_surface_area"=sum(Surface_Area),                                                           "total_prjected_area"= sum(Projected_area),
          "n"= n())

ggplot(knoll_mdata,aes(Site,Max_depth,fill = Site))+ 
    geom_violin()+ geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5,shp = 16)

ggplot(knoll_mdata,aes(Site,Min_depth,fill = Site))+ 
    geom_violin()+ geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5,shp = 16)

ggplot(knoll_mdata,aes(Site,Hight,fill = Site))+ 
    geom_violin()+ geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5,shp = 16)

```


5) plot size distrabiotions


```{r}

weight_boxplot<-ggplot(individual_data, aes(x=Site,y=log(class_weight) ,fill=Site)) +
  geom_boxplot()+ 
  stat_summary(fun=mean, geom="point")

weight_boxplot

weight_boxplot_cut<-ggplot(individual_data, aes(x=Site,y=class_weight ,fill=Site)) +
  geom_boxplot()+ ylim(0,20)

weight_boxplot_cut 


```


6) add the month to the year and remove the North Beach


```{r}
individual_data$year_month<-ifelse(month(individual_data$Date)>6,
                                paste(individual_data$Year,"b",sep= " "),
                                paste(individual_data$Year,"a",sep= " "))

individual_data<-individual_data %>% relocate(year_month,.after = 3) %>% 
  filter(Site != "North beach")
```


# calculate size spectra by MLE method


1. for the test ill take one site at one year

```{r}

NR<-individual_data %>% filter(Site == "NR") %>% filter(Year == 2016) %>% select(Length,midpoint,class_weight,true_weight)


Caves<-individual_data %>% filter(Site == "Caves") %>%  select(Length,midpoint,class_weight,true_weight)

Caves<-na.omit(Caves)
```


```{r}

x <- subset(Caves, true_weight >= 3)$true_weight

log.x = log(x)                      # to avoid keep calculating
sum.log.x = sum(x) 
xmin = min(x)
xmax = max(x)
```



```{r}
PL.bMLE = 1/( log(min(x)) - sum.log.x/length(x)) - 1

PLB.minLL <-  nlm(negLL.PLB, p=PL.bMLE, x=x, n=length(x),
    xmin=xmin, xmax=xmax, sumlogx=sum.log.x)

PLB.bMLE = PLB.minLL$estimate


```



# 95% confidence intervals for MLE method.

PLB.minNegLL = PLB.minLL$minimum
















```{r}

xmax<-max(NR$class_weight)
xmin<-min(NR$class_weight)
x=rPLB(n=8764,xmin = xmin,xmax = xmax)


num.bins = 8  

# Llin method - plotting binned data on log-linear axes then fitting regression.

hLlin = Llin.method(x,
                    num.bins = num.bins)

gap.barplot.cust(hLlin$counts,
                 midpoints = hLlin$mids,
                 breakpoints = hLlin$breaks,
                 xlim = c(-10,max(hLlin$breaks)+50),ylim = 10000,
                 col = rep("grey", length(hLlin$counts)),
                 xaxs = "i",
                 yaxs = "i"
                 )

```





```{r}
eight.results = eightMethodsMEE(NR$class_weight, num.bins = num.bins)
eight.methods.plot(eight.results)

```




## size spectra for each site 


```{r}
for (i in unique(individual_data$Site)) {
  
 data<-individual_data %>% filter(Site == i)
 data$class <- with(data,(ceiling(log(class_weight,2)))) 
  
 size_sp_group <- data %>% 
  group_by(class,year_month) %>% 
  dplyr::summarize(sum_biomass = sum(class_weight),individuals_in_class= n())

size_sp_group<-na.omit(size_sp_group)

size_sp_group$midpoint <- ((2^size_sp_group$class)+(2^(size_sp_group$class-1)))/2
size_sp_group$Width <- (2^((size_sp_group$class)))-((2^(size_sp_group$class-1)))
size_sp_group$corrected_biomass <- with(data = size_sp_group,expr = sum_biomass/Width)
size_sp_group$midpoint_log<- log10(size_sp_group$midpoint)

filter_size_sp_group<-size_sp_group %>% filter(class>(2))%>% filter(class<(11))


print(ggplot(size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+geom_point()+ggtitle(i))


print(ggplot(filter_size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+geom_point()+ggtitle(i))



print(ggplot(size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+geom_smooth(method = lm,se=F)+geom_point()+ggtitle(i))


print(ggplot(filter_size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+geom_smooth(method = lm,se=F)+geom_point()+ggtitle(i))


  
}
```
## size spectra for each site - all years toghther


```{r}
  
 data<-individual_data 

data<-data %>% filter(Site != "North beach") %>% mutate(Knoll = as.numeric(Knoll)) %>% drop_na(Knoll) %>% filter(Year != "2017")

data$class <- with(data,(ceiling(log(class_weight,2)))) 
  
 size_sp_group <- data %>% 
  group_by(class,Site) %>% 
  dplyr::summarize(sum_biomass = sum(class_weight),individuals_in_class= n())

size_sp_group<-na.omit(size_sp_group)

size_sp_group$midpoint <- ((2^size_sp_group$class)+(2^(size_sp_group$class-1)))/2
size_sp_group$Width <- (2^((size_sp_group$class)))-((2^(size_sp_group$class-1)))
size_sp_group$corrected_biomass <- with(data = size_sp_group,expr = sum_biomass/Width)
size_sp_group$midpoint_log<- log10(size_sp_group$midpoint)

filter_size_sp_group<-size_sp_group %>% filter(class>(1))


print(ggplot(size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = Site))+geom_point())


print(ggplot(filter_size_sp_group,aes(log10(midpoint),log10(corrected_biomass)))+geom_point()+ggtitle(i))



print(ggplot(size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = Site))+geom_smooth(method = lm,se=T)+geom_point()+ggtitle(i))


ggplot(filter_size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = Site))+ylab("log 10 normalized biomass")+
  xlab("Size class midpoint (g)")+
    scale_x_continuous(breaks = seq(0,5,by=0.25),labels = function(x) round(10^(x),0))+
    geom_smooth(method = lm,se=T,alpha = 0.1)+geom_point()+theme_classic()+
    theme(axis.text.x = element_text(size=16,angle=45,hjust = 1),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title = element_text(size=18),
        legend.text = element_text(size=16),
        strip.text.x = element_text(size = 16))



print(ggplot(filter_size_sp_group,aes(midpoint,corrected_biomass,color = Site))+ylab("log 10 normalized biomass")+
  xlab("log 10 size class midpoint")+
    geom_smooth(method = lm,se=F)+geom_point())  

```
## size spectra for each year - all sites toghther


```{r}
  
data<-individual_data 

data<-data %>% filter(Site != "North beach") %>%
  mutate(Knoll = as.numeric(Knoll)) %>% drop_na(Knoll)%>% filter(Year != "2017")

data$class <- with(data,(ceiling(log(class_weight,2)))) 
  
 size_sp_group <- data %>% 
  group_by(class,year_month) %>% 
  dplyr::summarize(sum_biomass = sum(class_weight),individuals_in_class= n())

size_sp_group<-na.omit(size_sp_group)

size_sp_group$midpoint <- ((2^size_sp_group$class)+(2^(size_sp_group$class-1)))/2
size_sp_group$Width <- (2^((size_sp_group$class)))-((2^(size_sp_group$class-1)))
size_sp_group$corrected_biomass <- with(data = size_sp_group,expr = sum_biomass/Width)
size_sp_group$midpoint_log<- log10(size_sp_group$midpoint)

filter_size_sp_group<-size_sp_group %>% filter(class>(1))


print(ggplot(size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+geom_point())


print(ggplot(filter_size_sp_group,aes(log10(midpoint),log10(corrected_biomass)))+geom_point()+ggtitle(i))



ggplot(size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+geom_smooth(method = lm,se=T)+geom_point()


ggplot(filter_size_sp_group,aes(log10(midpoint),log10(corrected_biomass),color = year_month))+ 
  ylab("log 10 normalized biomass")+
  xlab("Size class midpoint (g)")+ labs(color = "Survey")+
  scale_x_continuous(breaks = seq(0,5,by=0.25),labels = function(x) round(10^(x),0))+
  geom_smooth(method = lm,se=T,alpha= 0.1)+
  geom_point()+
  scale_color_brewer(palette="Dark2")+
  theme_classic()+
  theme(axis.text.x = element_text(size=16,angle=45,hjust = 1),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title = element_text(size=18),
        legend.text = element_text(size=16))
  

+ylab("log 10 normalized biomass")+
  xlab("Size class midpoint (g)")+
    scale_x_continuous(breaks = seq(0,5,by=0.25),labels = function(x) round(10^(x),0))+
    geom_smooth(method = lm,se=T,alpha = 0.1)+geom_point()+theme_classic()+
    theme(axis.text.x = element_text(size=16,angle=45,hjust = 1),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        legend.title = element_text(size=18),
        legend.text = element_text(size=16),
        strip.text.x = element_text(size = 16))


  

```
