---
title: "4.2 community structure"
author: "TG"
date: "2 2 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github

---


```{r include=FALSE}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
```


1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r include=FALSE}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r include=FALSE}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```

3) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```


create wide data for:

1. species by 2015 categories - wide_2015 (for each site comparison)

2. species by each year resolution (i.e Species column) - wide_data (for each year comparison)

3. no transient fish for 2015 division - wide_cryptic_2015

4. no transient fish for full species division - wide_cryptic


### 1. wide_2015

```{r include=FALSE}
all_fish_2015 <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_2015<-spread(all_fish_2015,Species_2015,aboundance,fill = 0)

wide_2015<-na.omit(wide_2015)

rm(all_fish_2015)
```

### 2. wide data


```{r include=FALSE}
all_fish <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species) %>%
  summarise("aboundance"=sum(Abundance))


wide_data<-spread(all_fish,Species,aboundance,fill = 0)

wide_data<-na.omit(wide_data)

rm(all_fish)
```


### 3.wide_cryptic_2015


```{r include=FALSE}
cryp_fish_2015 <- one_observer %>% filter(Status != "T") %>% 
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_cryptic_2015<-spread(cryp_fish_2015,Species_2015,aboundance,fill = 0)

wide_cryptic_2015<-na.omit(wide_cryptic_2015)

rm(cryp_fish_2015)
```

### 4. wide_cryptic



```{r include=FALSE}

all_fish <- one_observer %>% filter(Status != "T") %>% 
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species) %>%
  summarise("aboundance"=sum(Abundance))


wide_cryptic<-spread(all_fish,Species,aboundance,fill = 0)

wide_cryptic<-na.omit(wide_cryptic)

rm(all_fish)

```


# nMDS

## for all fish

### wisconsin transformation

for each year, without 2017


```{r echo=TRUE}

wide_data_no_2017<-wide_data %>% filter(Year != "2017")

```


```{r}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_data_no_2017$year_month)) {
  
  data <- wide_data_no_2017 %>% filter(year_month == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)

    print(ord$stress)
  
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  



```

for all fish but without schools...

```{r}

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")

wide_data_no_2017$`Pseudanthias squamipinnis`<-NULL
wide_data_no_2017$`Pseudanthias spp.`<-NULL
wide_data_no_2017$`Chromis viridis`<-NULL
wide_data_no_2017$`Neopomacentrus miryae`<-NULL

for (i in unique(wide_data_no_2017$year_month)) {
  
  data <- wide_data_no_2017 %>% filter(year_month ==i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = i)
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  


```


for each site


```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
   print(ord$stress)
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}




```
for each site with no schools


```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")


wide_2015$`Pseudanthias spp.`<-NULL
wide_2015$`Chromis viridis`<-NULL
wide_2015$`Neopomacentrus miryae`<-NULL

for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress) 
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}




```

### with no transformatiom


```{r}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_data_no_2017$year_month)) {
  
  data <- wide_data_no_2017 %>% filter(year_month == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
  site <- factor(data$Site)
   print(ord$stress)
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  



```


for each site


```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
   print(ord$stress)
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}




```


### sqrt transformation


```{r}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_data_no_2017$year_month)) {
  
  data <- wide_data_no_2017 %>% filter(year_month == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- sqrt(wis_data)
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
   print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  



```


for each site


```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- sqrt(wis_data)
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
  Year <- factor(data$year_month)
 print(ord$stress) 
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}




```

### hellinger transformation


```{r}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_data_no_2017$year_month)) {
  
  data <- wide_data_no_2017 %>% filter(year_month == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- decostand(wis_data,"hellinger")
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  



```
for each site


```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- decostand(wis_data,"hellinger")
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}




```
## for cryptic species only

for each year


```{r include=FALSE}

wide_cryptic_no_2017<-wide_cryptic %>% filter(Year != "2017")

```


```{r echo=FALSE}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_cryptic_no_2017$year_month)) {
  
  data <- wide_cryptic_no_2017 %>% filter(year_month == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  

```

for each site

```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}
```

### no transformation

all sites toghether

```{r}

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")

  data <- wide_cryptic_no_2017 
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n")
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)


```
all sites toghether - no schools

```{r}

color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")

  data <- wide_cryptic_no_2017 
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n",main = paste("Stress =",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)


```

```{r echo=FALSE}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_cryptic_no_2017$year_month)) {
  
  data <- wide_cryptic_no_2017 %>% filter(year_month == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  

```

for each site

```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
 
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}



```

#Do the style


all the site beside katza

```{r echo=FALSE}

#color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")


new_6<-c("#277da1",
         "#4d194d",
         "#43aa8b",
         "#d9ed92",
         "#f94144",
         "#f8961e")

new_7<-c("#277da1",
         "#4d194d",
         "#90be6d",
         "#43aa8b",
         "#d9ed92",
         "#f94144",
         "#f8961e")


wide_cryptic_2015$year_month<-as.character(wide_cryptic_2015$year_month)
wide_cryptic_2015$year_month <-ifelse(month(wide_cryptic_2015$Date) == 6,
                                     paste0("Spring ",year(wide_cryptic_2015$Date)),
                                     paste0("Fall " ,year(wide_cryptic_2015$Date)))


wide_cryptic_2015$year_month<-as.factor(wide_cryptic_2015$year_month)
wide_cryptic_2015$year_month<-ordered(wide_cryptic_2015$year_month,levels= c("Fall 2015",
    "Fall 2016",
    "Fall 2017",
    "Fall 2018",
    "Spring 2020",
    "Fall 2020" ,
    "Spring 2021"))

for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE) 
  
  if (i == "Katza") {color_y=new_7} else
  {color_y=new_6}
  
  print(color_y)
 
  Year <- factor(data$year_month)
  
  Year<-droplevels(Year)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.9,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)



}

if (i == "Katza"){legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)}

```

```{r echo=FALSE}

#color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")


new_6<-c("#277da1",
         "#4d194d",
         "#43aa8b",
         "#d9ed92",
         "#f94144",
         "#f8961e")

new_7<-c("#277da1",
         "#4d194d",
         "#90be6d",
         "#43aa8b",
         "#d9ed92",
         "#f94144",
         "#f8961e")


wide_cryptic_2015$year_month<-as.character(wide_cryptic_2015$year_month)
wide_cryptic_2015$year_month <-ifelse(month(wide_cryptic_2015$Date) == 6,
                                     paste0("Spring ",year(wide_cryptic_2015$Date)),
                                     paste0("Fall " ,year(wide_cryptic_2015$Date)))


wide_cryptic_2015$year_month<-as.factor(wide_cryptic_2015$year_month)
wide_cryptic_2015$year_month<-ordered(wide_cryptic_2015$year_month,levels= c("Fall 2015",
    "Fall 2016",
    "Fall 2017",
    "Fall 2018",
    "Spring 2020",
    "Fall 2020" ,
    "Spring 2021"))


wide_no_k<-wide_cryptic_2015 %>% filter(Site!="Katza")
wide_no_k$year_month<-droplevels(wide_no_k$year_month)

for (i in unique(wide_no_k$Site)) {
  
  data <- wide_no_k %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE) 
  
  color_y=new_6
 
  Year <- factor(data$year_month)
  
  Year<-droplevels(Year)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.5,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)



}

wide_k<-wide_cryptic_2015 %>% filter(Site=="Katza")

for (i in unique(wide_k$Site)) {
  
  data <- wide_k %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE) 
  
  color_y=new_7
 
  Year <- factor(data$year_month)
  
  Year<-droplevels(Year)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.5,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)



}

```






all sites
```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")


  
  data <- wide_cryptic_2015 
  wis_data <- data[,9:ncol(data)]
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n")
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)


```


### sqrt transformation


```{r echo=FALSE}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(wide_cryptic_no_2017$year_month)) {
  
  data <- wide_cryptic_no_2017 %>% filter(year_month == i)
  wis_data <- data[,9:ncol(data)]
  wis_data<-sqrt(wis_data)
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  site <- factor(data$Site)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  

```

for each site

```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
 wis_data<-sqrt(wis_data)
   ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  print(ord$stress)
  Year <- factor(data$year_month)
  
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}
```









## Presence/absence nMDS

Ill try one example for P/A data to see if it makes any difference

ill take the data for all the fish



## for all fish


for each year, without 2017


```{r echo=TRUE}


pa_data<-wide_data_no_2017[,9:ncol(wide_data_no_2017)]
pa_data[pa_data>0]<-1

pa_data_no_2017<-cbind(wide_data_no_2017[,1:8],pa_data)

```


```{r}



color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")



for (i in unique(pa_data_no_2017$year_month)) {
  
  data <- pa_data_no_2017 %>% filter(year_month == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
  site <- factor(data$Site)
  print(ord$stress)
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color[site],pch = 16)
ordiellipse(ord, groups = data$Site, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color)
legend("bottomright", legend = levels(site), bty = "n", col = color, pch = 15,cex=1.5)

}
  

```
```{r}
pa_data_site<-wide_2015[,9:ncol(wide_2015)]
pa_data_site[pa_data_site>0]<-1

pa_data<-cbind(wide_2015[,1:8],pa_data_site)
```

for each site

```{r echo=FALSE}

color_y<-c("#ed6028", "#159ed2", "#16a647","#c40032","#522c19","#fcf5c1")

for (i in unique(pa_data$Site)) {
  
  data <- pa_data %>% filter(Site == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  
  Year <- factor(data$year_month)
  print(ord$stress)
ordiplot(ord, type = "n", main = paste(i,", stress=",round(ord$stress,3)))
orditorp(ord, label = T, display = "sites", col = color_y[Year],pch = 16)
ordiellipse(ord, groups = data$year_month, draw = "polygon",alpha = 0.35,label=F ,lty = 1,col = color_y)
legend("bottomright", legend = levels(Year), bty = "n", col = color_y, pch = 15,cex=1.5)

}
```