---
title: "4.3.1 RDA"
author: "TG"
date: "3 2 2021"
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(sars)
library(lme4)
library(lmerTest)
library(MuMIn)
```

1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```

3) add the month to the sample

```{r}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```



```{r}
all_fish <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface.Area,Max.depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_data_all<-spread(all_fish,Species_2015,aboundance,fill = 0)

wide_data_all<-na.omit(wide_data_all)
```

# RDA - ALL FISH

wisconsin_wide_data_all<-wide_data_all 

```{r}


wisconsin_wide_data_all<- wisconsin(wide_data_all[,9:ncol(wide_data_all)])

```


```{r}
rda_all_fish_wisconsin <- rda(wisconsin_wide_data_all  ~ Site +year_month + Surface.Area, data=wide_data_all)

rda_all_fish_wisconsin


plot(rda_all_fish_wisconsin, scaling = 2)

plot(rda_all_fish_wisconsin, type = "n",scaling =2)
abline(h = 0, v = 0, col = "white", lwd = 3)
box()
points(rda_all_fish_wisconsin, display = "species")
text(rda_all_fish_wisconsin, display = "cn", col = "blue")


```

Lets do it for each Year seperatly

2015

```{r}
data_2015 <- wide_data_all %>% filter(Year == "2015")
wisconsin_2015<- wisconsin(data_2015[,9:ncol(wide_data_all)])

rda_2015 <- rda(wisconsin_2015  ~ Site + Surface.Area+ Max.depth, data=data_2015)

rda_2015

plot(rda_2015, scaling = 2)
text(rda_2015, display = "cn", col = "blue")



plot(rda_2015, type = "n",scaling =2)
abline(h = 0, v = 0, col = "white", lwd = 3)
box()
points(rda_2015, display = "species",col = "red",)
points(rda_2015, display = "sites")
text(rda_2015, display = "cn", col = "blue")

```
Error in match.arg(display, c("sites", "species", "wa", "lc", "bp", "cn", : 'arg' should be one of “sites”, “species”, “wa”, “lc”, “bp”, “cn”, “reg”

```{r}
rda_year<-list()


wide_data_no_2017<-wide_data_all %>% filter(Year != "2017")

for (i in unique(wide_data_no_2017$year_month)) {
  
  data<- wide_data_no_2017 %>%  filter(year_month == i)
  wisconsin_data <- wisconsin(data[,9:ncol(data)])
  
  rda <- rda(wisconsin_data  ~ Site + Surface.Area + Max.depth, data= data)

  rda_year[[i]]<-rda
  
  
}

plot(rda_year[[1]], scaling = 2,main = "2015" )
text(rda_year[[1]], display = "cn", col = "blue")

plot(rda_year[[2]], scaling = 2,main = "2016")
text(rda_year[[2]], display = "cn", col = "blue")

plot(rda_year[[3]], scaling = 2,main = "2018")
text(rda_year[[3]], display = "cn", col = "blue")

plot(rda_year[[4]], scaling = 2,main = "2020 a")
text(rda_year[[4]], display = "cn", col = "blue")

plot(rda_year[[5]], scaling = 2,main = "2020 b")
text(rda_year[[5]], display = "cn", col = "blue")
```


```{r}
rm(list=setdiff(ls(), c("one_observer","wide_data_all","wide_data_no_2017")))
```
