---
title: "5.1 beta diversity"
author: "TG"
date: "28 1 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(sars)
library(lme4)
library(lmerTest)
library(MuMIn)
```



```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```

filter the north beach
```{r}
one_observer<-one_observer %>% filter(Site !="North beach")
```


3) convert to wide data

```{r}
wide_data<- one_observer %>%
  group_by(Site,Year,Knoll,Date,survey_id,Species_grouped) %>%
  summarise("aboundance"=sum(Abundance))


wide_data<-spread(wide_data,Species_grouped,aboundance,fill = 0)
```
4) split to species matrix and site matrix and convert to mobr object


```{r}
species_data<-wide_data[,6:ncol(wide_data)]
site_data<-wide_data[,1:5]
site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")

mobr_data <- make_mob_in(species_data, site_data)

```


5) rarefaction (with schools!)

```{r}
plot_rarefaction(mobr_data, 'Site', 'IBR', lwd=4)
plot_rarefaction(mobr_data, 'Site', 'SBR', lwd=4)


```

```{r}
plot_rarefaction(mobr_data, 'Year', 'IBR', lwd=4)
plot_rarefaction(mobr_data, 'Year', 'SBR', lwd=4)
```



6) Species abundance distributions (SAD)

```{r}

plot_abu(mobr_data, 'Site', type='rad', pooled=T, log='x')

```


7) calculate indexes 

by site

```{r}
site_stats <- get_mob_stats(mobr_data, group_var = "Site",
n_perm = 200)

```
• S - Observed species richness

```{r}
plot(site_stats, 'S')
```

• N - Number of individuals

```{r}
plot(site_stats, 'N')
```


• S_n - Rarefied species richness (Hurlbert 1971)

```{r}
plot(site_stats, 'S_n')
```


• S_PIE - Effective number of species based on PIE (Jost 2007)

```{r}
plot(site_stats, 'S_PIE')
```


by Year (2017 is only katza!)

```{r}
year_stats <- get_mob_stats(mobr_data, group_var = "Year",
n_perm = 200)

```
• S - Observed species richness

```{r}
plot(year_stats, 'S')
```

• N - Number of individuals

```{r}
plot(year_stats, 'N')
```


• S_n - Rarefied species richness (Hurlbert 1971)

```{r}
plot(year_stats, 'S_n')
```


• S_PIE - Effective number of species based on PIE (Jost 2007)

```{r}
plot(year_stats, 'S_PIE')
```
