---
title: "5.2 beta diversity and rarefaction"
author: "TG"
date: "7 2 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r}
load("fish_full_data.rdata")
load("survey_list.Rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
#library(sars)
library(lme4)
library(lmerTest)
library(MuMIn)
```

1) set one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 0

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <1)

```

3) filter the north beach
```{r}

one_observer<-one_observer %>%
  filter(Site !="North beach",
         ! Knoll %in% c("Tamar","South bridge reserve","south bridge reserve",
                        "North bridge reserve","north bridge reserve"))
  
```



4) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```


```{r}
knoll_survey <- survey_list %>%
  drop_na(Knoll) %>%
  filter(Site != "North beach") %>% 
  group_by(Site,Year,month(Date)) %>%
  summarise("knolls" = n()) 
 
knoll_survey$Season<-ifelse(knoll_survey$`month(Date)`==6,
                            paste0("Spring_",knoll_survey$Year),
                            paste0("Fall_",knoll_survey$Year))
knoll_survey$`month(Date)`<-NULL
knoll_survey$Year<-NULL
knoll_survey<-spread(knoll_survey,Season,knolls)


```



create wide data for:

1. species by 2015 categories - wide_2015 (for each site comparison)

2. species by each year resolution (i.e Species column) - wide_data (for each year comparison)




### 1. mobr_2015_data

species by 2015 categories - wide_2015 (for each site comparison)


```{r include=FALSE}
all_fish_2015 <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,DD_X,DD_Y,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_2015<-spread(all_fish_2015,Species_2015,aboundance,fill = 0)


```

### 2. mobr_data

species by each year resolution (i.e Species column) - wide_data (for each year comparison)


```{r include=FALSE}
all_fish <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,DD_X,DD_Y,Surface_Area,Max_depth,Date,survey_id,Species) %>%
  summarise("aboundance"=sum(Abundance))


wide_data<-spread(all_fish,Species,aboundance,fill = 0)


```



5) rarefaction grouped by site (all years together) 




```{r}
data<- wide_2015 

data<-data %>% drop_na(Surface_Area)

species_data<-data[,11:ncol(data)]
site_data<-data[,1:11]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


ggplot(site_data,aes(Site,Surface_Area,fill = Site))+geom_violin()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.4,shape=16,fill="black")+
  theme_bw()

mobr_2015_data <- make_mob_in(species_data, site_data,coord_names = c("DD_Y","DD_X"))


plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')

plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright',pooled = F)

plot_rarefaction(mobr_2015_data, 'Site', method='sSBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'Site', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "Site",n_perm = 100,
                            boot_groups = TRUE,effort_samples = 50)
# plot(site_stats,"S")
# plot(site_stats,"S_n")

pdf(file="mobr_2015_allyaersNsites.pdf",width=10, height=7)
plot(site_stats)
dev.off()


jpeg(file="mobr_2015_allyaersNsites_N.jpeg",width=600, height=350)
plot(site_stats,"N")
dev.off()

plot(site_stats,'S_n')


```

maybe remove the: 
Pseudanthias squamipinnis (Pseudanthias spp.)
Neopomacentrus miryae
Chromis viridis
Atherinidae


```{r}
data<- wide_2015 

data<- data %>% filter(Year != "2017")

data$`Pseudanthias spp.`<-NULL
data$`Chromis viridis`<-NULL
data$`Neopomacentrus miryae`<-NULL
data$Atherinidae<-NULL

species_data<-data[,11:ncol(data)]

site_data<-data[,1:11]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


ggplot(site_data,aes(Site,Surface_Area,fill = Site))+geom_violin()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.4,shape=16,fill="black")+
  theme_bw()

mobr_2015_data <- make_mob_in(species_data, site_data)


plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')
plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright',pooled = F)
plot_rarefaction(mobr_2015_data, 'Site', method='SBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'Site', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "Site",n_perm = 100,
                            boot_groups = TRUE,effort_samples = c(10,100,1000))


plot(site_stats)

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\sp_data_2015_no_schools.pdf",width=10, height=7)
plot(site_stats)
dev.off()

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\sample_rare_2015_sp_data_no_schools.pdf",width=10, height=7)
plot_rarefaction(mobr_2015_data, 'Site', method='SBR', lwd=4,leg_loc='bottomright')
dev.off()

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\Ind_rare_2015_sp_data_no_schools.pdf",width=10, height=7)
plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')
dev.off()
```
in this type of data there is a plato of species that the sites can reach to (~ 80 in the list) so it make sense most of them did...

"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots"



```{r}
data<- wide_data
data<- data %>% filter(Year != "2017")

#data$`Pseudanthias spp.`<-NULL
#data$`Chromis viridis`<-NULL
#data$`Neopomacentrus miryae`<-NULL
#data$Atherinidae<-NULL
#data$`Pseudanthias squamipinnis`<-NULL

species_data<-data[,9:ncol(data)]
site_data<-data[,1:8]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


ggplot(site_data,aes(Site,Surface_Area,fill = Site))+geom_violin()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.4,shape=16,fill="black")+
  theme_bw()

mobr_2015_data <- make_mob_in(species_data, site_data)


plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')
plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright',pooled = F)
plot_rarefaction(mobr_2015_data, 'Site', method='SBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'Site', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "Site",n_perm = 100,
                            boot_groups = TRUE,effort_samples = c(10,100,1000))
# plot(site_stats,"S")
# plot(site_stats,"S_n")


plot(site_stats,'S_n')

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\full_sp_data_with_schools.pdf",width=10, height=7)
plot(site_stats)
dev.off()

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\sample_rare_full_sp_data_with_schools.pdf",width=10, height=7)
plot_rarefaction(mobr_2015_data, 'Site', method='SBR', lwd=4,leg_loc='bottomright')
dev.off()

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\Ind_rare_full_sp_data_with_schools.pdf",width=10, height=7)
plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')
dev.off()


```
only for high resolution data:


```{r}
data<- wide_data
data<- data %>% filter(!Year %in% c("2017", "2016", "2015")) %>% drop_na(Surface_Area)


data$`Pseudanthias spp.`<-NULL
data$`Chromis viridis`<-NULL
data$`Neopomacentrus miryae`<-NULL
data$Atherinidae<-NULL
data$`Pseudanthias squamipinnis`<-NULL

species_data<-data[,12:ncol(data)]
site_data<-data[,1:11]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


ggplot(site_data,aes(Site,Surface_Area,fill = Site))+geom_violin()+
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.4,shape=16,fill="black")+
  theme_bw()

mobr_2015_data <- make_mob_in(species_data, site_data)


plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')
plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright',pooled = F)
plot_rarefaction(mobr_2015_data, 'Site', method='SBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'Site', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "Site",n_perm = 100,
                            boot_groups = TRUE,effort_samples = c(10,100,1000))
# plot(site_stats,"S")
# plot(site_stats,"S_n")

plot(site_stats)
plot(site_stats,'S_n')

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\full_sp_data_with_schools.pdf",width=10, height=7)
plot(site_stats)
dev.off()

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\sample_rare_full_sp_data_with_schools.pdf",width=10, height=7)
plot_rarefaction(mobr_2015_data, 'Site', method='SBR', lwd=4,leg_loc='bottomright')
dev.off()

pdf(file="C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\rarefaction plots\\Ind_rare_full_sp_data_with_schools.pdf",width=10, height=7)
plot_rarefaction(mobr_2015_data, 'Site', method='IBR', lwd=4,leg_loc='bottomright')
dev.off()


```

5) rarefaction for each site along the years

first all sites together

```{r}
data<- wide_2015 

data
data<- data %>% filter(Year != "2017")
species_data<-data[,9:ncol(data)]
site_data<-data[,1:8]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


mobr_2015_data <- make_mob_in(species_data, site_data)


plot_rarefaction(mobr_2015_data,'year_month', method='IBR', lwd=4,leg_loc='bottomright')
plot_rarefaction(mobr_2015_data,'year_month',method='IBR', lwd=4,leg_loc='bottomright',pooled = F)
plot_rarefaction(mobr_2015_data,'year_month', method='SBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'year_month', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "year_month",n_perm = 100,
                            boot_groups = TRUE,effort_samples = c(10,100,1000))
plot(site_stats)

# plot(site_stats,"S")
# plot(site_stats,"S_n")

pdf(file="mobr_2015_sites_with_school.pdf",width=10, height=7)
plot(site_stats)
dev.off()


jpeg(file="mobr_2015_sites_with_school.jpeg",width=600, height=350)
plot(site_stats,"N")
dev.off()

plot(site_stats,'S_n')

```


all sites together - no schools


```{r}
data<- wide_2015 

data
data<- data %>% filter(Year != "2017")

data$`Pseudanthias spp.`<-NULL
data$`Chromis viridis`<-NULL
data$`Neopomacentrus miryae`<-NULL
data$Atherinidae<-NULL



species_data<-data[,11:ncol(data)]
site_data<-data[,1:11]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


mobr_2015_data <- make_mob_in(species_data, site_data)


plot_rarefaction(mobr_2015_data,'year_month', method='IBR', lwd=4,leg_loc='bottomright')
plot_rarefaction(mobr_2015_data,'year_month',method='IBR', lwd=4,leg_loc='bottomright',pooled = F)
plot_rarefaction(mobr_2015_data,'year_month', method='SBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'year_month', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "year_month",n_perm = 100,
                            boot_groups = TRUE,effort_samples = c(10,100,1000))
plot(site_stats)


# plot(site_stats,"S")
# plot(site_stats,"S_n")

pdf(file="mobr_2015_sites_with_noschool.pdf",width=10, height=7)
plot(site_stats)
dev.off()


jpeg(file="mobr_2015_sites_with_noschool_N.jpeg",width=600, height=350)
plot(site_stats,"N")
dev.off()

plot(site_stats,'S_n')

```










```{r}
for (i in unique(wide_2015$Site)){
  
data<- wide_2015 %>% filter(Site == i)  
species_data<-data[,9:ncol(data)]

site_data<-data[,1:8]
site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")

mobr_2015_data <- make_mob_in(species_data, site_data)

print(i)
plot_rarefaction(mobr_2015_data, 'year_month', method='IBR', lwd=4)
plot_rarefaction(mobr_2015_data, 'year_month',method= 'SBR', lwd=4)
plot_abu(mobr_2015_data, 'year_month', type='rad', pooled=T, log='x')    
}


```

6) rarefaction for each year along the years

```{r}

for (i in unique(wide_data$year_month)) {
  
  data<- wide_data %>% filter(year_month== i)
  species_data<-data[,11:ncol(data)]

site_data<-data[,1:11]
site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")

mobr_data <- make_mob_in(species_data, site_data)

print(i)
plot_rarefaction(mobr_data, 'Site',method=  'IBR', lwd=4)
plot_rarefaction(mobr_data, 'Site',method=  'SBR', lwd=4)
plot_abu(mobr_data, 'Site', type='rad', pooled=T, log='x')  
  
  
}

```



7) calculate indexes 

by site

• S - Observed species richness
• N - Number of individuals
• S_n - Rarefied species richness (Hurlbert 1971)
• S_PIE - Effective number of species based on PIE (Jost 2007)




```{r}
for (i in unique(wide_2015$Site)){
  
data<- wide_2015 %>% filter(Site == i)  
species_data<-data[,9:ncol(data)]

site_data<-data[,1:8]
site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")

mobr_2015_data <- make_mob_in(species_data, site_data)

site_stats <- get_mob_stats(mobr_2015_data, group_var = "year_month",n_perm = 200)
print(i)
plot(site_stats, 'S')
plot(site_stats, 'N')
plot(site_stats, 'S_n')
plot(site_stats, 'S_PIE')
   
}
```

for each year

```{r}
for (i in unique(wide_data$year_month)) {
  
  data<- wide_data %>% filter(year_month== i)
  species_data<-data[,9:ncol(data)]

site_data<-data[,1:8]
site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")

mobr_data <- make_mob_in(species_data, site_data)

print(i)
year_stats <- get_mob_stats(mobr_data, group_var = "Site",n_perm = 200)

plot(year_stats, 'S')
plot(year_stats, 'N')
plot(year_stats, 'S_n')
plot(year_stats, 'S_PIE') 
  
  
}
```
5) rarefaction grouped by storm




```{r}
data<- wide_2015 

data<-data %>% mutate(storm = if_else(Year %in% c("2020","2021"),"After","Before"),.after = 4)

species_data<-data[,10:ncol(data)]
site_data<-data[,1:9]

site_data$Site_Knoll<-paste(site_data$Site,site_data$Knoll,sep="_")


mobr_2015_data <- make_mob_in(species_data, site_data)


plot_rarefaction(mobr_2015_data, 'storm', method='IBR', lwd=4,leg_loc='bottomright')
plot_rarefaction(mobr_2015_data, 'storm', method='IBR', lwd=4,leg_loc='bottomright',pooled = F)
plot_rarefaction(mobr_2015_data, 'storm', method='SBR', lwd=4,leg_loc='bottomright')

plot_abu(mobr_2015_data, 'storm', type='rad', pooled=T, log='x')    


site_stats <- get_mob_stats(mobr_2015_data, group_var = "storm",n_perm = 100,
                            boot_groups = TRUE,effort_samples = c(10,100,1000))

site_stats <- get_mob_stats(mobr_2015_data, group_var = "storm",n_perm = 100,
                            boot_groups = TRUE,effort_samples = 500)

plot(site_stats)





# plot(site_stats,"S")
# plot(site_stats,"S_n")

pdf(file="mobr_2015_allyaersNsites.pdf",width=10, height=7)
plot(site_stats)
dev.off()


jpeg(file="mobr_2015_allyaersNsites_N.jpeg",width=600, height=350)
plot(site_stats,"N")
dev.off()

plot(site_stats,'S_n')


```

# finel?

here ill take the full species data for years 2018-2021 *and* keep only the knolls that were sampled in the storm so the effeort will be similar.
ill keep all the sites toghther for now...


## full data no schools

```{r}
data<- wide_data %>%
  mutate("Site_Knoll" = paste(Site,Knoll,sep="_"),.after = Knoll
) %>%
  filter(Year > 2017)  %>%
  drop_na(DD_X) 

data$`Pseudanthias spp.`<-NULL
data$`Chromis viridis`<-NULL
data$`Neopomacentrus miryae`<-NULL
data$Atherinidae<-NULL

species_data<-data[,12:ncol(data)]
site_data<-data[,1:11]

mobr_data <- make_mob_in(species_data, site_data,coord_names = c("DD_Y","DD_X"))

# rarefaction curves

plot_rarefaction(mobr_data, 'year_month',method=  'IBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'SBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'sSBR', lwd=4)
plot_abu(mobr_data, 'year_month', type='rad', pooled=T, log='x') 

```


```{r}


after_storm_knoll<- one_observer %>% filter(year_month =="2020 a") %>% distinct(knoll_id)

data<- wide_data %>%
  mutate("Site_Knoll" = paste(Site,Knoll,sep="_"),.after = Knoll
) %>%
  filter(year_month %in% c("2018 b","2020 a"))  %>%
  drop_na(DD_X) %>%
  filter(Site_Knoll %in% after_storm_knoll$knoll_id)
  
species_data<-data[,12:ncol(data)]
site_data<-data[,1:11]

mobr_data <- make_mob_in(species_data, site_data,coord_names = c("DD_Y","DD_X"))

# rarefaction curves

plot_rarefaction(mobr_data, 'year_month',method=  'IBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'SBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'sSBR', lwd=4)
plot_abu(mobr_data, 'year_month', type='rad', pooled=T, log='x') 

# 2 dim stats

site_stats <- get_mob_stats(mobr_data, group_var = "year_month",n_perm = 100,boot_groups = TRUE,effort_samples = 50)

plot(site_stats)


# 3 dim

delta <- get_delta_stats(mobr_data,env_var = "year_month",type="discrete",n_perm = 20)

plot(delta)

pdf(file="delta_2018_2020a.pdf",width=10, height=7)
plot(delta)
dev.off()

```
same bur without schools
```{r}
data<- wide_data %>%
  mutate("Site_Knoll" = paste(Site,Knoll,sep="_"),.after = Knoll
) %>%
  filter(year_month %in% c("2018 b","2020 a")) %>%
  drop_na(DD_X) %>%
  filter(Site_Knoll %in% after_storm_knoll$knoll_id)

data$`Pseudanthias spp.`<-NULL
data$`Chromis viridis`<-NULL
data$`Neopomacentrus miryae`<-NULL
data$Atherinidae<-NULL
 
species_data<-data[,12:ncol(data)]
site_data<-data[,1:11]

mobr_data <- make_mob_in(species_data, site_data,coord_names = c("DD_Y","DD_X"))

plot_rarefaction(mobr_data, 'year_month',method=  'SBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'sSBR', lwd=4)
plot_abu(mobr_data, 'year_month', type='rad', pooled=T, log='x') 

# 2 dim stats

site_stats <- get_mob_stats(mobr_data, group_var = "year_month",n_perm = 100,boot_groups = TRUE,effort_samples = 50)

plot(site_stats)


# 3 dim

delta <- get_delta_stats(mobr_data,env_var = "year_month",type="discrete",n_perm = 50)

plot(delta)

pdf(file="delta_2018_2020a_no_school.pdf",width=10, height=7)
plot(delta)
dev.off()

```
without filtering only mutal knolls

```{r}


data<- wide_data %>%
  mutate("Site_Knoll" = paste(Site,Knoll,sep="_"),.after = Knoll
) %>%
  filter(year_month %in% c("2018 b","2020 a"))  %>%
  drop_na(DD_X)

data$`Pseudanthias spp.`<-NULL
data$`Chromis viridis`<-NULL
data$`Neopomacentrus miryae`<-NULL
data$Atherinidae<-NULL
  
species_data<-data[,12:ncol(data)]
site_data<-data[,1:11]

mobr_data <- make_mob_in(species_data, site_data,coord_names = c("DD_Y","DD_X"))

# rarefaction curves

plot_rarefaction(mobr_data, 'year_month',method=  'IBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'SBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'sSBR', lwd=4)
plot_abu(mobr_data, 'year_month', type='rad', pooled=T, log='x') 

# 2 dim stats

site_stats <- get_mob_stats(mobr_data, group_var = "year_month",n_perm = 100,boot_groups = TRUE,effort_samples = 50)

plot(site_stats)


# 3 dim

delta <- get_delta_stats(mobr_data,env_var = "year_month",type="discrete",n_perm = 20)

plot(delta)

pdf(file="delta_2018_2020a_all_knoll_no school.pdf",width=10, height=7)
plot(delta)
dev.off()

```




```{r}

after_storm_knoll<- one_observer %>% filter(year_month =="2020 a") %>% distinct(knoll_id)

data<- wide_data %>%
  mutate("Site_Knoll" = paste(Site,Knoll,sep="_"),.after = Knoll
) %>%
  filter(year_month %in% c("2020 a","2021 a"))  %>%
  drop_na(DD_X) %>%
  filter(Site_Knoll %in% after_storm_knoll$knoll_id)
  
species_data<-data[,12:ncol(data)]
site_data<-data[,1:11]

mobr_data <- make_mob_in(species_data, site_data,coord_names = c("DD_Y","DD_X"))

# rarefaction curves

plot_rarefaction(mobr_data, 'year_month',method=  'IBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'SBR', lwd=4)
plot_rarefaction(mobr_data, 'year_month',method=  'sSBR', lwd=4)
plot_abu(mobr_data, 'year_month', type='rad', pooled=T, log='x') 

# 2 dim stats

site_stats <- get_mob_stats(mobr_data, group_var = "year_month",n_perm = 100,boot_groups = TRUE,effort_samples = 50)

plot(site_stats)


# 3 dim

delta <- get_delta_stats(mobr_data,env_var = "year_month",type="discrete",n_perm = 20)

plot(delta)

pdf(file="delta_2020a_2021a.pdf",width=10, height=7)
plot(delta)
dev.off()
```


