---
title: "5.3 fish rarefactions"
author: "TG"
date: "23 12 2021"
output: html_document
---

# Load and preper data

```{r}
load("fish_full_data.rdata")
load("survey_list.Rdata")
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(Rarefy)
library(rareNMtests)
```

1) set one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 1

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <1)

```

3) filter the north beach
```{r}
one_observer<-one_observer %>%
  filter(Site !="North beach",
         ! Knoll %in% c("Tamar","South bridge reserve","south bridge reserve",
                        "North bridge reserve","north bridge reserve"))
                                      
```



4) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```


```{r eval=FALSE, include=FALSE}
SURVEYS<-one_observer %>% group_by(Site,year_month) %>%  summarise(knoll_number = n_distinct(Knoll))



S1<-one_observer %>%select(Site,year_month,Knoll) %>% group_by(Site,year_month) %>%summarise (uni=unique(Knoll))



#write.csv(SURVEYS,"fish_survey_list.csv")


fish_list_appendix<-fish_full_data %>% select(Species,Species_2015,Family,Status) %>% distinct(Species,.keep_all = T) %>% arrange(Species)

fish_list_appendix$Status<-ifelse(fish_list_appendix$Status=="T","Transient","Reef associated")

#write.csv(fish_list_appendix,"fish_list_appendix.csv")
```


# Create wide data for 2015

1. species by 2015 categories - wide_2015 (for each site comparison)


```{r include=FALSE}
all_fish_2015 <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_2015<-spread(all_fish_2015,Species_2015,aboundance,fill = 0)


```

### ibr models by site(vegan)  with schools

```{r}

#create the data

ibr<-one_observer %>% filter(Year !="2017") %>% group_by(Site,Species_2015) %>% summarise(abundance= sum(Abundance))

ibr<-spread(ibr,Species_2015,abundance,fill = 0)

#playing a bit with the vegan functions

ibr_max<-min(rowSums(ibr[,2:ncol(ibr)]))

rarefy(ibr[,2:ncol(ibr)],ibr_max,se=TRUE)


rare_index_steps<-rarefy(ibr[2:ncol(ibr)],se=T,sample=100)




```



### individual rarefaction curve with confidence interval pooled to the site level

```{r}

steps <- data.frame("ind" =  seq(10, ibr_max, by=10)) 
rare_list<-list()
loop<-1
site_name<-ibr$Site

for(i in 1:nrow(steps)){
  
  ind <- steps$ind[i]
  
 S<-rarefy(ibr[,2:ncol(ibr)], ind, se = TRUE)
 S<-t(S)
 S<-as.data.frame(S)
 S$Site<-site_name
 S$ind<-rep(ind)

 rare_list[[loop]]  <- S
 
 loop<-loop+1
  
}

rarefection<-bind_rows(rare_list)




ggplot(data = rarefection,aes(x=ind,y=S,color=Site))+geom_line(size=1.5)+
   geom_ribbon(aes(ymin=S-se, ymax=S+se),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Individuals number") + ggtitle("Fish")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


#ggsave("fish_ibr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")

```

#Fish sbr (with schools)


```{r}
srb_list<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$Site)) {
  
  site<- wide_2015 %>% filter(Site == i)
  
  site_richness<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=0,
                          random=200,
                          mean = F)
  
  site_richness$Site<-rep(i)
  site_richness$sample<-1:nrow(site_richness)
  
  
  srb_list[[loop]]<-site_richness
  loop<-loop+1
  
  
}

site_richness<-bind_rows(srb_list)


ggplot(site_richness,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.05, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


#ggsave("fish_sbr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```

#Fish shannon sbr (with schools)




```{r}
srb_list_shannon<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$Site)) {
  
  site<- wide_2015 %>% filter(Site == i)
  
  site_shannon<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  site_shannon$Site<-rep(i)
  site_shannon$sample<-1:nrow(site_shannon)
  
  
  srb_list_shannon[[loop]]<-site_shannon
  loop<-loop+1
  
  
}

site_richness_shannon<-bind_rows(srb_list_shannon)



ggplot(site_richness_shannon,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.05, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("fish_shannon_sbr.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```
# SHANNON FOR SITE NO SCHOOLS


```{r}

srb_list_shannon_ns<-list()
loop<-1

wide_2015_ns<-wide_2015 %>% filter(Year !="2017")

wide_2015_ns$`Pseudanthias spp.`<-NULL
wide_2015_ns$`Chromis viridis`<-NULL
wide_2015_ns$`Neopomacentrus miryae`<-NULL

for (i in unique(wide_2015_ns$Site)) {
  
  site<- wide_2015_ns %>% filter(Site == i)
  
  site_shannon_ns<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=1,
                          random=200,
                          mean = F)
  
  site_shannon_ns$Site<-rep(i)
  site_shannon_ns$sample<-1:nrow(site_shannon_ns)
  
  
  srb_list_shannon_ns[[loop]]<-site_shannon_ns
  loop<-loop+1
  
  
}

site_richness_shannon_ns<-bind_rows(srb_list_shannon_ns)


ggplot(site_richness_shannon_ns,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.05, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish - no schools")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("fish_shannon_sbr_no_schools.png", width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```

# fish SBR for the full species list

data prep

```{r}
all_fish <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species) %>%
  summarise("aboundance"=sum(Abundance))


wide_fish<-spread(all_fish,Species,aboundance,fill = 0)

wide_fish<-wide_fish %>%  filter(Year > 2017)

```

## site sbr full data

```{r}
srb_list_full<-list()
loop<-1



for (i in unique(wide_fish$Site)) {
  
  site<- wide_fish %>% filter(Site == i)
  
  site_richness_full<-Rarefy::rare_alpha(site[9:ncol(site)],
                          method="hill",
                          q=0,
                          random=100,
                          mean = F)
  
  site_richness_full$Site<-rep(i)
  site_richness_full$sample<-1:nrow(site_richness_full)
  
  
  srb_list_full[[loop]]<-site_richness_full
  loop<-loop+1
  
  
}

site_richness_full<-bind_rows(srb_list_full)


ggplot(site_richness_full,aes(x= sample,y=Rarefaction,color=Site))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.2, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = color)


ggsave("fish_sbr_full.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```

## year sbr full data

```{r}
srb_list_year_full<-list()
loop<-1



for (i in unique(wide_fish$year_month)) {
  
  year<- wide_fish %>% filter(year_month == i)
  
  year_richness_full<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=0,
                          random=100,
                          mean = F)
  
  year_richness_full$Year<-rep(i)
  year_richness_full$sample<-1:nrow(year_richness_full)
  
  
  srb_list_year_full[[loop]]<-year_richness_full
  loop<-loop+1
  
  
}

year_richness_full<-bind_rows(srb_list_year_full)


ggplot(year_richness_full,aes(x= sample,y=Rarefaction,color=Year))+geom_line(size=1.2)+
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.05, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all,name = "Year",
                           labels = c(                                    "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))


ggsave("fish_sbr_year_full.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```


# hill number by site

```{r}
renyi_profile_site<-renyi(ibr[,2:ncol(ibr)],scales = c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf), hill = T)

renyi_data<-data.frame(t(renyi_profile_site))

rownames(renyi_data)=NULL

renyi_data$q<-c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf)

colnames(renyi_data)<-c("Caves","Katza","NR","Princess","Q")

renyi_data<-gather(renyi_data,"Site","richness",1:4)

renyi_data$Q<-ordered(renyi_data$Q,levels = c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf))

ggplot(renyi_data,aes(x=Q,y=richness,color = Site))+
  geom_point(size=2)+
  scale_color_manual(values = color)+
  xlab("Diversity number")+
  ylab("Diversity")+theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  ggtitle("Fish")
  

ggsave("hill_profile_fish.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")


```




#IBR by year

ibr models by site(vegan)  with schools

```{r}

#create the data

ibr_year<-one_observer %>% filter(Year !="2017") %>% group_by(year_month,Species_2015) %>% summarise(abundance= sum(Abundance))

ibr_year<-spread(ibr_year,Species_2015,abundance,fill = 0)

#playing a bit with the vegan functions

ibr_max_year<-min(rowSums(ibr_year[,2:ncol(ibr_year)]))

rarefy(ibr_year[,2:ncol(ibr_year)],ibr_max_year,se=TRUE)


rare_index_steps<-rarefy(ibr_year[2:ncol(ibr_year)],se=T,sample=100)




```



individual rarefaction curve with confidence interval
pooled to the site level

```{r}

steps_year <- data.frame("ind" =  seq(10, ibr_max_year , by=10)) 

rare_list_year<-list()
loop<-1
Year<-ibr_year$year_month


for(i in 1:nrow(steps_year)){
  
  ind <- steps_year$ind[i]
  
 S<-rarefy(ibr_year[,2:ncol(ibr_year)], ind, se = TRUE)
 S<-t(S)
 S<-as.data.frame(S)
 S$Year<-Year
 S$ind<-rep(ind)

 rare_list_year[[loop]]  <- S
 
 loop<-loop+1
  
}

rarefection_year<-bind_rows(rare_list_year)




ggplot(data = rarefection_year,aes(x=ind,y=S,color=Year))+geom_line(size=1.5)+
   geom_ribbon(aes(ymin=S-se, ymax=S+se),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Individuals number") + ggtitle("Fish")+
  ylab("Richness")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=fish_all,
                           name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))


#ggsave("fish_ibr_year.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")

```

# fish sbr by year



```{r}

l_type<-c("solid","solid","solid","twodash","solid","solid")

srb_list_year<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$year_month)) {
  
  year<- wide_2015 %>% filter(year_month == i)
  
  year_richness<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=0,
                          random=100,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  srb_list_year[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness<-bind_rows(srb_list_year)

labs(color  = "Year", linetype = "Year")


ggplot(year_richness,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year, linetype = Year), size = 1.2) +
  scale_linetype_manual(name = "Year",values = l_type,labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021")) +
  geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Richness")+ 
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all,name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("fish_sbr_year.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")



```

### boxplot for rarefied richness

Ill rarefied to 52 knolls as this is the lowest knoll number sampled by year

```{r}
rare_sbr<-year_richness %>% filter(sample == 52)


ggplot(data = rare_sbr,aes(x=Year,y=Rarefaction,fill = Year))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=IC_low,ymax= IC_up),width =0.1)+
  scale_fill_manual(values = fish_all)+
  xlab("Year")+ 
  ylab('Rarefied richness')+
  theme_bw()+
  geom_text(aes(label = sample,y=5), size = 7,color="White")+ 
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
    scale_x_discrete(labels = c("Fall 2015", "Fall 2016","Fall 2017",
                                      "Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))
#ggsave("fish_box_plot_sbr_year.png" , width = 10, height = 8, dpi = 300,path = "C:/Users/TalG/My Drive/PHD/chapter 1 strom effect/R Scripts")
```




# fish sbr shonnon by year



```{r}
srb_list_year_sh<-list()
loop<-1

wide_2015<-wide_2015 %>% filter(Year !="2017")

for (i in unique(wide_2015$year_month)) {
  
  year<- wide_2015 %>% filter(year_month == i)
  
  year_richness_shannon<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=1,
                          random=500,
                          mean = F)
  
  year_richness_shannon$Year<-rep(i)
  year_richness_shannon$sample<-1:nrow(year_richness_shannon)
  
  
  srb_list_year_sh[[loop]]<-year_richness_shannon
  loop<-loop+1
  
  
}

year_richness_shannon<-bind_rows(srb_list_year_sh)


ggplot(year_richness_shannon,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year), size = 1.2) +
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all,name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("fish_sbr_year_shannon.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```
### boxplot rarefied shannon


```{r}

rare_sbr_shanon<-year_richness_shannon %>% filter(sample == 52)


ggplot(data = rare_sbr_shanon,aes(x=Year,y=Rarefaction,fill = Year))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=IC_low,ymax= IC_up),width =0.1)+
  scale_fill_manual(values = fish_all)+
  xlab("Year")+ 
  ylab('Shannon rarefied richness')+
  theme_bw()+geom_text(aes(label = sample,y=1.5), size = 7,color="White")+ 
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
    scale_x_discrete(labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))


ggsave("fish_box_plot_sbrshannon_year.png" , width = 10, height = 8, dpi = 300,path = "C:/Users/TalG/My Drive/PHD/chapter 1 strom effect/R Scripts")

```



# fish sbr shonnon by year - full data



```{r}
srb_list_year_sh_full<-list()
loop<-1

wide_data<-wide_fish 

for (i in unique(wide_data$year_month)) {
  
  year<- wide_data %>% filter(year_month == i)
  
  year_richness_shannon<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=1,
                          random=500,
                          mean = F)
  
  year_richness_shannon$Year<-rep(i)
  year_richness_shannon$sample<-1:nrow(year_richness_shannon)
  
  
  srb_list_year_sh_full[[loop]]<-year_richness_shannon
  loop<-loop+1
  
  
}

year_richness_shannon<-bind_rows(srb_list_year_sh_full)


ggplot(year_richness_shannon,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year), size = 1.2) +
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all[3:6],name = "Year",
                           labels = c(                                  "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))




rare_sbr_shanon<-year_richness_shannon %>% filter(sample == 52)


ggplot(data = rare_sbr_shanon,aes(x=Year,y=Rarefaction,fill = Year))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=IC_low,ymax= IC_up),width =0.1)+
  scale_fill_manual(values = fish_all[3:6])+
  xlab("Year")+ 
  ylab('Shannon rarefied richness')+
  theme_bw()+geom_text(aes(label = sample,y=1.5), size = 7,color="White")+ 
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
    scale_x_discrete(labels = c(                                    "Fall 2018","Spring 2020",                                  "Fall 2020" ,                                     "Spring 2021"))

ggsave("fish_box_plot_sbrshannon_year_full_data.png" , width = 10, height = 8, dpi = 300,path = "C:/Users/TalG/My Drive/PHD/chapter 1 strom effect/R Scripts")

#ggsave("fish_sbr_year_shannon_full_comm.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```

# fish sbr shonnon by year- no schools

```{r}
srb_list_year_sh_nosc<-list()
loop<-1

wide_2015_ns<-wide_2015 

wide_2015_ns$`Pseudanthias spp.`<-NULL
wide_2015_ns$`Chromis viridis`<-NULL
wide_2015_ns$`Neopomacentrus miryae`<-NULL


for (i in unique(wide_2015_ns$year_month)) {
  
  year<- wide_2015_ns %>% filter(year_month == i)
  
  year_richness_shannon_ns<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=1,
                          random=500,
                          mean = F)
  
  year_richness_shannon_ns$Year<-rep(i)
  year_richness_shannon_ns$sample<-1:nrow(year_richness_shannon_ns)
  
  
  srb_list_year_sh_nosc[[loop]]<-year_richness_shannon_ns
  loop<-loop+1
  
  
}

year_richness_shannon_ns<-bind_rows(srb_list_year_sh_nosc)


ggplot(year_richness_shannon_ns,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year), size = 1.2) +
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish - without schools")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all,name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("fish_sbr_year_shannon_no_s.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")
```

# fish sbr shonnon by year full data- no schools


```{r}
srb_list_year_sh_full<-list()
loop<-1

wide_data<-wide_fish 

wide_data$`Pseudanthias spp.`<-NULL
wide_data$`Chromis viridis`<-NULL
wide_data$`Neopomacentrus miryae`<-NULL

for (i in unique(wide_data$year_month)) {
  
  year<- wide_data %>% filter(year_month == i)
  
  year_richness_shannon<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=1,
                          random=500,
                          mean = F)
  
  year_richness_shannon$Year<-rep(i)
  year_richness_shannon$sample<-1:nrow(year_richness_shannon)
  
  
  srb_list_year_sh_full[[loop]]<-year_richness_shannon
  loop<-loop+1
  
  
}

year_richness_shannon<-bind_rows(srb_list_year_sh_full)


ggplot(year_richness_shannon,aes(x= sample,y=Rarefaction,color=Year))+
  geom_line(aes(color = Year), size = 1.2) +
   geom_ribbon(aes(ymin=IC_low, ymax=IC_up),alpha=0.1, linetype = 0)+
  theme_bw()+xlab("Random sample number") + ggtitle("Fish")+
  ylab("Shannon diversity")+
  theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values = fish_all[3:6],name = "Year",
                           labels = c(                                  "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021"))




rare_sbr_shanon<-year_richness_shannon %>% filter(sample == 52)


ggplot(data = rare_sbr_shanon,aes(x=Year,y=Rarefaction,fill = Year))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=IC_low,ymax= IC_up),width =0.1)+
  scale_fill_manual(values = fish_all[3:6])+
  xlab("Year")+ 
  ylab('Shannon rarefied richness')+
  theme_bw()+geom_text(aes(label = sample,y=1.5), size = 7,color="White")+ 
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
    scale_x_discrete(labels = c(                                    "Fall 2018","Spring 2020",                                  "Fall 2020" ,                                     "Spring 2021"))

ggsave("fish_box_plot_sbrshannon_year_full_data.png" , width = 10, height = 8, dpi = 300,path = "C:/Users/TalG/My Drive/PHD/chapter 1 strom effect/R Scripts")
```











Hill per year



# hill number

```{r}
renyi_profile_year<-renyi(ibr_year[,2:ncol(ibr_year)],
                          scales = c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf),
                          hill = T)

renyi_data_year<-data.frame(t(renyi_profile_year))

rownames(renyi_data_year)=NULL

renyi_data_year$q<-c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf)

colnames(renyi_data_year)<-c("Fall 2015",
                             "Fall 2016",
                             "Fall 2018",
                             "Spring 2020",
                             "Fall 2020" ,
                             "Spring 2021",
                             "Q")

renyi_data_year<-gather(renyi_data_year,"Year","richness",1:6)

renyi_data_year$Q<-ordered(renyi_data_year$Q,levels = c(0, 0.25, 0.5, 1, 2, 4, 8, 16, 32, 64, Inf))

ggplot(renyi_data_year,aes(x=Q,y=richness,color = Year))+
  geom_point(size=2)+
  scale_color_manual(values = fish_all,name = "Year",
                           labels = c("Fall 2015", "Fall 2016",
                                      "Fall 2018","Spring 2020", "Fall 2020" ,
                                      "Spring 2021","Fall 2021"))+
  xlab("Diversity number")+
  ylab("Diversity")+theme_bw()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  ggtitle("Fish")
  

ggsave("hill_profile_fish_year.png" , width = 14, height = 5, dpi = 300,path = "C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots nov 2021")


```




