---
title: "5.6 biomass per area"
author: "TG"
date: "23 12 2021"
output: html_document
---




```{r}
load("fish_full_data.rdata")
select <- dplyr::select
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)

fish_k<-c("#669bbc","#336683","#1a4b66","#003049","#ed3140","#c1121f","#710912")
fish_all<-c("#669bbc","#336683","#003049","#ed3140","#c1121f","#710912")
color<-c("#56dde0", "#33658a", "#f5af2d","#f2641a")


```


### general steps:

1) filter confidence = 1

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

biomass_area<-fish_full_data
biomass_area$Confidence[is.na(biomass_area$Confidence)]<-(0)
biomass_area<- biomass_area %>% filter(Confidence < 1)

```


2) remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

biomass_area<-biomass_area %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")

```

3) keep one observer

```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(biomass_area$survey_id)) {
  
  data <- biomass_area %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


the connection between resident fish and surface area


```{r}
only_cryptic <- one_observer %>% filter(Status != "T")

```


4) unify the fish lenghts (to be log2 scale)

```{r}
only_cryptic$class <- with(only_cryptic,(ceiling(log(Length,2))))


only_cryptic<-only_cryptic %>%
  mutate(lenght_class =(2^class)) %>% relocate(lenght_class,.after = 12)

only_cryptic$midpoint <- ((2^only_cryptic$class) + (2^(only_cryptic$class-1)))/2

only_cryptic<-only_cryptic %>% relocate(midpoint,.after = 13) %>% relocate(class,.after = 14)


```


5) calculate the weight (g) per individual

```{r}
only_cryptic<- only_cryptic %>% mutate("true_weight_ind" =  a*Length^b,.after = 14) # to use only when compering years
only_cryptic<- only_cryptic %>% mutate("class_weight_ind" =  a*midpoint^b,.after = 15)# to use only when compering sites


# the weight for all ind in the same line

only_cryptic<-only_cryptic %>%
  mutate("true_weight_total"=true_weight_ind*Abundance,
         "class_weight_total" = class_weight_ind*Abundance)
```





3) set the number of species per knoll


```{r}

biomass_knoll<- only_cryptic %>%
  group_by(Site,Knoll,year_month,Surface_Area,Max_depth,coral_cover) %>% 
  summarise("knoll_biomass_true" = sum(true_weight_total),
            "knoll_biomass_class" = sum(class_weight_total))
  
  
ggplot(data = biomass_knoll,aes(x=log(Surface_Area),y=log(knoll_biomass_class),color=Site)) +geom_point() +geom_smooth()


```

* ill leave 2007 where only kazza was sampled out

```{r}
biomass_knoll<-biomass_knoll %>% filter(Year != 2017)


```

add month 

```{r}
biomass_knoll$month<-month(biomass_knoll$Date)
biomass_knoll$month<-as.factor(biomass_knoll$month)

biomass_knoll$storm<-as.factor(biomass_knoll$storm)
biomass_knoll$Season <- ifelse(biomass_knoll$month == 6,"Spring","Fall")

biomass_knoll$Year<-as.factor(biomass_knoll$Year)
biomass_knoll$Site<-as.factor(biomass_knoll$Site)
```



remove outlier knolls

```{r}

no_nr_outliar<-biomass_knoll %>% filter(knoll_id != "Katza_19",knoll_id != "NR_17")

```

best model from previos scrips (2.4 species area)

