---
title: "6.2 knoll dissimilarity"
author: "TG"
date: "28 2 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r include=FALSE}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(sars)
library(lme4)
library(lmerTest)
library(MuMIn)
```


1) choose only one surveyor form each survey
at this point it's random! later maybe consider ranking or other method

```{r include=FALSE}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r include=FALSE}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```

3) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```


create wide data for:

1. species by 2015 categories - wide_2015 (for each site comparison)

2. no transient fish for 2015 division - wide_cryptic_2015




### 1. wide_2015

```{r include=FALSE}
all_fish_2015 <- one_observer %>%
  group_by(Site,Year,year_month,Knoll,Surface.Area,Max.depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_2015<-spread(all_fish_2015,Species_2015,aboundance,fill = 0)

wide_2015<-na.omit(wide_2015)

rm(all_fish_2015)
```


### 2.wide_cryptic_2015


```{r include=FALSE}
cryp_fish_2015 <- one_observer %>% filter(Status != "T") %>% 
  group_by(Site,Year,year_month,Knoll,Surface.Area,Max.depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_cryptic_2015<-spread(cryp_fish_2015,Species_2015,aboundance,fill = 0)

wide_cryptic_2015<-na.omit(wide_cryptic_2015)

rm(cryp_fish_2015)
```




# nMDS

## for all fish

### wisconsin

```{r}

area_list<-list()


for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col =data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- wisconsin(data_no_j[,9:ncol(data_no_j)])
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
  orditorp(ord_n, label = T, display = "sites",col =data_no_j$Knoll,pch = 16)
  
  area_nj <-ordihull(ord_n, groups = data_no_j$Knoll,
                     draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n<-summary(area_nj)
  n<-as.data.frame(n)
  n<-n[3,]
  n<-gather(n,Knoll,area_no_june,1:ncol(n))

  area_data<-left_join(p,n,by="Knoll")
  
  area_data$Site<-rep(i)

  area_list[[i]]<-area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)


```

### no transformation

```{r}

area_list<-list()


for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col =data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- data_no_j[,9:ncol(data_no_j)]
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
  orditorp(ord_n, label = T, display = "sites",col =data_no_j$Knoll,pch = 16)
  
  area_nj <-ordihull(ord_n, groups = data_no_j$Knoll,
                     draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n<-summary(area_nj)
  n<-as.data.frame(n)
  n<-n[3,]
  n<-gather(n,Knoll,area_no_june,1:ncol(n))

  area_data<-left_join(p,n,by="Knoll")
  
  area_data$Site<-rep(i)

  area_list[[i]]<-area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)

```

### square-root





```{r}

area_list<-list()


for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- sqrt(wis_data) 
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2, autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  #orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col = data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- data_no_j[,9:ncol(data_no_j)]
  wis_data_no<- sqrt(wis_data_no)
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
 # orditorp(ord_n, label = F, display = "sites",col =data_no_j$Knoll,pch = 16)
  area_nj <- ordihull(ord_n, groups = data_no_j$Knoll,
                      draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n <- summary(area_nj)
  n <- as.data.frame(n)
  n <- n[3,]
  n <- gather(n,Knoll,area_no_june,1:ncol(n))

  area_data <- left_join(p,n,by="Knoll")
  
  area_data$Site <- rep(i)

  area_list[[i]]<- area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)

```




### forth-root

```{r}

area_list<-list()


for (i in unique(wide_2015$Site)) {
  
  data <- wide_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- wis_data^0.25 
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2, autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  #orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col = data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- data_no_j[,9:ncol(data_no_j)]
  wis_data_no<- wis_data_no ^ 0.25
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
 # orditorp(ord_n, label = F, display = "sites",col =data_no_j$Knoll,pch = 16)
  area_nj <- ordihull(ord_n, groups = data_no_j$Knoll,
                      draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n <- summary(area_nj)
  n <- as.data.frame(n)
  n <- n[3,]
  n <- gather(n,Knoll,area_no_june,1:ncol(n))

  area_data <- left_join(p,n,by="Knoll")
  
  area_data$Site <- rep(i)

  area_list[[i]]<- area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)
  

```

## for core species only

##core fish only

### wisconson 

```{r}

area_list<-list()


for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- wisconsin(data[,9:ncol(data)])
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col =data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- wisconsin(data_no_j[,9:ncol(data_no_j)])
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
  orditorp(ord_n, label = T, display = "sites",col =data_no_j$Knoll,pch = 16)
  
  area_nj <-ordihull(ord_n, groups = data_no_j$Knoll,
                     draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n<-summary(area_nj)
  n<-as.data.frame(n)
  n<-n[3,]
  n<-gather(n,Knoll,area_no_june,1:ncol(n))

  area_data<-left_join(p,n,by="Knoll")
  
  area_data$Site<-rep(i)

  area_list[[i]]<-area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_cryptic_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = delta_area,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = "gam",alpha=0.1)


```

### no transformation

```{r}

area_list<-list()


for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col =data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- data_no_j[,9:ncol(data_no_j)]
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
  orditorp(ord_n, label = T, display = "sites",col =data_no_j$Knoll,pch = 16)
  
  area_nj <-ordihull(ord_n, groups = data_no_j$Knoll,
                     draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n<-summary(area_nj)
  n<-as.data.frame(n)
  n<-n[3,]
  n<-gather(n,Knoll,area_no_june,1:ncol(n))

  area_data<-left_join(p,n,by="Knoll")
  
  area_data$Site<-rep(i)

  area_list[[i]]<-area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_cryptic_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)

```

### square-root





```{r}

area_list<-list()


for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- sqrt(wis_data) 
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2, autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  #orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col = data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- data_no_j[,9:ncol(data_no_j)]
  wis_data_no<- sqrt(wis_data_no)
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
 # orditorp(ord_n, label = F, display = "sites",col =data_no_j$Knoll,pch = 16)
  area_nj <- ordihull(ord_n, groups = data_no_j$Knoll,
                      draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n <- summary(area_nj)
  n <- as.data.frame(n)
  n <- n[3,]
  n <- gather(n,Knoll,area_no_june,1:ncol(n))

  area_data <- left_join(p,n,by="Knoll")
  
  area_data$Site <- rep(i)

  area_list[[i]]<- area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_cryptic_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)

```




### forth-root

```{r}

area_list<-list()


for (i in unique(wide_cryptic_2015$Site)) {
  
  data <- wide_cryptic_2015 %>% filter(Site == i)
  wis_data <- data[,9:ncol(data)]
  wis_data <- wis_data^0.25 
  
  ord<- metaMDS(wis_data,distance = "bray" ,trace = F,k=2, autotransform = FALSE)   
  data$Knoll<-as.factor(data$Knoll)
  ordiplot(ord, type = "n",main = i)
  #orditorp(ord, label = T, display = "sites",col =data$Knoll,pch = 16)
  
  area <-ordihull(ord, groups = data$Knoll, draw = "polygon",alpha = 0.35,col = data$Knoll,label=F)
  
  p<-summary(area)
  p<-as.data.frame(p)
  p<-p[3,]
  p<-gather(p,Knoll,area,1:ncol(p))

 
 
  data_no_j <- data %>% filter(year_month != "2020 a")
  wis_data_no <- data_no_j[,9:ncol(data_no_j)]
  wis_data_no<- wis_data_no ^ 0.25
  
  ord_n<- metaMDS(wis_data_no,distance = "bray" ,trace = F,k=2,autotransform = FALSE)   
  data_no_j$Knoll<-as.factor(data_no_j$Knoll)
  
  ordiplot(ord_n, type = "n",main = i)
 # orditorp(ord_n, label = F, display = "sites",col =data_no_j$Knoll, pch = 16)
  area_nj <- ordihull(ord_n, groups = data_no_j$Knoll,
                      draw = "polygon",alpha = 0.35,col=data_no_j$Knoll,label=F)
  
  n <- summary(area_nj)
  n <- as.data.frame(n)
  n <- n[3,]
  n <- gather(n,Knoll,area_no_june,1:ncol(n))

  area_data <- left_join(p,n,by="Knoll")
  
  area_data$Site <- rep(i)

  area_list[[i]]<- area_data
  
}

area<-bind_rows(area_list)

area<-area %>% 
  mutate(Knoll = as.numeric(Knoll),Knoll_id = paste(Site,Knoll,sep = "_")) %>%
  select(Site,Knoll,Knoll_id,area,area_no_june) %>% arrange(Site,Knoll) %>%
  mutate("delta_area"= area - area_no_june)

surface<- wide_cryptic_2015 %>% mutate("Knoll_id" = paste(Site,Knoll,sep = "_")) %>% ungroup() %>%  select(Knoll_id,Surface.Area) %>% distinct(Knoll_id,.keep_all = T)

 
area<-left_join(area,surface,by="Knoll_id")

ggplot(data = area, aes(x=Surface.Area,y = area_no_june,color =Site))+
  geom_point()+ylab("Stability")+
  geom_smooth(method = 'gam',alpha=0.1)
  

```














