---
title: "8.1 dark diversity first trail"
author: "TG"
date: "26 3 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(vegan)
library(mobr)


```


for the years 2015-2017 where there is no confidance value i set the value 0

```{r}
dark_data<-fish_full_data
dark_data$Confidence[is.na(dark_data$Confidence)]<-(0)
dark_data<-dark_data %>% filter(Confidence <2)

```

 filter the north beach
```{r}
dark_data<-dark_data %>% filter(Site !="North beach")
```



 add the month to the sample

```{r include=FALSE}

dark_data$year_month<-ifelse(month(dark_data$Date)>6,
                                paste(dark_data$Year,"b",sep= " "),
                                paste(dark_data$Year,"a",sep= " "))


```


first I will build species list for each site for each year
ill do it for data collected after 2016 where fish were identified to species level

* ill use presence/absance of species form both surviors


```{r}
dark_data<- dark_data %>% filter(Year != "2015",Year !="2016", Year !="2017") %>% select(Site,year_month,Species) %>% group_by(Site,year_month,Species) %>% summarise(Abundance = n())

dark_data$Abundance<-1

dark_data<-spread(dark_data,Species,Abundance)


dark_data[is.na(dark_data)]<-0

```


apply the Beals algorithm


```{r}

beals<-beals(dark_data[,3:ncol(dark_data)],include = FALSE)


pa <- decostand(dark_data[3:243],"pa")
boxplot(as.vector(beals) ~ unlist(pa), xlab="Presence", ylab="Beals")

beals_df<-as.data.frame(beals)
d_meta<-dark_data[,1:2]

beals_data<-cbind(d_meta,beals_df)





```

see for each species if it was present or absent form a site/year along side with its Beals probability  

```{r}
g_beals<-gather(beals_data,"Species","beal",3:ncol(beals_data))
g_beals<-g_beals %>% arrange(Site,year_month)

g_pa<-gather(dark_data,"Species","pa",3:ncol(dark_data))
g_pa<-g_pa %>% arrange(Site,year_month)



dark_diversity<-left_join(g_beals,g_pa,by=c("Site","year_month","Species")) 

```

set a threshold for being included in dark diversity


```{r}

threshold <- dark_diversity %>% filter(pa == 1) %>% group_by(Species) %>% summarise(avg_beal_pr =mean(beal))

dark_diversity<- left_join(dark_diversity,threshold,by="Species")

dark_diversity$avg_plus_five<-dark_diversity$avg_beal_pr+0.05

dark_diversity$dark_sp<-ifelse(dark_diversity$pa != "1" & dark_diversity$beal > dark_diversity$avg_plus_five,
                               "dark",
                               "normal")

```

No species is define as dark!

- the common species make noise in the beals probability?
- the scale is too small?
- too little sample units

- maybe i should try using abundance data instead of p/a
- maybe i should try using only cryptic species