---
title: "9.2 fuzzyQ"
author: "TG"
date: "15 5 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(vegan)
library(mobr)
library(FuzzyQ)

```

# Data organizing



```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```

filter the north beach

```{r}
one_observer<-one_observer %>% filter(Site !="North beach")
```


add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```


add knoll id to see the surface area distrabution and filter small or big knolls

based on the mean & sd the range is 8.2-29.6 

```{r}
one_observer<- one_observer %>% mutate(site_knoll=paste(Site,Knoll,sep="_"))

knoll<- one_observer %>% distinct(site_knoll,.keep_all = T)
#hist(knoll$Surface.Area)

ggplot(data = knoll, aes(x=Surface_Area)) + 
  geom_histogram(binwidth = 2,color="black", fill="lightblue")+ theme_bw() +
  scale_x_continuous(breaks = seq(0, 70, 2),lim = c(0,70))

mean(knoll$Surface_Area,na.rm = T)
sd(knoll$Surface_Area,na.rm = T)

filter_knoll<-knoll %>% filter(between(Surface_Area,8.2,29.6))

filter_knoll %>% group_by(Site) %>% summarise(n=n())


```

filter the data...

```{r}

fuzzy_prep<-one_observer %>% filter(site_knoll %in% filter_knoll$site_knoll)

fuzzy_prep<- fuzzy_prep %>%
  group_by(site_knoll,year_month,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


fuzzy_prep<-spread(fuzzy_prep,Species_2015,aboundance,fill = 0)
```

test for 1 year only

```{r eval=FALSE, include=FALSE}
fuzzy_prep_2015<-fuzzy_prep %>% filter(year_month=="2015 b")

matric_fuzzy<-fuzzy_prep_2015[,3:ncol(fuzzy_prep_2015)]

fuzzy <- fuzzyq(matric_fuzzy)

paste(name,"fuzz",sep = "_")<-fuzzyq(matric_fuzzy)

AOplot(fuzzy)
AOplot(fuzzy,log.y = T)
AOplot(fuzzy,log.y = T,log.x = T)

fuzzy_boot <- fuzzyqBoot(matric_fuzzy)
fuzzy_boot_global <- fuzzyqBoot(matric_fuzzy,level = "global")
fuzzyci <- fuzzyqCI(fuzzy_boot)
sort<-sortClus(fuzzyci,fuzzy)


spp <- fuzzy$spp
col.RC <- c("brown2", "turquoise3") 

plot(spp[, 3], cex.axis = 0.8, xaxt= 'n', ylab = "Commoness index",
ylim = c(0, max(sort)), xlab = "Species", col = col.RC[spp[, 1] + 1],
pch = 16, cex = 0.8, las = 1)
ebar.int <- seq_len(nrow(spp))
arrows(ebar.int, sort["Lower", ], ebar.int, sort["Upper", ],
length= 0, col = col.RC[spp[, 1] + 1])
axis(1, at = ebar.int, labels = rownames(spp), las = 2, cex.axis = 0.6)

fuzzy_sp_index<-fuzzy[["spp"]]
fuzzy_sp_index<-fuzzy_sp_index %>% as_data_frame(fuzzy_sp_index, rownames = "Species")
fuzzy_sp_index<-fuzzy_sp_index %>% arrange(sil_width)
fuzzy_sp_index$cluster<-as.factor(fuzzy_sp_index
                                  $cluster)



ggplot(data = fuzzy_sp_index[,],aes(x= reorder(Species,sil_width),y= sil_width,fill = cluster))+geom_bar(stat = "identity")+ theme(axis.text.x = element_text(angle = 90))

ggplot(data = fuzzy_sp_index[1:30,],aes(x= reorder(Species,sil_width),y= sil_width,fill = cluster))+geom_bar(stat = "identity")+ theme(axis.text.x = element_text(angle = 90))

boxplot(fuzzy_boot_global[["fq.rep"]])


```

loop it for all year:


```{r}

loop<-1
fuz_list<-list()

for (i in unique(fuzzy_prep$year_month)) {
  
  
  one_yr<- fuzzy_prep %>% filter(year_month == i)
  name <- unique(one_yr$year_month)
  
  one_yr<-one_yr[,3:ncol(one_yr)]

  
  fuzzy <- fuzzyq(one_yr)
  
  data<-fuzzy[["spp"]]

  data<-data %>% as_data_frame(data, rownames = "Species") %>% mutate("year"=rep(name))
  data<-data %>% arrange(sil_width)
  data$cluster<-as.factor(data$cluster)
  
  fuz_list[[loop]]<-data
  loop<-loop+1
  assign(paste("fuzzyQ",name,sep = "_"),fuzzy) 
  
}

all_fuzz<-bind_rows(fuz_list)

all_fuzz<-all_fuzz %>% separate(year,c("year","month"))
all_fuzz$month<-ifelse(all_fuzz$month == "a","10_1","6_1")
all_fuzz<-all_fuzz %>% unite(year_month,year,month,sep = "_")
all_fuzz$year_month<-ymd(all_fuzz$year_month)
all_fuzz<-all_fuzz %>% arrange(Species,year_month)



ggplot(all_fuzz[1:90,],aes(x=year_month,y=Common.I,color = Species))+
  geom_line()+
  geom_point()+
 geom_hline(yintercept=0.5, linetype="dashed")

ggplot(all_fuzz[91:198,],aes(x=year_month,y=Common.I,color = Species))+
  geom_line()+
  geom_point()+
 geom_hline(yintercept=0.5, linetype="dashed")



all_fuzz$cluster <-as.character(all_fuzz$cluster)
all_fuzz$cluster <-as.numeric(all_fuzz$cluster)
all_fuzz<-all_fuzz %>% filter(year_month != "2017-06-01") %>% group_by(Species) %>% mutate("group_sum"= sum(cluster)) 

all_fuzz$group_sum<-ifelse(all_fuzz$group_sum == 0,"allways rare",
                           ifelse(all_fuzz$group_sum == 5,"allways common","shift"))

ggplot(all_fuzz,aes(x=year_month,y=Common.I,color = group_sum,group = Species))+
  geom_line(aes(group = Species))+
  geom_point()+ 
 geom_hline(yintercept=0.5, linetype="dashed")


ggplot(all_fuzz,aes(x=year_month,y=Common.I,color = group_sum))+
  geom_point(aes(group = Species))+geom_jitter()+
 geom_smooth()+
 geom_hline(yintercept=0.5, linetype="dashed")


shift_sp<-all_fuzz %>% filter(group_sum =="shift") %>% filter(year_month!="2017-06-01")

ggplot(shift_sp[1:25,],aes(x=year_month,y=Common.I,color = Species))+
  geom_line()+
  geom_point()+
 geom_hline(yintercept=0.5, linetype="dashed")

ggplot(shift_sp[26:50,],aes(x=year_month,y=Common.I,color = Species))+
  geom_line()+
  geom_point()+
 geom_hline(yintercept=0.5, linetype="dashed")

ggplot(shift_sp[51:75,],aes(x=year_month,y=Common.I,color = Species))+
  geom_line()+
  geom_point()+
 geom_hline(yintercept=0.5, linetype="dashed")

ggplot(shift_sp[76:105,],aes(x=year_month,y=Common.I,color = Species))+
  geom_line()+
  geom_point()+
 geom_hline(yintercept=0.5, linetype="dashed")




sp_stat_list <- all_fuzz %>% select(Species,group_sum) %>% distinct(Species,.keep_all = T) 

sp_stat_list %>% group_by(group_sum) %>% summarise(n=n())
```

find out which is the most common species in each year

```{r}

all_fuzz<-all_fuzz %>% arrange(year_month,Common.I,decreasing = FALSE)

```

