---
title: "9.4 fuzzyq for proposal"
author: "TG"
date: "21 5 2021"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---



```{r}
load("fish_full_data.rdata")
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(vegan)
library(mobr)
library(FuzzyQ)
library(forcats)
library(lme4)
library(lmerTest)
library(MuMIn)

path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots for research proposal"
```



```{r}
one_observer<-list()
loop<-1
set.seed(1)
for (i in unique(fish_full_data$survey_id)) {
  
  data <- fish_full_data %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

one_observer<- bind_rows(one_observer)

rm(data,filter_data)
```


2) filter confidence = 2

for the years 2015-2017 where there is no confidance value i set the value 0

```{r}

one_observer$Confidence[is.na(one_observer$Confidence)]<-(0)
one_observer<-one_observer %>% filter(Confidence <2)

```

filter the north beach
```{r}
one_observer<-one_observer %>% filter(Site !="North beach")
```


4) add the month to the sample

```{r include=FALSE}

one_observer$year_month<-ifelse(month(one_observer$Date)>6,
                                paste(one_observer$Year,"b",sep= " "),
                                paste(one_observer$Year,"a",sep= " "))


```

convert to wide data

```{r}
wide_data<- one_observer %>%
  group_by(Site,year_month,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


wide_data<-spread(wide_data,Species_2015,aboundance,fill = 0)



```


# FuzzyQ


```{r}
matric_fuzzy<-wide_data[,3:ncol(wide_data)]

fuzzy <- fuzzyq(matric_fuzzy)


AOplot(fuzzy,log.y = T)


fuzzy_boot <- fuzzyqBoot(matric_fuzzy)
global_fuzz_boot<-fuzzyqBoot(matric_fuzzy,level = "global")
fuzzyci <- fuzzyqCI(fuzzy_boot)
sort<-sortClus(fuzzyci,fuzzy)

fuzzy_sp_index <- fuzzy$spp
col.RC <- c("#017977","#593a8c") # two colors to plot rare and common

fuzzy_sp_index<-fuzzy_sp_index %>% arrange(sil_width)
spp<-fuzzy$spp


ebar.int <- seq_len(nrow(spp))

#pdf(file="fuzzyq_commoness_index.pdf", width = 10, height = 7)

par(mar = c(12, 5, 5, 5)) 
plot(spp[, 3] ,cex.axis = 0.8, xaxt= 'n', ylab = "Commoness index",
ylim = c(0, max(sort)), xlab = "", col = col.RC[spp[, 1] + 1],
pch = 16, cex = 0.8, las = 1)
arrows(ebar.int, sort["Lower", ], ebar.int, sort["Upper", ],length= 0, col = col.RC[spp[, 1] + 1])
axis(1, at = ebar.int, labels = rownames(spp), las = 2, cex.axis = 0.8)


#dev.off()



fuzzy_sp_index<-fuzzy[["spp"]]
fuzzy_sp_index<-fuzzy_sp_index %>% as_data_frame(fuzzy_sp_index, rownames = "Species")
fuzzy_sp_index<-fuzzy_sp_index %>% arrange(sil_width)
fuzzy_sp_index$cluster<-as.factor(fuzzy_sp_index
                                  $cluster)

fuzzy_sp_index$cluster<-as.factor(fuzzy_sp_index
                                  $cluster)

fuzzy_sp_index<-fuzzy_sp_index %>% arrange(sil_width)

fuzzy_sp_index %>% 
  mutate(ordering =  as.numeric(cluster) + sil_width,
         Species = fct_reorder(Species, ordering, .desc = T)) %>% 
  ggplot(aes(Species, sil_width, fill = cluster)) + geom_col()+
  theme(axis.text.x = element_text(angle = 70, hjust = 1))


boot_spp<- as.data.frame(fuzzy_boot[["fq.rep"]])

boot_glob<-as.data.frame(global_fuzz_boot[["fq.rep"]])
boot_glob_long<-boot_glob %>% gather("var","value",1:6)

boot_glob_long %>% filter(var %in% c("silw.rar","silw.all","silw.com")) %>% ggplot(aes(var,value,fill=var))+geom_boxplot()

boot_glob_long %>% filter(var %in% c("commI.rar","commI.com")) %>% ggplot(aes(var,value,fill=var))+geom_boxplot()+theme_classic()+
  scale_fill_manual(values=c("#593a8c","#017977"),labels = c("Common species", "Rare species"))+
  xlab("")+
  ylab("Average commonness indices")+
  theme(axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=14),
        legend.title = element_blank(),
        legend.text = element_text(size = 12))+
  scale_x_discrete(labels=c("commI.com" = "Common species", "commI.rar" = "Rare species"))
  

ggsave("ave_commI.pdf",path = path )

ggsave("ave_commI.png",path = path )

boot_glob_long %>% filter(var %in% c("N.Dunn")) %>% ggplot(aes(var,value))+geom_boxplot(fill="#911946")+theme_classic()+
 xlab("")+
  ylab("Normalized Dunn's coefficient")+
  theme(axis.title.x = element_blank(),
        axis.title.y = element_text(size=16),
        axis.text.x = element_blank(),
        axis.text.y = element_text(size=14),
        legend.title = element_blank(),
        legend.text = element_text(size = 14))

#ggsave("dunn.pdf",path = path )

#ggsave("dunn.png",path = path )


boot_glob_long %>% group_by(var) %>% summarise(min = min(value),max=max(value))

```


make alist 


```{r}
fuzzy_sp_list<-fuzzy_sp_index %>% select(Species,cluster,Common.I) %>% rename(Species_2015 = Species, part_of_common = Common.I ) 
fuzzy_sp_list<-fuzzy_sp_list%>% mutate( part_of_rare = 1 - part_of_common)


```

clean the env

```{r}
rm(list=setdiff(ls(), c("fish_full_data","fuzzy_sp_list","fuzzy_sp_index","one_observer")))
```

connect the cluster to the data-frame

```{r}

one_observer<-left_join(one_observer,fuzzy_sp_list,by="Species_2015")

```



give weight richness and diversity for all knolls

```{r}

weighted_richness<-one_observer %>%  select(knoll_id,Site,year_month,Species_2015,Abundance,part_of_common,part_of_rare)

weighted_richness<-weighted_richness %>% filter(knoll_id != "NR_South bridge reserve",knoll_id != "NR_north bridge reserve",knoll_id != "NR_south bridge reserve",knoll_id != "NR_North bridge reserve",knoll_id !="NR_Tamar")

weighted_richness<-weighted_richness %>% group_by(knoll_id,Site,year_month,Species_2015,part_of_common,part_of_rare) %>% summarise("abundance"=sum(Abundance))

weighted_richness<-weighted_richness %>% mutate("common_abundance" = abundance*part_of_common,"rare_abundance" = abundance*part_of_rare)


weighted_richness <- weighted_richness %>%
  group_by(year_month,Site,knoll_id) %>%  
  summarise("common_ind_number"= sum(common_abundance),
            "rare_ind_number"= sum(rare_abundance),
            "common_richness" = sum(part_of_common),
            "rare_richness" = sum(part_of_rare),
            "common_shannon" = diversity(common_abundance, index = "shannon") ,
            "rare_shannon" = diversity(rare_abundance, index = "shannon"))%>%
  arrange(Site,knoll_id,year_month) %>% group_by(knoll_id)%>%
  mutate("delta_common_richness"=lag(common_richness)-common_richness,
         "delta_rare_richness"=lag(rare_richness)-rare_richness,
         "delta_common_shannon"=lag(common_shannon)-common_shannon,
         "delta_rare_shannon"=lag(rare_shannon)-rare_shannon)
```

log ratio between 2018-2016 and 2020-2018

```{r}
log_ratio <- weighted_richness %>% filter(year_month != "2015 b",
                                          year_month != "2017 b",
                                          year_month != "2020 a") %>% 
  select(-delta_common_richness,-delta_rare_richness,-delta_common_shannon,-delta_rare_shannon,
         -common_ind_number,-rare_ind_number)


ratio_list<-list()
loop<-1

for (i in unique(log_ratio$knoll_id)){
  one_knoll<- log_ratio %>% filter(knoll_id== i)
  
  if(nrow(one_knoll) == 3) {
  
  data<-data.frame("Site" = unique(one_knoll$Site),"knoll_id" = unique(one_knoll$knoll_id))
  
  # common 
  
  data$common_rich_ratio_after <- one_knoll$common_richness[one_knoll$year_month == "2020 b"]/one_knoll$common_richness[one_knoll$year_month == "2018 b"] 
  
   data$common_rich_ratio_before <-one_knoll$common_richness[one_knoll$year_month == "2018 b"]/one_knoll$common_richness[one_knoll$year_month == "2016 b"] 
  
   data$common_shan_ratio_after <- one_knoll$common_shannon[one_knoll$year_month == "2020 b"]/one_knoll$common_shannon[one_knoll$year_month == "2018 b"] 
  
   data$common_shan_ratio_before <-one_knoll$common_shannon[one_knoll$year_month == "2018 b"]/one_knoll$common_shannon[one_knoll$year_month == "2016 b"]  
  
    # rare
   
    data$rare_rich_ratio_after <- one_knoll$rare_richness[one_knoll$year_month == "2020 b"]/one_knoll$rare_richness[one_knoll$year_month == "2018 b"] 
  
   data$rare_rich_ratio_before <-one_knoll$rare_richness[one_knoll$year_month == "2018 b"]/one_knoll$rare_richness[one_knoll$year_month == "2016 b"] 
  
   data$rare_shan_ratio_after <- one_knoll$rare_shannon[one_knoll$year_month == "2020 b"]/one_knoll$rare_shannon[one_knoll$year_month == "2018 b"] 
  
   data$rare_shan_ratio_before <-one_knoll$rare_shannon[one_knoll$year_month == "2018 b"]/one_knoll$rare_shannon[one_knoll$year_month == "2016 b"]  
   
   ratio_list[[loop]] <- data
   loop<-loop+1

   } 
   
}


ratio_data<-bind_rows(ratio_list)


  
``` 

richness plots


```{r}
path<-"C:\\Users\\TalG\\Documents\\PhD\\Eco camp\\R Scripts\\plots for research proposal"


#on log scale
# after

ggplot(ratio_data,aes(x=log(common_rich_ratio_after)))+
  geom_histogram(fill = "#593a8c",
                 color = 'black',
                 alpha = 0.7)+
  geom_histogram(aes(x=log(rare_rich_ratio_after)),
                 fill = "#017977",
                 color = "black",
                 alpha = 0.6)+
  theme_classic()+
  geom_vline(xintercept = 0, linetype="dashed", color = "black", size=1) +
  geom_vline(xintercept = mean(log(ratio_data$common_rich_ratio_after)),
             linetype="dashed", color = "#c18cbf",size=1)+
  geom_vline(xintercept = mean(log(ratio_data$rare_rich_ratio_after)),
             linetype="dashed", color = "#88c76d", size=1)+
  ggtitle("After the storm") +
  xlab("Log(species richness ratio)") + 
  ylab("Knolls frequency")+
    theme(plot.title = element_text(size=18, face="bold.italic"),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=16))

ggsave("log_frq_ratio_after.pdf",path = path )
ggsave("log_frq_ratio_after.png",path = path )
  

mean(log(ratio_data$common_rich_ratio_after))
mean(log(ratio_data$rare_rich_ratio_after))     

median(log(ratio_data$common_rich_ratio_after))
median(log(ratio_data$rare_rich_ratio_after))

# on log scale 
# before

ggplot(ratio_data,aes(x=log(common_rich_ratio_before)))+
  geom_histogram(fill = "#593a8c",
                 color = 'black',
                 alpha = 0.6)+
  geom_histogram(aes(x=log(rare_rich_ratio_before)),
                 fill = "#017977",
                 color = "black",
                 alpha = 0.6)+
  theme_classic()+
  geom_vline(xintercept = 0, linetype="dashed",color = "black", size=1.1)+
  geom_vline(xintercept = mean(log(ratio_data$common_rich_ratio_before)),
             linetype="dashed", color = "#c18cbf",size=1)+
  geom_vline(xintercept = mean(log(ratio_data$rare_rich_ratio_before)),
             linetype="dashed", color = "#88c76d", size=1)+
  ggtitle("Before the storm") +
  xlab("Log(species richness ratio)") + 
  ylab("Knolls frequency")+
  theme(plot.title = element_text(size=18, face="bold.italic"),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=16))



ggsave("log_frq_ratio_before.pdf",path = path )
ggsave("log_frq_ratio_before.png",path = path )
 
mean(log(ratio_data$common_rich_ratio_before))
mean(log(ratio_data$rare_rich_ratio_before))

median(log(ratio_data$common_rich_ratio_before))
median(log(ratio_data$rare_rich_ratio_before))

```


```{r}


# regular scale
# after

ggplot(ratio_data,aes(x=common_rich_ratio_after))+
  geom_histogram(fill = "#593a8c",
                 color = 'black',
                 alpha = 0.7)+
  geom_histogram(aes(x=rare_rich_ratio_after),
                 fill = "#017977",
                 color = "black",
                 alpha = 0.6)+
  theme_classic()+
  geom_vline(xintercept = 1, linetype="dashed", color = "black", size=1) +
  ggtitle("After the storm") +
  xlab("Species richness ratio") + 
  ylab("Knolls frequency")+
    theme(plot.title = element_text(size=18, face="bold.italic"),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=16))

    

#ggsave("frq_ratio_after.pdf",path = path )
#ggsave("frq_ratio_after.png",path = path )

# regular scale 
# before

ggplot(ratio_data,aes(x=common_rich_ratio_before),)+
  geom_histogram(fill = "#593a8c",
                 color = 'black',
                 alpha = 0.6)+
  geom_histogram(aes(x=rare_rich_ratio_before),
                 fill = "#017977",
                 color = "black",
                 alpha = 0.6)+
  theme_classic()+
  geom_vline(xintercept = 1,color = "black" ,linetype="dashed",color = "black", size=1.1)+
  ggtitle("Before the storm") +
  xlab("Species richness ratio") +
  ylab("Knolls frequency")+
  theme(plot.title = element_text(size=16, face="bold.italic"),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=14))

#ggsave("frq_ratio_before.pdf",path = path )
#ggsave("frq_ratio_before.png",path = path )
 
```


shannon plots

```{r}
ggplot(ratio_data,aes(x=log(common_shan_ratio_after)))+
  geom_histogram(fill = "#593a8c",
                 color = 'black',
                 alpha = 0.7)+
  geom_histogram(aes(x=log(rare_shan_ratio_after)),
                 fill = "#017977",
                 color = "black",
                 alpha = 0.6)+
  theme_classic()+
  geom_vline(xintercept = 0, linetype="dashed", color = "black", size=1) +
  geom_vline(xintercept = mean(log(ratio_data$common_shan_ratio_after)),
             linetype="dashed", color = "#c18cbf",size=1)+
  geom_vline(xintercept = mean(log(ratio_data$rare_shan_ratio_after)),
             linetype="dashed", color = "#88c76d", size=1)+
  ggtitle("After the storm") +
  xlab("Log(Shannon diversity ratio)") + 
  ylab("Knolls frequency")+
    theme(plot.title = element_text(size=18, face="bold.italic"),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=16))
    

ggsave("log_frq_shannon_ratio_after.pdf",path = path )
ggsave("log_frq_shannon_ratio_after.png",path = path )

mean(log(ratio_data$common_shan_ratio_after))
mean(log(ratio_data$rare_shan_ratio_after))

median(log(ratio_data$common_shan_ratio_after))
median(log(ratio_data$rare_shan_ratio_after))

ggplot(ratio_data,aes(x=log(common_shan_ratio_before)))+
  geom_histogram(fill = "#593a8c",
                 color = 'black',
                 alpha = 0.7)+
  geom_histogram(aes(x=log(rare_shan_ratio_before)),
                 fill = "#017977",
                 color = "black",
                 alpha = 0.6)+
  theme_classic()+
  geom_vline(xintercept = 0, linetype="dashed", color = "black", size=1) +
  geom_vline(xintercept = mean(log(ratio_data$common_shan_ratio_before)),
             linetype="dashed", color = "#c18cbf",size=1)+
  geom_vline(xintercept = mean(log(ratio_data$rare_shan_ratio_before)),
             linetype="dashed", color = "#88c76d", size=1)+
  ggtitle("Before the storm") +
  xlab("Log(Shannon diversity ratio)") + 
  ylab("Knolls frequency")+
    theme(plot.title = element_text(size=18, face="bold.italic"),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=16))
    

ggsave("log_frq_shannon_ratio_before.pdf",path = path )
ggsave("log_frq_shannon_ratio_before.png",path = path )

mean(log(ratio_data$common_shan_ratio_before))
mean(log(ratio_data$rare_shan_ratio_before))

median(log(ratio_data$common_shan_ratio_before))
median(log(ratio_data$rare_shan_ratio_before))
```

  
make species list - appendix 1 research proposal

```{r}

species_list <- fish_full_data %>% 
  select(Family,Species,Species_2015) %>%
  distinct(Species,.keep_all = T) %>% arrange(Family,Species)

write.csv(species_list,"species_list_full.csv")



```

